### 2017-03-16 snippets/qcoh.tex
Proof of the quasicoherence criterion in the big Zariski topos

\documentclass[10pt,reqno]{amsart}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{etex}
\usepackage{amsmath,amsthm,amssymb,stmaryrd,color,graphicx,multirow}
\usepackage{mathtools}
\usepackage{setspace}
\usepackage{bussproofs}
\usepackage{xspace}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{array}
\usepackage[protrusion=true,expansion=true]{microtype}
\usepackage[bookmarksdepth=2,pdfencoding=auto]{hyperref}
\usepackage[all]{xy}
\usepackage{soul}\setul{0.3ex}{}

\usepackage{tikz}
\usetikzlibrary{calc,shapes.callouts,shapes.arrows}
\newcommand{\hcancel}[5]{%
    \tikz[baseline=(tocancel.base)]{
        \node[inner sep=0pt,outer sep=0pt] (tocancel) {#1};
        \draw[red, line width=0.3mm] ($(tocancel.south west)+(#2,#3)$) -- ($(tocancel.north east)+(#4,#5)$);
    }%
}

\usepackage[natbib=true,style=numeric]{biblatex}
\usepackage[babel]{csquotes}
\bibliography{bibliography}

\theoremstyle{definition}
\newtheorem{defn}{Definition}[section]
\newtheorem{ex}[defn]{Example}

\theoremstyle{plain}

\newtheorem{prop}[defn]{Proposition}
\newtheorem{cor}[defn]{Corollary}
\newtheorem{lemma}[defn]{Lemma}
\newtheorem{thm}[defn]{Theorem}

\theoremstyle{remark}
\newtheorem{rem}[defn]{Remark}
\newtheorem{question}[defn]{Question}

\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\FF}{\mathbb{F}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\A}{\mathcal{A}}
\renewcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\renewcommand{\G}{\mathcal{G}}
\renewcommand{\H}{\mathcal{H}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\N}{\mathcal{N}}
\newcommand{\M}{\mathcal{M}}
\renewcommand{\L}{\mathcal{L}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\GG}{\mathbb{G}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\ppp}{\mathfrak{p}}
\newcommand{\qqq}{\mathfrak{q}}
\newcommand{\mmm}{\mathfrak{m}}
\newcommand{\nnn}{\mathfrak{n}}
\newcommand{\Hom}{\mathrm{Hom}}
\newcommand{\HOM}{\mathcal{H}\mathrm{om}}
\newcommand{\id}{\mathrm{id}}
\newcommand{\GL}{\mathrm{GL}}
\newcommand{\placeholder}{\underline{\quad}}
\let\oldul\ul
\renewcommand{\ul}[1]{\text{\oldul{$#1$}}}
\newcommand{\Set}{\mathrm{Set}}
\newcommand{\Grp}{\mathrm{Grp}}
\newcommand{\Vect}{\mathrm{Vect}}
\newcommand{\Sh}{\mathrm{Sh}}
\newcommand{\PSh}{\mathrm{PSh}}
\newcommand{\Zar}{\mathrm{Zar}}
\newcommand{\Et}{\mathrm{\acute{E}t}}
\newcommand{\Sch}{\mathrm{Sch}}
\newcommand{\Mod}{\mathrm{Mod}}
\newcommand{\Alg}{\mathrm{Alg}}
\newcommand{\Ring}{\mathrm{Ring}}
\newcommand{\LRL}{\mathrm{LRL}}
\newcommand{\pt}{\mathrm{pt}}
\newcommand{\tors}{\mathrm{tors}}
\DeclareMathOperator{\Spec}{Spec}
\newcommand{\RelSpek}{\operatorname{\ul{\mathrm{Spec}}}}
\newcommand{\QcohSpec}[2]{\mathrm{Spec}^{\mathrm{qcoh}}_{#1}{#2}}
\newcommand{\op}{\mathrm{op}}
\DeclareMathOperator{\colim}{colim}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\Ann}{Ann}
\DeclareMathOperator{\Int}{int}
\DeclareMathOperator{\Clos}{cl}
\DeclareMathOperator{\Kernel}{ker}
\DeclareMathOperator{\supp}{supp}
\newcommand{\Ass}{\mathrm{Ass}}
\newcommand{\Open}{\T}
\newcommand{\?}{\,{:}\,}
\renewcommand{\_}{\mathpunct{.}\,}
\newcommand{\speak}[1]{\ulcorner\text{\textnormal{#1}}\urcorner}
\newcommand{\Ll}{\vcentcolon\Longleftrightarrow}
\newcommand{\notat}[1]{{!#1}}
\newcommand{\lra}{\longrightarrow}
\newcommand{\lhra}{\ensuremath{\lhook\joinrel\relbar\joinrel\rightarrow}}
\newcommand{\hra}{\hookrightarrow}
\newcommand{\brak}[1]{{\llbracket{#1}\rrbracket}}
\newcommand{\sdense}{{\widehat\Box}}
\newcommand{\sdenseother}{\Box}
\newcommand{\ie}{i.\,e.\@\xspace}
\newcommand{\eg}{e.\,g.\@\xspace}
\newcommand{\vs}{vs.\@\xspace}
\newcommand{\resp}{resp.\@\xspace}
\newcommand{\inv}{inv.\@}
\newcommand{\notnot}{\emph{not not}\xspace}
\newcommand{\affl}{\ensuremath{{\ul{\AA}^1_S}}\xspace}
\newcommand{\afflx}{\ensuremath{{\ul{\AA}^1_X}}\xspace}
\newcommand{\affla}{\ensuremath{{\ul{\AA}^1_{\Spec A}}}\xspace}
\newcommand{\xra}{\xrightarrow}

\newcommand{\XXX}[1]{\textbf{XXX: #1}}

\newcommand{\defeq}{\vcentcolon=}
\newcommand{\defequiv}{\vcentcolon\equiv}
\newcommand{\seq}[1]{\mathrel{\vdash\!\!\!_{#1}}}

\newcommand{\stacksproject}[1]{\cite[{\href{http://stacks.math.columbia.edu/tag/#1}{Tag~#1}}]{stacks-project}}

\definecolor{gray}{rgb}{0.7,0.7,0.7}

\begin{document}

\begin{defn}An~$R$-module~$E$ is \emph{synthetically quasicoherent} if and only if,
for any finitely presented~$R$-algebra~$A$, the canonical~$R$-algebra
homomorphism
\[ E \otimes_R A \longrightarrow [\Spec(A), E] = [[A, R]_R, E] \]
which maps a pure tensor~$x \otimes f$ to~$(\varphi \mapsto \varphi(f) x)$ is
bijective.\end{defn}

\begin{thm}Let~$E \in \Zar(S)$ be an~$\affl$-algebra.
If~$E$ is quasicoherent, that is of
the form~$\E_0^\Zar$ for some quasicoherent~$\O_S$-module~$\E_0$,
then~$E$ is synthetically quasicoherent from the internal point of view of~$\Zar(S)$.
The converse holds in any of the following situations:
\begin{enumerate}
\item The site used to construct~$\Zar(S)$ is one of the parsimonious sites.
\item The base scheme~$S$ is concentrated (quasicompact and quasiseparated) and
the functor~$E$ maps directed limits of inverse systems of~$S$-schemes with
affine transition morphisms to colimits in~$\Set$.
\end{enumerate}
\end{thm}

\begin{proof}To verify that~$E$ is synthetically quasicoherent, we have to
verify a condition for~$\affl$-algebras~$A$ in any
slice~$\Zar(S)/\ul{T}$. If such an algebra is finitely presented from
the internal point of view, then there is a covering~$T = \bigcup_i T_i$ such
that the each of the restrictions of the algebra to the~$T_i$ is of the
form~$(\A_0)^\Zar$ for some finitely presented~$\O_{T_i}$-algebra~$\A_0$.
Without loss of generality, we will just assume that~$A$ itself is of the
form~$(\A_0)^\Zar$ for a finitely presented~$\O_S$-algebra~$\A_0$.

By Proposition~\ref{XXX}, the internal expression~$\Spec(A)$ is the functor of
points of~$\RelSpek_S \A_0$. For any~$S$-scheme~$f : T \to S$ contained in the site
used to define~$\Zar(S)$, we consider the fiber product
\[ \xymatrix{
  \RelSpek_T(f^*\A_0) \ar[r]^{f'} \ar[d]_{p'} & \RelSpek_S\A_0 \ar[d]^p \\
  T \ar[r]_f & S.
} \]
Since~$\RelSpek_T(f^*\A_0) \to S$ is contained in the site (for any of our
admissible sites), we may conclude using the following chain of isomorphisms:
\begin{align*}
  [\Spec(A), E](T) &\cong
  \Hom_{\Zar(S)}(\ul{T}, [\Spec(A), E])
  \cong \Hom_{\Zar(S)}(\ul{T} \times \Spec(A), E) \\
  &\cong \Hom_{\Zar(S)}(\underline{T \times_S \RelSpek_S\A_0}, E)
  \cong E(\RelSpek_T(f^*\A_0)) \\
  &\cong \Gamma(\RelSpek_T(f^*\A_0), (p')^* f^* \E_0)
  \cong \Gamma(T, (p')_* (p')^* f_* \E_0) \\
  &\cong \Gamma(T, f^*\E_0 \otimes_{\O_T} f^*\A_0)
  \cong \Gamma(T, (\E_0 \otimes_{\O_S} \A_0)^\Zar) \\
  &\cong \Gamma(T, \E_0^\Zar \otimes_\affl \A_0^\Zar)
  \cong \Gamma(T, E \otimes_\affl A).
\end{align*}
The antepenultimate isomorphism is because pullback of modules in~$\Sh(S)$ to
modules in~$\Sh(T)$ commutes with tensor product. The penultimate isomorphism
is because pullback of a sheaf in~$\Sh(S)$ to a sheaf in~$\Zar(S)$ commutes
with tensor product (Lemma~\ref{lemma:zar-tensor-product-commutes}).

For the converse direction, we first verify that the restrictions~$E|_{\Sh(T)}$
to the little Zariski topos of each~$S$-scheme~$T$ contained in the site used
to define~$\Zar(S)$ are quasicoherent~$\O_T$-modules. For this, we employ the
quasicoherence criterion of Theorem~\ref{thm:qcoh-sheafchar}: For any open
affine subset~$T' \subseteq T$ and any function~$h \in \Gamma(T', \O_T)$ we
verify that the canonical morphism
\begin{equation}\label{eqn:want-iso}\tag{$\dagger$}
  E|_{\Sh(T)}[h^{-1}] \longrightarrow j_*(E|_{\Sh(D(h))})
\end{equation}
is an isomorphism, where~$j : D(h) \hookrightarrow T'$ denotes the inclusion.
This follows from the assumption of synthetic quasicoherence by considering
the~$\affl$-module~$A \defeq \affl[h^{-1}]$ (in the slice~$\Zar(S)/\ul{T'}$):
This expresses that the canonical morphism
\begin{equation}\label{eqn:have-iso}\tag{$\ddagger$}
  E \otimes_\affl \affl[h^{-1}] \longrightarrow [\Spec(A), E]
\end{equation}
is an isomorphism (of~$\affl$-modules in~$\Zar(S)/\ul{T'}$). Restricting the
domain to~$\Sh(T')$ yields the sheaf~$E|_{\Sh(T')} \otimes_{\O_{T'}}
\O_{T'}[h^{-1}]$, since restricting commutes with the geometric constructions
``forming the tensor product'' and ``localizing away from~$h$''.
Since~$\Spec(A)$ is the functor of points of~$D(h)$, restricting
the codomain to~$\Sh(T')$ yields the sheaf~$j_*(E|_{\Sh(D(h))})$.
The canonical morphism~\eqref{eqn:want-iso} which we want to recognize as an
isomorphism is therefore the restriction of the canonical
morphism~\eqref{eqn:have-iso} which we know to be an isomorphism.

A natural candidate for an quasicoherent~$\O_S$-module~$\E_0$ with~$E \cong
\E_0^\Zar$ is~$\E_0 \defeq E|_{\Sh(S)}$. We'll show that this is indeed true.
Let~$f : T \to S$ be any~$S$-scheme contained in the site used to
define~$\Zar(S)$. We assume for the time being that~$f$ is of finite
presentation and affine, so~$T \cong \RelSpek_S \A_0$ for some finitely
presented~$\O_S$-algebra~$\A_0$. We want to verify that the canonical morphism
\begin{equation}\label{eqn:want-iso2}\tag{§}
  f^*(E|_{\Sh(S)}) \longrightarrow E|_{\Sh(T)}
\end{equation}
is an isomorphism. Since the functor~$f_*$ from quasicoherent~$\O_T$-modules to
quasicoherent~$\O_S$-modules is fully faithful (the morphism~$f$ being affine)
and domain and codomain of that morphism are quasicoherent, it suffices to
verify that its image under~$f_*$ is an isomorphism. This image is the
canonical morphism
\[ E|_{\Sh(S)} \otimes_{\O_S} \A_0 \longrightarrow f_*(E|_{\Sh(T)}). \]
The assumption of synthetic quasicoherence, applied to the~$\affl$-algebra~$A
\defeq \A_0^\Zar$, shows that this morphism is an isomorphism.

In situation~(1), the only step left to do is to generalize the argument is the
previous paragraph to morphisms~$f : T \to S$ which are locally of finite
presentation. This works out because there are open covers of~$S$ and~$T$ such
that the appropriate restrictions of~$f$ are of finite presentation and affine.
The assumption of synthetic quasicoherence then needs to be applied to
to~$\affl$-algebras in suitable slices of~$\Zar(S)$, showing that the canonical
morphism~\eqref{eqn:want-iso2} is locally an isomorphism and therefore globally
as well.

In situation~(2), we employ the technique of approximating general~$S$-schemes
by~$S$-schemes of finite presentation. Specifically, let~$f : T \to S$ be an
arbitrary~$S$-scheme contained in the site used to define~$\Zar(S)$. Without
loss of generality, we may assume that~$T$ is an affine scheme. Thus~$T$ is
quasicompact and quasiseparated, and~$S$ is quasiseparated. We may therefore
apply the lemma of relative approximation~\stacksproject{09MV} to deduce
that~$T$ is a directed limit of an inverse system of~$S$-schemes~$f_i : T_i \to
S$ of finite presentation with affine transition maps. These~$S$-schemes
inherit quasicompactness and quasiseparatedness from~$S$. Therefore we can
apply a comparison result on the categories of quasicoherent
modules~\stacksproject{01Z0}:
\[
  E(T) = E(\lim_i T_i) \cong \colim_i E(T_i) \cong
  \colim_i \Gamma(T_i, f_i^* \E_0)
  \cong \Gamma(T, f^* \E_0). \qedhere
\]
\end{proof}

\end{document}