### 2018-04-05 slides-pssl103.tex
Slides for PSSL 103 (WIP)

\documentclass[12pt,utf8,notheorems,compress,t]{beamer}
\usepackage{etex}

\usepackage[english]{babel}

\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{multicol}
\usepackage{tabto}
\usepackage{xstring}
\usepackage{soul}\setul{0.3ex}{}
\usepackage[all]{xy}
\xyoption{rotate}
\usepackage{tikz}
\usetikzlibrary{calc,shapes.callouts,shapes.arrows,patterns}
\hypersetup{colorlinks=true}

\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\graphicspath{{images/}}

\usepackage[protrusion=true,expansion=true]{microtype}

\setlength\parskip{\medskipamount}
\setlength\parindent{0pt}

\title{Synthetic algebraic geometry and the phenomenon of nongeometric sequents}
\author{Ingo Blechschmidt}
\date{April 7th, 2018}

\useinnertheme[shadow=true]{rounded}
\useoutertheme{split}
\usecolortheme{orchid}
\usecolortheme{whale}
\setbeamerfont{block title}{size={}}

\useinnertheme{rectangles}

\usecolortheme{seahorse}
\definecolor{mypurple}{RGB}{150,0,255}
\setbeamercolor{structure}{fg=mypurple}
\definecolor{myred}{RGB}{150,0,0}
\setbeamercolor*{title}{bg=myred,fg=white}
\setbeamercolor*{titlelike}{bg=myred,fg=white}

\usefonttheme{serif}
\usepackage[T1]{fontenc}
\usepackage{libertine}

\newcommand{\A}{\mathcal{A}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\renewcommand{\G}{\mathcal{G}}
\newcommand{\GG}{\mathbb{G}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\ZZ}{\mathbb{Z}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\ppp}{\mathfrak{p}}
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\defeqv}{\vcentcolon\equiv}
\newcommand{\Sh}{\mathrm{Sh}}
\newcommand{\GL}{\mathrm{GL}}
\newcommand{\Zar}{\mathrm{Zar}}
\newcommand{\op}{\mathrm{op}}
\newcommand{\Set}{\mathrm{Set}}
\newcommand{\Eff}{\mathrm{Ef{}f}}
\newcommand{\Sch}{\mathrm{Sch}}
\newcommand{\Aff}{\mathrm{Aff}}
\newcommand{\LRS}{\mathrm{LRS}}
\newcommand{\Hom}{\mathrm{Hom}}
\DeclareMathOperator{\Spec}{Spec}
\newcommand{\lra}{\longrightarrow}
\newcommand{\RelSpec}{\operatorname{Spec}}
\renewcommand{\_}{\mathpunct{.}}
\newcommand{\?}{\,{:}\,}
\newcommand{\speak}[1]{\ulcorner\text{\textnormal{#1}}\urcorner}
\newcommand{\ull}[1]{\underline{#1}}
\newcommand{\affl}{\ensuremath{{\ull{\AA}^1}}}
\newcommand{\Ll}{\vcentcolon\!\Longleftrightarrow}

\setbeamertemplate{blocks}[rounded][shadow=false]

\newcommand{\fmini}[2]{\framebox{\begin{minipage}{#1}#2\end{minipage}}}
\makeatletter
\def\underunbrace#1{\mathop{\vtop{\m@th\ialign{##\crcr
      $\hfil\displaystyle{#1}\hfil$\crcr\noalign{\kern3\p@\nointerlineskip}
      \crcr\noalign{\kern3\p@}}}}\limits}
\def\overunbrace#1{\mathop{\vbox{\m@th\ialign{##\crcr\noalign{\kern3\p@}
      \crcr\noalign{\kern3\p@\nointerlineskip}
      $\hfil\displaystyle{#1}\hfil$\crcr}}}\limits}
\makeatother

\newenvironment{changemargin}[2]{%
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \setlength{\leftmargin}{#1}%
    \setlength{\rightmargin}{#2}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\itemindent}{\parindent}%
    \setlength{\parsep}{\parskip}%
  }%
  \item[]}{\end{list}}

\newcommand{\pointthis}[2]{%
  \tikz[remember picture,baseline]{\node[anchor=base,inner sep=0,outer sep=0]%
    (#1) {#1};\node[overlay,rectangle callout,%
    callout relative pointer={(-0.5cm,1.0cm)},fill=blue!20] at ($(#1.north)+(0.9cm,-1.6cm)$) {#2};}%
}%

\newcommand{\hcancel}[5]{%
  \tikz[baseline=(tocancel.base)]{
    \node[inner sep=0pt,outer sep=0pt] (tocancel) {#1};
    \draw[red, line width=0.4mm] ($(tocancel.south west)+(#2,#3)$) -- ($(tocancel.north east)+(#4,#5)$);
  }%
}

\newcommand{\slogan}[1]{%
  \begin{center}%
    \setlength{\fboxrule}{2pt}%
    \setlength{\fboxsep}{8pt}%
    {\usebeamercolor[fg]{item}\fbox{\usebeamercolor[fg]{normal text}\parbox{0.93\textwidth}{#1}}}%
  \end{center}%
}

\newcommand{\sloganwithoutborder}[1]{%
  \begin{center}%
    \setlength{\fboxrule}{0pt}%
    \setlength{\fboxsep}{-14pt}%
    {\usebeamercolor[fg]{item}\fbox{\usebeamercolor[fg]{normal
    text}\parbox{0.9\textwidth}{\begin{center}#1\end{center}}}}%
  \end{center}%
}

% Adapted from https://latex.org/forum/viewtopic.php?t=2251 (Stefan Kottwitz)
\newenvironment<>{varblock}[2]{
  \begin{center}
    \begin{minipage}{#1}
      \setlength{\textwidth}{#1}
      \begin{actionenv}#3
	\def\insertblocktitle{\centering #2}
	\par
	\usebeamertemplate{block begin}}{
        \par
        \usebeamertemplate{block end}
      \end{actionenv}
    \end{minipage}
  \end{center}}

\setbeamertemplate{frametitle}{%
  \vskip0.7em%
  \leavevmode%
  \begin{beamercolorbox}[dp=1ex,center]{}%
      \usebeamercolor[fg]{item}{\textbf{{\Large \insertframetitle}}}
  \end{beamercolorbox}%
}

\setbeamertemplate{navigation symbols}{}

\newcounter{framenumberpreappendix}
\newcommand{\backupstart}{
  \setcounter{framenumberpreappendix}{\value{framenumber}}
}
\newcommand{\backupend}{
  \addtocounter{framenumberpreappendix}{-\value{framenumber}}
  \addtocounter{framenumber}{\value{framenumberpreappendix}} 
}

\setbeamertemplate{footline}{%
  \begin{beamercolorbox}[wd=\paperwidth,ht=2.25ex,dp=1ex,right,rightskip=1mm,leftskip=1mm]{}%
    \inserttitle \hfill
    \insertframenumber\,/\,\inserttotalframenumber
  \end{beamercolorbox}%
  \vskip0pt%
}

\newcommand{\hil}[1]{{\usebeamercolor[fg]{item}{\textbf{#1}}}}

\IfSubStr{\jobname}{\detokenize{nonotes}}{
  \setbeameroption{hide notes}
}{
  \setbeameroption{show notes}
}
\setbeamertemplate{note page}[plain]
\setbeameroption{hide notes}

\begin{document}

\addtocounter{framenumber}{-1}

\begin{frame}[c]
  \centering
  \includegraphics[scale=0.3]{images/external-internal-small}
  \medskip

  \hil{Synthetic algebraic geometry} \\
  \small\emph{a case study in applied topos theory}

  \large\&

  \normalsize

  \hil{the phenomenon of nongeometric sequents}

  \medskip

  \scriptsize
  Ingo Blechschmidt \\
  University of Augsburg
  \medskip

  103rd Peripatetic Seminar on Sheaves and Logic (Brno) \\
  April 7th, 2018
  \par
\end{frame}

\section{Preliminaries}

\subsection{Synthetic differential geometry}

\begin{frame}{Synthetic differential geometry}
  \vspace*{-1em}
  \begin{varblock}{9.5cm}{The axiom of microaffinity}
    \justifying
    Let~$\Delta = \{ \varepsilon \in \RR \,|\, \varepsilon^2 = 0 \}$.
    For any function $f : \Delta \to \RR$,
    there are unique numbers~$a, b \in \RR$ such that
    $f(\varepsilon) = a + b \varepsilon$
    for all~$\varepsilon \in \Delta$.
  \end{varblock}

  \begin{center}
    \includegraphics[width=0.5\textwidth]{microaffinity}
  \end{center}

  \vspace*{-1em}
  \begin{itemize}
    \item The \hil{derivative} of~$f$ as above at zero is~$b$.
    \item Manifolds are \hil{just sets}.
    \item A \hil{tangent vector} to~$M$ is a map~$\Delta \to M$.
  \end{itemize}

  \centering
  \hil{Toposes provide models for this theory.}
\end{frame}

\subsection{Toposes}

\begin{frame}{The internal universe of a topos}
  A \hil{topos} is a category which has finite limits,
  is cartesian closed and has a subobject classifier, for instance
  \begin{itemize}
    \item \hil{$\boldsymbol{\Set}$}, the category of sets;
    \item \hil{$\boldsymbol{\Sh(X)}$}, the category of set-valued sheaves over
    a space~$X$;
    \item \hil{$\boldsymbol{\Eff}$}, the ef{}fective topos (roughly: a
    category of data types).
  \end{itemize}

  \slogan{Any topos supports an \hil{internal language},
  which is sound with respect to
  \only<1>{\hil{intuitionistic reasoning}}%
  \only<2->{\hil{\pointthis{intuitionistic reasoning}{
    \textnormal{\textcolor{black}{
    no $\varphi \vee \neg\varphi$,\ \
    no $\neg\neg\varphi \Rightarrow \varphi$,\ \
    no axiom of choice}}}}}.}
\end{frame}

\newcommand{\intex}[3]{#1 \qquad #2\par #3\medskip}

\begin{frame}{Curious universes}
  \small
  \intex{
    $\Eff \models \text{``There are infinitely many prime numbers.''}$
  }{\textcolor{green!90}{\cmark}}{
    External meaning: There is a (Turing) \hil{machine} producing arbitrarily many prime
    numbers.
  }

  \intex{
    $\Eff \not\models \text{``Any Turing machine halts or doesn't halt.''}$
  }{\textcolor{red!80}{\xmark}}{
    External meaning: There is a \hil{halting oracle} which determines whether
    any given machine halts or doesn't halt.
  }

  \intex{
    $\Sh(X) \not\models \text{``Any cont. function with opposite signs has
    a zero.''}$
  }{\textcolor{red!80}{\xmark}}{
    External meaning: Zeros can be picked \hil{locally continuously} in
    continuous families of continuous functions.
  }
\end{frame}

\end{document}

"Any map from the reals to the reals is smooth." This statement holds in the
context of synthetic differential geometry, a well-developed account of smooth
manifolds and related notions which allows us to bring formal reasoning much
closer to geometric intuition.

We employ the internal language of the big Zariski topos of a base scheme to
give a similar account of algebraic geometry, reminiscent of the language of
the classical Italian school and incorporating Grothendieck's functor-of-points
philosophy. From the point of view of the Zariski topos, a scheme over the base
will look like a plain old set, and the affine line will look like a certain
field. Central to the synthetic account is the notion of "synthetic
quasicoherence", which doesn't have an analog in synthetic differential
geometry and which gives the account a distinct algebraic flavor.

The higher-order axiom of synthetic quasicoherence implies all known internal
properties of the affine line, for instance that it is a field and that it is
algebraically closed in a weak sense. We surmise that this is for a deeper
reason, related to the age-old question "which nongeometric sequents hold in
the classifying topos of a geometric theory?". The second part of the talk
reports on work in progress about this topic.