### 2018-04-15 paper-generic-freeness.tex
On weakly Noetherian modules

\documentclass{amsart}

\usepackage[utf8]{inputenc}
\usepackage{amsthm,mathtools}
\usepackage[protrusion=true,expansion=true]{microtype}
\usepackage{hyperref}

\title{An elementary and constructive proof of Grothendieck's generic freeness
lemma}
\author{Ingo Blechschmidt}

\theoremstyle{definition}
\newtheorem{defn}{Definition}[]
\newtheorem{ex}[defn]{Example}

\theoremstyle{plain}
\newtheorem{prop}[defn]{Proposition}
\newtheorem{cor}[defn]{Corollary}
\newtheorem{lemma}[defn]{Lemma}
\newtheorem{thm}[defn]{Theorem}
\newtheorem{scholium}[defn]{Scholium}

\theoremstyle{remark}
\newtheorem{rem}[defn]{Remark}
\newtheorem{question}[defn]{Question}
\newtheorem{speculation}[defn]{Speculation}
\newtheorem{caveat}[defn]{Caveat}
\newtheorem{conjecture}[defn]{Conjecture}

\newcommand{\XXX}[1]{\textbf{XXX: #1}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\bbb}{\mathfrak{b}}
\newcommand{\defeq}{\vcentcolon=}
\DeclareMathOperator{\Spec}{Spec}

\begin{document}

\begin{abstract}
  We present a new and direct proof of Grothendieck's generic freeness
  lemma in its general form. Unlike the previously published proofs, it doesn't
  proceed in a series of reduction steps and is fully constructive, not
  involving the axiom of choice or even the law of excluded middle. It was
  found by unwinding the result of a general topos-theoretic technique.
\end{abstract}

\maketitle

We prove Grothendieck's generic freeness lemma for rings and modules in the
following form.

\begin{thm}\label{thm:algebraic}Let~$A$ be a reduced ring. Let~$B$ be
an~$A$-algebra of finite type. Let~$M$ be a finitely generated~$B$-module.
If~$f = 0$ is the only element of~$A$ such that
\begin{enumerate}
\item the~$A[f^{-1}]$-modules $B[f^{-1}]$ and $M[f^{-1}]$ are free,
\item the~$A[f^{-1}]$-algebra~$B[f^{-1}]$ is of finite presentation, and
\item the~$B[f^{-1}]$-module~$M[f^{-1}]$ is finitely presented,
\end{enumerate}
then~$1 = 0$ in~$A$.
\end{thm}

Previously known proofs proceed in a series of intermediate steps, reducing
to the case that~$A$ is a Noetherian integral domain where one can argue by
\emph{dévissage}; but in fact, a direct proof is possible and shorter.

Grothendieck's generic freeness lemma is often presented in contrapositive form
or in the following geometric variant:

\begin{thm}\label{thm:geometric}Let~$A$ be a reduced ring. Let~$B$ be
an~$A$-algebra of finite type. Let~$M$ be a finitely generated~$B$-module. Then
the space~$\Spec(A)$ contains a dense open~$U$ such that over~$U$,
\begin{enumerate}
\item[(a)] $B^\sim$ and~$M^\sim$ are free as sheaves of~$A^\sim$-modules,
\item[(b)] $B^\sim$ is of finite presentation as a sheaf of~$A^\sim$-algebras, and
\item[(c)] $M^\sim$ is finitely presented as a sheaf of~$B^\sim$-modules.
\end{enumerate}
\end{thm}

Theorem~\ref{thm:geometric} immediately follows from
Theorem~\ref{thm:algebraic} by defining~$U$ as the union of all the basic
opens~$D(f)$ such that~(1),~(2), and~(3) hold. It's clear that~(a),~(b),
and~(c) hold over~$U$, and~$U$ is dense for if~$V$ is an arbitrary open
such that~$U \cap V = \emptyset$, the open~$V$ is itself empty: Let~$h \in A$
such that~$D(h) \subseteq V$. The hypothesis implies the assumptions of
Theorem~\ref{thm:algebraic} for the datum~$(A[h^{-1}], B[h^{-1}], M[h^{-1}])$.
Thus~$1 = 0 \in A[h^{-1}]$, so~$h$ is nilpotent and~$D(h) = \emptyset$.

The new proof was found using a general topos-theoretical technique which we
believe to be useful in other situations as well. This technique allows to view
reduced rings and their modules from a different point of view, one from which
reduced rings look like fields. Since Grothendieck's generic freeness
is trivial for fields, this technique yields a trivial proof for reduced rings.
The proof presented here was obtained by unwinding the topos-theoretic proof,
yielding a self-contained argument without any references to topos theory.
We refer readers who want to learn about this technique to an accompanying
note~\cite{blechschmidt:wlog}.

\XXX{Acknowledgments, especially Martin Brandenburg}

\XXX{References to prior proofs: EGA, Stacks Project, Charles Staats. Check
\href{https://math.stackexchange.com/questions/1048496/question-on-generic-freeness-ref-matsumura-page-185}{this
MO question}}

\section{The proof of the general case}

\begin{lemma}\label{lemma:basis}
Let~$A$ be a ring. Let~$M$ be an~$A$-module with generating
family~$(x_i)_{i \in I}$ where~$I$ is a totally ordered set. Assume that the
only element~$g \in A$ such that one of the~$x_i$ is an~$A[g^{-1}]$-linear
combination in~$M[g^{-1}]$ of other generators with smaller index is~$g = 0$.
Then~$M$ is free with~$(x_i)_{i \in I}$ as a basis.
\end{lemma}

\begin{proof}Let~$\sum_i a_i x_i = 0$. Starting with the greatest
index~$i$ which appears in that sum, we see that in~$M[a_i^{-1}]$, the
element~$x_i$ is an~$A[g^{-1}]$-linear combination of other generators with
smaller index. Thus~$a_i = 0$ by assumption.\end{proof}

\begin{thm}Let~$A$ be a reduced ring. Let~$B$ be an~$A$-algebra of finite type.
Let~$M$ be a finitely generated~$B$-module. If~$f = 0$ is the only element
of~$A$ such that~$M[f^{-1}]$ is free as an~$A[f^{-1}]$-module, then~$1 = 0$
in~$A$.\end{thm}

\begin{proof}Let~$B$ be generated by~$(x_1,\ldots,x_n)$ as an~$A$-algebra and
let~$M$ be generated by~$(v_1,\ldots,v_m)$ as an~$B$-module. We endow the set
\[ I \defeq \{ (\ell, j_1,\ldots,j_n) \,|\,
  \ell \in \{ 1,\ldots,m \},
  j_1,\ldots,j_n \in \{ 0,1,\ldots \} \} \]
with the lexicographic order. The family~$(w_i)_{i \in I} \defeq (x_1^{j_1}
\ldots x_n^{j_n} v_\ell)_{(\ell,j_1,\ldots,j_n) \in I}$ thus generates~$M$ as
an~$A$-module, and we'll call a subfamily~$(w_i)_{i \in I' \subseteq I}$
\emph{good} if and only if for all~$i \in I$, the vector~$w_i$ is a linear
combination of the vectors~$(w_{i'})_{i' \in I, i' \preceq i}$,
and if~$(\ell,j_1,\ldots,j_n) \not\in I'$ implies~$(\ell,k_1,\ldots,k_n) \not\in
I'$ for all~$k_1 \geq j_1, \ldots, k_n \geq j_n$. \XXX{Looks like in Figure XXX}

We induct on the shape of a good generating system, starting with~$(w_i)_{i \in I}$.

We show that~$(w_i)_{i \in I}$ is a basis of~$M$ by verifying the assumptions
of Lemma~\ref{lemma:basis}. Thus let~$g \in A$ be given such that one of
the~$w_i$ is an $A[g^{-1}]$-linear combination of generators with smaller index
in~$M[g^{-1}]$. Removing~$w_i = x_1^{j_1} \ldots x_n^{j_n} v_\ell$ and also all
vectors~$x_1^{k_1} \ldots x_n^{k_n} v_\ell$ where~$k_1 \geq j_1, \ldots, k_n
\geq j_n$, we obtain a subfamily which is still good. By induction, applied
to~$A[g^{-1}]$ and its module~$M[g^{-1}]$, it therefore follows that $A[g^{-1}] = 0$.
This implies that $g = 0$ since~$A$ is reduced.

Thus~$M$ is free. We finish by using the assumption for~$f = 1$.
\end{proof}

Recall that a module is Noetherian if and only if all of its submodules are
finitely generated. We call a module~$M$ over an~$A$-algebra~$B$ \emph{weakly
Noetherian} (over~$A$) if and only if, for all~$B$-submodules~$U \subseteq M$
and all~$f \in A$ the condition
\begin{multline*}
  \text{for all~$g \in A$, if $U[(fg)^{-1}]$ is finitely generated as
an~$B[(fg)^{-1}]$-module,} \\
  \text{then~$fg$ is nilpotent in~$A$} \end{multline*}
implies that~$f$ is nilpotent. Geometrically, this expresses that all
quasicoherent sheaves of submodules of~$M^\sim$ are, over a dense open
of~$\Spec(A)$, of finite type over~$B^\sim$. An~$A$-algebra~$B$
is \emph{weakly Noetherian} if and only if it is weakly Noetherian as a module
over itself.

This notion is interesting for the following reason.

\begin{lemma}\label{lemma:self-noetherian}
Let~$A$ be a reduced ring. Then~$A$, regarded as an algebra over itself, is
weakly Noetherian.\end{lemma}

\begin{proof}Let~$\aaa \subseteq A$ be an ideal. Let~$f \in A$ such that, for
any~$g \in A$, if~$\aaa[(fg)^{-1}]$ is finitely generated, then~$fg = 0$.
We verify that~$f = 0$.

We first show that~$\aaa[f^{-1}] = (0)$. Let~$g \in \aaa$. Then~$\aaa[(fg)^{-1}] =
(1)$ is finitely generated. Thus~$fg = 0$, hence~$g = 0$ in~$A[f^{-1}]$.

Since~$\aaa[f^{-1}]$ is finitely generated, we finish by using the
assumption for~$g = 1$.
\end{proof}

The weak Noetherian condition enjoys the same basic stability properties as the
usual Noetherian condition:

\begin{lemma}\label{lemma:noetherian-stability}
Let~$A$ be a ring. Let~$B$ be an~$A$-algebra.
\begin{enumerate}
\item If~$B$ is weakly Noetherian, then so is~$B[X]$.
\item Let~$0 \to M' \to M \to M'' \to 0$ be a short exact sequence
of~$B$-modules. Then~$M$ is weakly Noetherian if and only if~$M'$ and~$M''$
are.
\end{enumerate}
\end{lemma}

We can use these observations to finish the proof of Grothendieck's generic
freeness lemma.

\begin{prop}Let~$A$ be a reduced ring. Let~$B$ be an~$A$-algebra of finite type.
If~$f = 0$ is the only element of~$A$ such that~$B[f^{-1}]$ is of finite
presentation as an~$A[f^{-1}]$-algebra, then~$1 = 0$ in~$A$.
\end{prop}

\begin{proof}We write~$B = A[X_1,\ldots,X_n]/\aaa$. By Lemma~\ref{lemma:self-noetherian} and
Lemma~\ref{lemma:noetherian-stability}, the~$A$-algebra~$A[X_1,\ldots,X_n]$ is
weakly Noetherian. We can therefore conclude if we can show, for any~$g \in A$,
that if the ideal~$\aaa[g^{-1}]$ is finitely generated, then~$g = 0$. This
follows immediately from the assumption (for~$f = g$).
\end{proof}

\begin{prop}Let~$A$ be a reduced ring. Let~$B$ be an~$A$-algebra of finite type.
Let~$M$ be a finitely generated~$B$-module. If~$f = 0$ is the only element
of~$A$ such that~$M[f^{-1}]$ is finitely presented as an~$B[f^{-1}]$-module,
then~$1 = 0$ in~$A$.
\end{prop}

\begin{proof}We write~$M = B^n/U$ for an~$B$-submodule~$U$. By Lemma~\ref{lemma:self-noetherian} and
Lemma~\ref{lemma:noetherian-stability}, the~$B$-module~$B^n$ is weakly
Noetherian. We can therefore conclude if we can show, for any~$g \in A$, that if
the submodule~$U[g^{-1}]$ is finitely generated, then~$g = 0$. This follows
immediately from the assumption (for~$f = g$).
\end{proof}

\section{The proof of the finitely-generated case}

The following proposition is just an instance of Grothendieck's generic
freeness lemma. We prove it here because it admits an easier proof.

\begin{prop}Let~$A$ be a reduced ring. Let~$M$ be a finitely
generated~$A$-module. If~$f = 0$ is the only element of~$A$ such
that~$M[f^{-1}]$ is a finite free~$A[f^{-1}]$-module, then~$1 = 0$ in~$A$.
\end{prop}

\begin{proof}We proceed by induction on the length of a given generating family
of~$M$. Let~$M$ be generated by~$(v_1,\ldots,v_m)$.

We show that the family~$(v_1,\ldots,v_m)$ is linear independent. Let~$\sum_i
a_i v_i = 0$. Over~$A[a_i^{-1}]$, the vector~$v_i \in M[a_i^{-1}]$ is a linear
combination of the other generators. Thus~$M[a_i^{-1}]$ can be generated as
an~$A[a_i^{-1}]$-module by fewer than~$m$ generators. The induction hypothesis,
applied to this module, yields that~$1 = 0$ in~$A[a_i^{-1}]$. Since~$A$ is
reduced, this amounts to~$a_i = 0$.

We finish by using the assumption for~$f = 1$.
\end{proof}

\end{document}