### 2018-05-15 paper-generic-freeness.tex
Improved the presentation; fought the index war.

\documentclass{amsart}

\usepackage[utf8]{inputenc}
\usepackage{amsthm,mathtools,tikz,float,caption}
\usetikzlibrary{patterns,matrix}
\usepackage[protrusion=true,expansion=true]{microtype}
\usepackage{hyperref}

\usepackage[natbib=true,style=numeric,maxnames=10]{biblatex}
\usepackage[babel]{csquotes}
\bibliography{paper-generic-freeness.bib}

\title[Grothendieck's generic freeness lemma]{An elementary and constructive
proof of Grothendieck's generic freeness lemma}
\author{Ingo Blechschmidt}

\theoremstyle{definition}
\newtheorem{defn}{Definition}[]
\newtheorem{ex}[defn]{Example}

\theoremstyle{plain}
\newtheorem{prop}[defn]{Proposition}
\newtheorem{cor}[defn]{Corollary}
\newtheorem{lemma}[defn]{Lemma}
\newtheorem{thm}[defn]{Theorem}
\newtheorem{scholium}[defn]{Scholium}

\theoremstyle{remark}
\newtheorem{rem}[defn]{Remark}
\newtheorem{question}[defn]{Question}
\newtheorem{speculation}[defn]{Speculation}
\newtheorem{caveat}[defn]{Caveat}
\newtheorem{conjecture}[defn]{Conjecture}

\newcommand{\XXX}[1]{\textbf{XXX: #1}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\bbb}{\mathfrak{b}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\defeq}{\vcentcolon=}
\DeclareMathOperator{\Spec}{Spec}

\newcommand{\stacksproject}[1]{\cite[{\href{https://stacks.math.columbia.edu/tag/#1}{Tag~#1}}]{stacks-project}}

\begin{document}

\begin{abstract}
  We present a new and direct proof of Grothendieck's generic freeness
  lemma in its general form. Unlike the previously published proofs, it doesn't
  proceed in a series of reduction steps and is fully constructive, not
  involving the axiom of choice or even the law of excluded middle. It was
  found by unwinding the result of a general topos-theoretic technique.
\end{abstract}

\maketitle
\thispagestyle{empty}

\noindent
We prove Grothendieck's generic freeness lemma for rings and modules in the
following form.

\begin{thm}\label{thm:algebraic}Let~$A$ be a reduced ring. Let~$B$ be
an~$A$-algebra of finite type. Let~$M$ be a finitely generated~$B$-module.
If~$f = 0$ is the only element of~$A$ such that
\begin{enumerate}
\item the~$A[f^{-1}]$-modules $B[f^{-1}]$ and $M[f^{-1}]$ are free,
\item the~$A[f^{-1}]$-algebra~$B[f^{-1}]$ is of finite presentation and
\item the~$B[f^{-1}]$-module~$M[f^{-1}]$ is finitely presented,
\end{enumerate}
then~$1 = 0$ in~$A$.
\end{thm}

Previously known proofs either only cover the case where~$A$ is a Noetherian
integral domain, where one can argue by \emph{dévissage} (see for
instance~\cite[Lemme~6.9.2]{ega-4-2},
\cite[Thm.~24.1]{matsumura:commutative-ring-theory} or
\cite[Thm.~14.4]{eisenbud:commutative-algebra}), or proceed in a series of
intermediate steps, reducing to that case (see for
instance~\cite{staats:generic-freeness} or~\stacksproject{051Q}); but in fact,
a direct proof is possible and shorter.

Grothendieck's generic freeness lemma is often presented in contrapositive form
or in the following geometric variant:

\begin{thm}\label{thm:geometric}Let~$A$ be a reduced ring. Let~$B$ be
an~$A$-algebra of finite type. Let~$M$ be a finitely generated~$B$-module. Then
the space~$\Spec(A)$ contains a dense open~$U$ such that over~$U$,
\begin{enumerate}
\item[(a)] $B^\sim$ and~$M^\sim$ are free as sheaves of~$A^\sim$-modules,
\item[(b)] $B^\sim$ is of finite presentation as a sheaf of~$A^\sim$-algebras and
\item[(c)] $M^\sim$ is finitely presented as a sheaf of~$B^\sim$-modules.
\end{enumerate}
\end{thm}

Theorem~\ref{thm:geometric} immediately follows from
Theorem~\ref{thm:algebraic} by defining~$U$ as the union of all those basic
opens~$D(f)$ such that~(1),~(2) and~(3) hold. It's clear that~(a),~(b)
and~(c) hold over~$U$, and~$U$ is dense for if~$V$ is an arbitrary open
such that~$U \cap V = \emptyset$, the open~$V$ is itself empty: Let~$h \in A$ be
such that~$D(h) \subseteq V$. The hypothesis implies the assumptions of
Theorem~\ref{thm:algebraic} for the datum~$(A[h^{-1}], B[h^{-1}], M[h^{-1}])$.
Thus~$1 = 0 \in A[h^{-1}]$, so~$h$ is nilpotent and~$D(h) = \emptyset$.

The new proof was found using a general topos-theoretical technique which we
believe to be useful in other situations as well. This technique allows to view
reduced rings and their modules from a different point of view, one from which
reduced rings look like fields. Since Grothendieck's generic freeness
is trivial for fields, this technique yields a trivial proof for reduced rings.
The proof presented here was obtained by unwinding the topos-theoretic proof,
yielding a self-contained argument without any references to topos theory.
We refer readers who want to learn about this technique to a forthcoming companion 
paper~\cite{blechschmidt:wlog}.

\textbf{Acknowledgments.} The proof presented here was prompted by
user~HeinrichD on MathOverflow~\cite{mo:kernel} and greatly
benefited from discussions with Martin Brandenburg, who employed the constructive
version in a paper of his~\cite{brandenburg:schur}. I'm grateful to Marc
Nieper-Wißkirchen for carefully guiding my PhD studies at the University of
Augsburg, where most of the work for this paper was carried out.

\section{The proof of the finitely-generated case}

The following proposition is just a special instance of Grothendieck's generic
freeness lemma. Its proof is easier and shorter than the proof of the general
case, which is why we present it here. The general proof
will not refer to this one.

\begin{prop}Let~$A$ be a reduced ring. Let~$M$ be a finitely
generated~$A$-module. If~$f = 0$ is the only element of~$A$ such
that~$M[f^{-1}]$ is a finite free~$A[f^{-1}]$-module, then~$1 = 0$ in~$A$.
\end{prop}

\begin{proof}We proceed by induction on the length of a given generating family
of~$M$. Let~$M$ be generated by~$(v_1,\ldots,v_m)$.

We show that the family~$(v_1,\ldots,v_m)$ is linearly independent. Let~$\sum_i
a_i v_i = 0$. Over~$A[a_i^{-1}]$, the vector~$v_i \in M[a_i^{-1}]$ is a linear
combination of the other generators. Thus~$M[a_i^{-1}]$ can be generated as
an~$A[a_i^{-1}]$-module by fewer than~$m$ generators. The induction hypothesis,
applied to this module, yields that~$1 = 0$ in~$A[a_i^{-1}]$. Since~$A$ is
reduced, this amounts to~$a_i = 0$.

We finish by using the assumption for~$f = 1$.
\end{proof}

We remark that proof takes a somewhat curious course: Our goal is to
verify~$1 = 0$, but as an intermediate step we verify that~$M$ is free, which
after the fact will be a trivial statement.

\section{The proof of the general case}

\begin{proof}[Proof of Theorem~\ref{thm:algebraic}]
Let~$B$ be generated by~$(x_1,\ldots,x_n)$ as an~$A$-algebra and
let~$M$ be generated by~$(v_1,\ldots,v_m)$ as a~$B$-module. We endow the sets
\begin{align*}
  \I &\defeq \{ (i_1,\ldots,i_n) \,|\, i_1,\ldots,i_n \geq 0 \}
  \quad\text{and} \\
  \J &\defeq \{ (\ell, i_1,\ldots,i_n) \,|\, 1 \leq \ell \leq m, i_1,\ldots,i_n \geq 0 \}
\end{align*}
with the lexicographic order. The family~$(w_J)_{J \in \J} \defeq (x_1^{i_1}
\cdots x_n^{i_n} v_\ell)_{(\ell,i_1,\ldots,i_n) \in \J}$ generates~$M$ as
an~$A$-module, and we'll call a subfamily~$(w_J)_{J \in \J' \subseteq \J}$
\emph{good} if and only if for all~$J \in \J$, the vector~$w_J$ is a linear
combination of the vectors~$(w_{J'})_{J' \in \J', J' \preceq J}$,
and if~$(\ell,i_1,\ldots,i_n) \not\in \J'$ implies~$(\ell,k_1,\ldots,k_n) \not\in
\J'$ for all~$k_1 \geq i_1, \ldots, k_n \geq i_n$. Figure~\ref{fig:good} shows
how a good generating family can look like.
Similarly, we define when a subfamily of the canonical generating
family~$(x_1^{i_1} \cdots x_n^{i_n})_{(i_1,\ldots,i_n) \in \I}$ of~$B$ is good
(which is just the special case~$m = 1$).

\begin{figure}[t]
  \captionsetup{width=.9\linewidth}
  \begin{tikzpicture}[scale=0.9]
    \draw[step=1cm,gray,very thin] (0,0) grid (8,8);

    \draw
      (0,8) -- (0,0) -- (8,0);

    \fill[pattern=north east lines,pattern color=gray,opacity=0.5]
      (2,8) -- (2,5) -- (5,5) -- (5,3) -- (6,3) -- (6,2) -- (8,2) -- (8,8);
    \draw
      (2,8) -- (2,5) -- (5,5) -- (5,3) -- (6,3) -- (6,2) -- (8,2);

    \fill[fill=black] (2,5.25) -- (2,5) -- (2.25,5) -- (2.25,5.25);
    \fill[fill=black] (5,3.25) -- (5,3) -- (5.25,3) -- (5.25,3.25);
    \fill[fill=black] (6,2.25) -- (6,2) -- (6.25,2) -- (6.25,2.25);

    \fill[fill=blue!40!white,opacity=0.5]
      (0,8) -- (0,0) -- (4,0) -- (4,3) -- (3,3) -- (3,5) -- (2,5) -- (2,8);

    \fill[fill=red!40!white,opacity=0.5]
      (3,3) -- (4,3) -- (4,4) -- (3,4);

    % !perl -we 'for my $y (reverse 0..7) { for my $x (0..7) { print % "\$x^$x\Ey^$y\$ & " } print "\\\\ \n" }'
    \begin{scope}[yshift=4cm, xshift=4cm]
      \matrix[matrix of nodes,nodes={inner sep=0pt,text width=0.9cm,align=center,minimum height=0.9cm}]{
        \tiny $x^0y^7v_1$ & \tiny $x^1y^7v_1$ & \tiny $x^2y^7v_1$ & \tiny $x^3y^7v_1$ & \tiny $x^4y^7v_1$ & \tiny $x^5y^7v_1$ & \tiny $x^6y^7v_1$ & \tiny $x^7y^7v_1$ & \\
        \tiny $x^0y^6v_1$ & \tiny $x^1y^6v_1$ & \tiny $x^2y^6v_1$ & \tiny $x^3y^6v_1$ & \tiny $x^4y^6v_1$ & \tiny $x^5y^6v_1$ & \tiny $x^6y^6v_1$ & \tiny $x^7y^6v_1$ & \\
        \tiny $x^0y^5v_1$ & \tiny $x^1y^5v_1$ & \tiny $x^2y^5v_1$ & \tiny $x^3y^5v_1$ & \tiny $x^4y^5v_1$ & \tiny $x^5y^5v_1$ & \tiny $x^6y^5v_1$ & \tiny $x^7y^5v_1$ & \\
        \tiny $x^0y^4v_1$ & \tiny $x^1y^4v_1$ & \tiny $x^2y^4v_1$ & \tiny $x^3y^4v_1$ & \tiny $x^4y^4v_1$ & \tiny $x^5y^4v_1$ & \tiny $x^6y^4v_1$ & \tiny $x^7y^4v_1$ & \\
        \tiny $x^0y^3v_1$ & \tiny $x^1y^3v_1$ & \tiny $x^2y^3v_1$ & \tiny $x^3y^3v_1$ & \tiny $x^4y^3v_1$ & \tiny $x^5y^3v_1$ & \tiny $x^6y^3v_1$ & \tiny $x^7y^3v_1$ & \\
        \tiny $x^0y^2v_1$ & \tiny $x^1y^2v_1$ & \tiny $x^2y^2v_1$ & \tiny $x^3y^2v_1$ & \tiny $x^4y^2v_1$ & \tiny $x^5y^2v_1$ & \tiny $x^6y^2v_1$ & \tiny $x^7y^2v_1$ & \\
        \tiny $x^0y^1v_1$ & \tiny $x^1y^1v_1$ & \tiny $x^2y^1v_1$ & \tiny $x^3y^1v_1$ & \tiny $x^4y^1v_1$ & \tiny $x^5y^1v_1$ & \tiny $x^6y^1v_1$ & \tiny $x^7y^1v_1$ & \\
        \tiny $x^0y^0v_1$ & \tiny $x^1y^0v_1$ & \tiny $x^2y^0v_1$ & \tiny $x^3y^0v_1$ & \tiny $x^4y^0v_1$ & \tiny $x^5y^0v_1$ & \tiny $x^6y^0v_1$ & \tiny $x^7y^0v_1$ & \\
      };
    \end{scope}
  \end{tikzpicture}
  \caption{\label{fig:good}A graphical depiction of a good generating family (the non-hatched cells) in
  the special case~$n = 2, m = 1$, writing~``$x$'' and~``$y$'' for~$x_1$ and~$x_2$. The hatched cells indicate vectors which
  have already been removed from the family. The small black squares indicate
  \emph{corners}. If the vector in the red cell will be found to be expressible
  as a linear combination of vectors with smaller index (blue cells), it will
  be removed, along with the vectors in all cells to the top and to the right
  of the red cell.}
\end{figure}

We then proceed by induction on the
shapes of a given good generating family~$(w_J)_{J \in \J'}$ for~$M$ and a given
good generating family~$(s_I)_{I \in \I'}$ for~$B$, starting with the canonical ones. It's
reasonably obvious that this induction is well-founded; the formal statement
that it is so is known as \emph{Dickson's Lemma} in mathematical logic.

We show that~$(w_J)_{J \in \J'}$ is a basis of~$M$ by verifying linear
independence. Thus let~$\sum_J a_J w_J = 0$ in~$M$. We show that all
coefficients in this sum are zero, starting with the largest
appearing index~$J$: In the module~$M[a_J^{-1}]$ over the localized ring~$A[a_J^{-1}]$,
the vector~$w_J$ is a linear combination of generators with
smaller index. Removing~$w_J = x_1^{i_1} \cdots x_n^{i_n} v_\ell$ and also all
vectors~$x_1^{k_1} \cdots x_n^{k_n} v_\ell$ where~$k_1 \geq i_1, \ldots,
k_n \geq i_n$, we obtain a subfamily which is still good for
the localized module. The induction hypothesis, applied
to~$A[a_J^{-1}]$ and its module~$M[a_J^{-1}]$, therefore implies that $A[a_J^{-1}] = 0$.
Thus $a_J = 0$ since~$A$ is reduced.
\enlargethispage{2em}

Similarly, we show that the given good generating family~$(s_I)_{I \in \I'}$ is
a basis. Thus~$M$ and~$B$ are free over~$A$. We fix for any corner~$J$
of~$\J'$, as indicated in Figure~\ref{fig:good},
a way of expressing~$w_J = \sum_K a_{JK} w_K$ as a linear combination of
generators with strictly smaller index. Let~$\widehat w_{(\ell,i_1,\ldots,i_n)}
\defeq x_1^{i_1} \cdots x_n^{i_n} V_\ell$ in the free~$B$-module~$B\langle
V_1,\ldots, V_m\rangle$. The canonical map
\[ \widehat M \defeq B\langle V_1,\ldots, V_m\rangle/
  (\widehat w_J - \textstyle\sum_K a_{JK} \widehat w_K)_{\text{$J$ corner of~$\J'$}} \longrightarrow M
\]
is trivially well-defined and surjective. It is also injective, since any
element of~$\widehat M$ can be written as an~$A$-linear combination of the
vectors~$(\widehat w_J)_{J \in \J'}$ by employing the corner relations a finite
number of times.  Therefore~$M$ is finitely presented as a~$B$-module.

In a similar vein, a quotient algebra of~$A[X_1,\ldots,X_n]$, where
we mod out by a suitable ideal with as many generators as corners of~$\I'$, is
isomorphic to~$B$. Thus~$B$ is finitely presented as an~$A$-algebra.

We finish by using the assumption for~$f = 1$.
\end{proof}

\printbibliography

\end{document}

Recall that a module is Noetherian if and only if all of its submodules are
finitely generated. We call a module~$M$ over an~$A$-algebra~$B$ \emph{weakly
Noetherian} (over~$A$) if and only if, for all~$B$-submodules~$U \subseteq M$
and all~$f \in A$ the condition
\begin{multline*}
  \text{for all~$g \in A$, if $U[(fg)^{-1}]$ is finitely generated as
an~$B[(fg)^{-1}]$-module,} \\
  \text{then~$fg$ is nilpotent in~$A$} \end{multline*}
implies that~$f$ is nilpotent. Geometrically, this expresses that all
quasicoherent sheaves of submodules of~$M^\sim$ are, over a dense open
of~$\Spec(A)$, of finite type over~$B^\sim$. An~$A$-algebra~$B$
is \emph{weakly Noetherian} if and only if it is weakly Noetherian as a module
over itself.

This notion is interesting for the following reason.

\begin{lemma}\label{lemma:self-noetherian}
Let~$A$ be a reduced ring. Then~$A$, regarded as an algebra over itself, is
weakly Noetherian.\end{lemma}

\begin{proof}Let~$\aaa \subseteq A$ be an ideal. Let~$f \in A$ such that, for
any~$g \in A$, if~$\aaa[(fg)^{-1}]$ is finitely generated, then~$fg = 0$.
We verify that~$f = 0$.

We first show that~$\aaa[f^{-1}] = (0)$. Let~$g \in \aaa$. Then~$\aaa[(fg)^{-1}] =
(1)$ is finitely generated. Thus~$fg = 0$, hence~$g = 0$ in~$A[f^{-1}]$.

Since~$\aaa[f^{-1}]$ is finitely generated, we finish by using the
assumption for~$g = 1$.
\end{proof}

The weak Noetherian condition enjoys the same basic stability properties as the
usual Noetherian condition:

\begin{lemma}\label{lemma:noetherian-stability}
Let~$A$ be a ring. Let~$B$ be an~$A$-algebra.
\begin{enumerate}
\item If~$B$ is weakly Noetherian, then so is~$B[X]$.
\item Let~$0 \to M' \to M \to M'' \to 0$ be a short exact sequence
of~$B$-modules. Then~$M$ is weakly Noetherian if and only if~$M'$ and~$M''$
are.
\end{enumerate}
\end{lemma}

We can use these observations to finish the proof of Grothendieck's generic
freeness lemma.

\begin{prop}Let~$A$ be a reduced ring. Let~$B$ be an~$A$-algebra of finite type.
If~$f = 0$ is the only element of~$A$ such that~$B[f^{-1}]$ is of finite
presentation as an~$A[f^{-1}]$-algebra, then~$1 = 0$ in~$A$.
\end{prop}

\begin{proof}We write~$B = A[X_1,\ldots,X_n]/\aaa$. By Lemma~\ref{lemma:self-noetherian} and
Lemma~\ref{lemma:noetherian-stability}, the~$A$-algebra~$A[X_1,\ldots,X_n]$ is
weakly Noetherian. We can therefore conclude if we can show, for any~$g \in A$,
that if the ideal~$\aaa[g^{-1}]$ is finitely generated, then~$g = 0$. This
follows immediately from the assumption (for~$f = g$).
\end{proof}

\begin{prop}Let~$A$ be a reduced ring. Let~$B$ be an~$A$-algebra of finite type.
Let~$M$ be a finitely generated~$B$-module. If~$f = 0$ is the only element
of~$A$ such that~$M[f^{-1}]$ is finitely presented as an~$B[f^{-1}]$-module,
then~$1 = 0$ in~$A$.
\end{prop}

\begin{proof}We write~$M = B^n/U$ for an~$B$-submodule~$U$. By Lemma~\ref{lemma:self-noetherian} and
Lemma~\ref{lemma:noetherian-stability}, the~$B$-module~$B^n$ is weakly
Noetherian. We can therefore conclude if we can show, for any~$g \in A$, that if
the submodule~$U[g^{-1}]$ is finitely generated, then~$g = 0$. This follows
immediately from the assumption (for~$f = g$).
\end{proof}

\end{document}