### 2018-05-17 paper-wlog.tex
Beginnings of the "without loss of generality" paper

\documentclass{amsart}

\usepackage[utf8]{inputenc}
\usepackage{amsthm,mathtools,tikz,float,caption,xspace}
\usetikzlibrary{patterns,matrix}
\usepackage[protrusion=true,expansion=true]{microtype}
\usepackage{hyperref}
\usepackage{cleveref}

\usepackage[natbib=true,style=numeric,maxnames=10]{biblatex}
\usepackage[babel]{csquotes}
\bibliography{paper-wlog.bib}

\title{Without loss of generality, \\ any reduced ring is a field}
\author{Ingo Blechschmidt}

\theoremstyle{definition}
\newtheorem{defn}{Definition}[]
\newtheorem{ex}[defn]{Example}

\theoremstyle{plain}
\newtheorem{prop}[defn]{Proposition}
\newtheorem{cor}[defn]{Corollary}
\newtheorem{lemma}[defn]{Lemma}
\newtheorem{thm}[defn]{Theorem}
\newtheorem{scholium}[defn]{Scholium}

\theoremstyle{remark}
\newtheorem{rem}[defn]{Remark}
\newtheorem{question}[defn]{Question}
\newtheorem{speculation}[defn]{Speculation}
\newtheorem{caveat}[defn]{Caveat}
\newtheorem{conjecture}[defn]{Conjecture}

\crefname{defn}{Definition}{Definitions}
\crefname{thm}{Theorem}{Theorems}
\crefname{lemma}{Lemma}{Lemmas}
\crefname{prop}{Proposition}{Propositions}
\crefname{ex}{Example}{Examples}
\crefname{section}{Section}{Sections}

% Workaround from David Carlisle described at https://tex.stackexchange.com/a/316037
\makeatletter
\let\xx@thm\@thm
\AtBeginDocument{\let\@thm\xx@thm}
\makeatother

\newcommand{\XXX}[1]{\textbf{XXX: #1}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\bbb}{\mathfrak{b}}
\newcommand{\mmm}{\mathfrak{m}}
\newcommand{\ppp}{\mathfrak{p}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\defeq}{\vcentcolon=}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\Quot}{Quot}
\newcommand{\?}{\,{:}\,}
\newcommand{\notnot}{\emph{not not}\xspace}
\newcommand{\negg}{\neg\!\!\!\neg}

\newcommand{\stacksproject}[1]{\cite[{\href{https://stacks.math.columbia.edu/tag/#1}{Tag~#1}}]{stacks-project}}

\begin{document}

\begin{abstract}
  XXX
\end{abstract}

\maketitle
\thispagestyle{empty}

\noindent
XXX

\section{Case studies}

\subsection{Linear surjections between finite free modules}

\begin{prop}\label{surjection}
Let~$M \in A^{n \times m}$ be a matrix over a ring~$A$ such that~$n
> m$. Assume that the induced linear map~$A^m \to A^n$ is surjective. Then~$1 =
0 \in A$.
\end{prop}

\begin{proof}[Proof~A (classical)]Assume to the contrary that~$1 \neq 0 \in A$.
Then there exists a maximal ideal~$\mmm \subseteq A$. Tensoring the given linear map
over~$A$ with~$A/\mmm$ yields a linear map~$(A/\mmm)^m \to (A/\mmm)^n$ which is
still surjective, since tensoring is right exact. Since~$A/\mmm$ is a field,
this is a contradiction by basic linear algebra.
\end{proof}

This proof uses the axiom of choice in the guise of the lemma that nontrivial
rings contain maximal ideals in order to reduce from general rings to
fields.\footnote{It's possible to avoid the full axiom of choice here and only
assume the Boolean Prime Ideal Theorem, at the expense of slightly complicating
the proof. This theorem is strong enough to conclude that a nontrivial ring has
a prime ideal~$\ppp$. One then first passes to the quotient~$A/\ppp$, where
surjectivity is preserved thanks to right-exactness of tensoring, and then to
the field of fractions~$\Quot(A/\ppp)$, where surjectivity is preserved thanks
to exactness of localization.} In fact, the statement even admits a fully
constructive proof. It's not hard to give such a proof explicitly, for instance
using determinants~\cite{richman:trivial-rings}, but in this section we want to
see how the forcing technique can be used to translate the classical proof into
a constructive one.

Starting point is the observation that \Cref{surjection} admits an easy proof
in the case that the ring~$A$ is local:

\begin{proof}[Proof~B (constructive, but only of the special case
that~$A$ is a local ring)]
Since~$M$ is surjective, some linear combination of its columns sums to the
first canonical basis vector~$e_1$. Since~$A$ is local, some entry of the first row
is invertible. By employing elementary row and column transformations, we can
change this entry to the unit of the ring, move it to the top left, and kill
any remaining entries in the first row and in the first column. Continuing in
this fashion, we can assume~$A$ to be a rectangular diagonal matrix. Since~$A$
has more rows than columns, all entries in the last row are zero. On the other
hand, one of its entries is the unit. Thus~$1 = 0$.
\end{proof}

\begin{proof}[Proof~C (constructive, general case)]
By~XXX, the given matrix~$M$ is also surjective as a matrix over~$A^\sim$.
Since~$A^\sim$ is a local ring, the previous proof applies and shows
that~$A^\sim \models 1 = 0 \? A^\sim$. This is equivalent to~$1 = 0 \in A$ by
definition.
\end{proof}

This attempt succeeds in eliminating the axiom of choice and the law of
excluded middle; but it isn't yet a faithful translation of Proof~A, since the
standard linear algebra proof referenced in Proof~A proceeds slightly
differently than Proof~B. To obtain such a translation, we use the special
field property of~$A^\sim$.

\begin{proof}[Proof~D (constructive, general case)]
By passing from~$A$ to its quotient ring~$A/\sqrt{(0)}$, which preserves
surjectivity since tensoring is right-exact, we may assume that~$A$ is a
reduced ring. Thus~$A^\sim$ is a field in the sense that non-invertible
elements are zero. It therefore suffices to give a constructive proof of
\Cref{surjection} in the special that~$A$ is such a kind of field.

Every entry of the given matrix is \notnot invertible or not invertible. Since
the statement we want to verify,~$1 = 0$, is~$\neg\neg$-stable, we may
assume that every entry is actually invertible or not invertible. If
one of the entries is invertible, we can change it to the unit, move it to the
top left, kill the remaining entries in the first row and in the first
column, and then continue in this fashion. We may thus assume that~$M$ is a
rectangular diagonal matrix having only zeros and ones on the diagonal.

XXX
\end{proof}

\subsection{Linear injections between finite free modules: McCoy's theorem}

\begin{proof}Let~$a \? A^\sim$ such that~$a \cdot (\Lambda^m M) = (0)$. We want
to verify that~$a = 0$. Instead of employing~$\neg\neg$ like on previous
occasions, we employ~$\negg\negg$, where~$\negg\varphi$ doesn't mean~$(\varphi
\Rightarrow \bot)$, but~$(\varphi \Rightarrow a = 0)$.

For any entry~$M_{ij}$ of the matrix, we~$\negg\negg$-have that it's invertible
or~$\negg($invertible$)$. The latter is equivalent to some power XXX
\end{proof}

\subsection{Nilpotent polynomials}

XXX

\subsection{Grothendieck's generic freeness lemma}

XXX

\end{document}