### 2018-05-25 slides-filmat2018.tex
Nicer title slide and slide on applications

\documentclass[12pt,utf8,notheorems,compress,t]{beamer}
\usepackage{etex}

\usepackage[english]{babel}

\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{multicol}
\usepackage{tabto}
\usepackage{xstring}
\usepackage{ifthen}
\usepackage{soul}\setul{0.3ex}{}
\usepackage[all]{xy}
\xyoption{rotate}
\usepackage{tikz}
\usetikzlibrary{calc,shapes,shapes.callouts,shapes.arrows,patterns,fit,backgrounds,decorations.pathmorphing}
\hypersetup{colorlinks=true}
\usepackage{multimedia}
\newcommand{\video}[2]{\movie[width=#2,height=#2,autostart,loop,poster]{}{#1}}
\hypersetup{colorlinks=false}

\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\graphicspath{{images/}}

\usepackage[protrusion=true,expansion=true]{microtype}

\setlength\parskip{\medskipamount}
\setlength\parindent{0pt}

\title{Exploring mathematical objects from custom-tailored mathematical universes}
\author{Ingo Blechschmidt}
\date{May 26th, 2018}

\useinnertheme[shadow=true]{rounded}
\useoutertheme[subsection=false]{miniframes}
\setbeamerfont{block title}{size={}}
\addtobeamertemplate{headline}{\vskip-0.5ex}{}

\useinnertheme{rectangles}

\usecolortheme{orchid}
\usecolortheme{seahorse}
\definecolor{mypurple}{RGB}{150,0,255}
\setbeamercolor{structure}{fg=mypurple}
\definecolor{myred}{RGB}{150,0,0}
\setbeamercolor*{title}{bg=myred,fg=white}
\setbeamercolor*{titlelike}{bg=myred,fg=white}

\usefonttheme{serif}
\usepackage[T1]{fontenc}
\usepackage{libertine}

\newcommand{\A}{\mathcal{A}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\renewcommand{\G}{\mathcal{G}}
\newcommand{\GG}{\mathbb{G}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\ZZ}{\mathbb{Z}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\ppp}{\mathfrak{p}}
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\defeqv}{\vcentcolon\equiv}
\newcommand{\Sh}{\mathrm{Sh}}
\newcommand{\GL}{\mathrm{GL}}
\newcommand{\Zar}{\mathrm{Zar}}
\newcommand{\op}{\mathrm{op}}
\newcommand{\Set}{\mathrm{Set}}
\newcommand{\Eff}{\mathrm{Ef{}f}}
\newcommand{\Sch}{\mathrm{Sch}}
\newcommand{\Aff}{\mathrm{Aff}}
\newcommand{\LRS}{\mathrm{LRS}}
\newcommand{\Hom}{\mathrm{Hom}}
\newcommand{\Spec}{\mathrm{Spec}}
\newcommand{\lra}{\longrightarrow}
\newcommand{\RelSpec}{\operatorname{Spec}}
\renewcommand{\_}{\mathpunct{.}}
\newcommand{\?}{\,{:}\,}
\newcommand{\speak}[1]{\ulcorner\text{\textnormal{#1}}\urcorner}
\newcommand{\ull}[1]{\underline{#1}}
\newcommand{\affl}{\ensuremath{{\ull{\AA}^1}}}
\newcommand{\Ll}{\vcentcolon\!\Longleftrightarrow}
\newcommand{\inv}{inv.\@}
\newcommand{\seq}{\vdash_{\!\!\!\vec x}}

\setbeamertemplate{blocks}[rounded][shadow=false]

% Adapted from https://latex.org/forum/viewtopic.php?t=2251 (Stefan Kottwitz)
\newenvironment<>{hilblock}{
  \begin{center}
    \begin{minipage}{9.05cm}
      \setlength{\textwidth}{9.05cm}
      \begin{actionenv}#1
        \def\insertblocktitle{}
        \par
        \usebeamertemplate{block begin}}{
        \par
        \usebeamertemplate{block end}
      \end{actionenv}
    \end{minipage}
  \end{center}}

\newenvironment{changemargin}[2]{%
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \setlength{\leftmargin}{#1}%
    \setlength{\rightmargin}{#2}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\itemindent}{\parindent}%
    \setlength{\parsep}{\parskip}%
  }%
  \item[]}{\end{list}}

\tikzset{
  invisible/.style={opacity=0,text opacity=0},
  visible on/.style={alt={#1{}{invisible}}},
  alt/.code args={<#1>#2#3}{%
    \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}}}
}

\newcommand{\pointthis}[3]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=0,outer sep=0] (#2) {#2};
    \node[visible on=#1,overlay,rectangle callout,rounded corners,callout relative pointer={(0.3cm,0.5cm)},fill=blue!20] at ($(#2.north)+(-0.1cm,-1.1cm)$) {#3};
  }%
}

% Adapted from https://latex.org/forum/viewtopic.php?t=2251 (Stefan Kottwitz)
\newenvironment<>{varblock}[2]{
  \begin{center}
    \begin{minipage}{#1}
      \setlength{\textwidth}{#1}
      \begin{actionenv}#3
	\def\insertblocktitle{\centering #2}
	\par
	\usebeamertemplate{block begin}}{
        \par
        \usebeamertemplate{block end}
      \end{actionenv}
    \end{minipage}
  \end{center}}

\setbeamertemplate{frametitle}{%
  \vskip0.7em%
  \leavevmode%
  \begin{beamercolorbox}[dp=1ex,center]{}%
      \usebeamercolor[fg]{item}{\textbf{{\Large \insertframetitle}}}
  \end{beamercolorbox}%
}

\setbeamertemplate{navigation symbols}{}

\newcounter{framenumberpreappendix}
\newcommand{\backupstart}{
  \setcounter{framenumberpreappendix}{\value{framenumber}}
}
\newcommand{\backupend}{
  \addtocounter{framenumberpreappendix}{-\value{framenumber}}
  \addtocounter{framenumber}{\value{framenumberpreappendix}} 
}

\setbeamertemplate{footline}{%
  \begin{beamercolorbox}[wd=\paperwidth,ht=2.25ex,dp=1ex,right,rightskip=1mm,leftskip=1mm]{}%
    \inserttitle
    \hfill
    \insertframenumber\,/\,\inserttotalframenumber
  \end{beamercolorbox}%
  \vskip0pt%
}

\newcommand{\hil}[1]{{\usebeamercolor[fg]{item}{\textbf{#1}}}}

\IfSubStr{\jobname}{\detokenize{nonotes}}{
  \setbeameroption{hide notes}
}{
  \setbeameroption{show notes}
}
\setbeamertemplate{note page}[plain]
\setbeameroption{hide notes}

\begin{document}

\addtocounter{framenumber}{-1}

{\title{}
\begin{frame}[c]
  \centering
  \medskip
  \includegraphics[height=0.4\textheight]{acquaviva-platani-street}%
  \includegraphics[height=0.4\textheight]{mussomeli-sheep}%
  \includegraphics[height=0.4\textheight]{mussomeli-performance}%
  \includegraphics[height=0.4\textheight]{mussomeli-castle}%
  \medskip

  \hil{Exploring mathematical objects from \\ custom-tailored mathematical universes} \\

  \emph{-- an invitation --}

  \bigskip

  \scriptsize
  Ingo Blechschmidt \\
  University of Augsburg / \\
  Max Planck Institute for Mathematics in the Sciences, Leipzig \\

  \medskip

  Third international conference of the \\
  \emph{Italian Network for the Philosophy of Mathematics} \\
  in Mussomeli
  \medskip

  May 26th, 2018
  \par
\end{frame}}

\section[Toposophic landscape]{A glimpse of the toposophic landscape}

\begin{frame}[fragile]{A glimpse of the toposophic landscape}
  \tikzstyle{topos} = [draw=mypurple, very thick, rectangle, rounded corners, inner sep=5pt, inner ysep=10pt]
  \tikzstyle{title} = [fill=mypurple, text=white]

  \input{images/primes.tex}
  %\renewcommand{\sieve}[2]{SIEVE}
  %\renewcommand{\fakesieve}[2]{SIEVE}

  \newcommand{\drawbox}[4]{
    \node[topos, #4] [fit = #3] (#1) {};
    \node[title] at (#1.north) {#2};
  }

  \newcommand{\muchstuff}{
    \includegraphics[height=3em]{filmat}
    \scalebox{0.5}{\sieve{14}{2}}
  }

  \newcommand{\muchstuffplaceholder}{
    \includegraphics[height=3em]{filmat-placeholder}
    \scalebox{0.5}{\fakesieve{14}{2}}
  }

  \newcommand{\fewstuff}{
    \includegraphics[height=3em]{filmat}
    \scalebox{0.5}{\sieve{7}{2}}
  }

  \newcommand{\threeblobs}{
    \colorbox{mypurple}{\ \ }\quad
    \colorbox{mypurple}{\ \ }\quad
    \colorbox{mypurple}{\ \ }
  }

  \begin{tikzpicture}
    \node (objs-set0) at (0,0) {
      \only<1-2>{\muchstuffplaceholder}
      \only<3>{\muchstuff}
    };
    \node[scale=0.4] (objs-set1) at (-3.5,-2.5) {
      \only<1->{\fewstuff}
    };
    \node[scale=0.4] (objs-eff1) at (3.5,-2.5) {
      \only<2->{\fewstuff}
    };
    \node[scale=0.4] (objs-sh1)  at (0,-2.5) {
      \only<2->{\fewstuff}
    };

    \node (prop-set1) [below of=objs-set1, align=left] {
      \only<1->{%
        The usual laws \\
        of logic hold.
      }
    };

    \node (prop-eff1) [below of=objs-eff1, align=left] {
      \only<2->{%
        Every function \\
        is computable.
      }
    };

    \node (prop-sh1) [below of=objs-sh1, align=left] {
      \only<2->{%
        The axiom of \\
        choice fails.
      }
    };

    \node (more-eff1) [below of=prop-eff1, visible on=<3->] {
      \threeblobs
    };
    \node (more-sh1)  [below of=prop-sh1, visible on=<3->] {
      \threeblobs
    };
    \node (more-set1) [below of=prop-set1, visible on=<3->] {
      \threeblobs
    };

    \begin{scope}[visible on=<1->]
      \drawbox{set1}{$\mathrm{Set}$}{(objs-set1) (prop-set1) (more-set1)}{}
    \end{scope}
    \begin{scope}[visible on=<2->]
      \drawbox{eff1}{Ef{}f}{(objs-eff1) (prop-eff1) (more-eff1)}{tape}
    \end{scope}
    \begin{scope}[visible on=<2->]
      \drawbox{sh1}{$\mathrm{Sh}\, X$}{(objs-sh1) (prop-sh1) (more-sh1)}{draw=none}
      \def\R{8pt}
      \begin{pgfonlayer}{background}
      \draw[decoration={bumps,segment length=8pt}, decorate, very thick, draw=mypurple, visible on=<2->]
        ($(sh1.south west) + (\R, 0)$) arc(270:180:\R) --
        ($(sh1.north west) + (0, -\R)$) arc(180:90:\R) --
        ($(sh1.north east) + (-\R, 0)$) arc(90:0:\R) --
        ($(sh1.south east) + (0, \R)$) arc(0:-90:\R) --
        cycle;
      \end{pgfonlayer}
    \end{scope}
    \begin{scope}[visible on=<3->]
      \drawbox{set0}{$\mathrm{Set}$}{(objs-set0) (set1) (eff1) (sh1)}{}
    \end{scope}
  \end{tikzpicture}
\end{frame}

\section[Internal universe]{The internal universe of a topos}

\begin{frame}{The internal universe of a topos}
  For any topos~$\E$ and any statement~$\varphi$, we define the meaning of
  \vspace*{-0.5em}
  \[
    \text{``$\E \models \varphi$''} \quad
    \text{(``$\varphi$ holds in the internal universe of~$\E$'')}
  \]

  \vspace*{-1.0em}
  using the \hil{Kripke--Joyal semantics}.

  \vspace*{-1em}
  \begin{columns}
    \def\insertblocktitle{}
    \begin{column}{0.3\textwidth}\usebeamertemplate{block begin}
      \centering
      $\Set \models \varphi$ \\
      ``$\varphi$ holds in the usual sense.''
    \usebeamertemplate{block end}\end{column}

    \begin{column}{0.3\textwidth}\usebeamertemplate{block begin}
      \centering
      $\Sh(X) \models \varphi$ \\
      ``$\varphi$ holds continuously.''
    \usebeamertemplate{block end}\end{column}

    \begin{column}{0.3\textwidth}\usebeamertemplate{block begin}
      \centering
      $\Eff \models \varphi$ \\
      ``$\varphi$ holds computably.''
    \usebeamertemplate{block end}\end{column}
  \end{columns}
  \medskip

  \pause
  Any topos supports \hil{mathematical reasoning}:

  \vspace*{-1em}
  \begin{hilblock}
    If~$\E \models \varphi$ and if~$\varphi \vdash \psi$
    \pointthis{<3>}{intuitionistically}{%
      no $\varphi \vee \neg\varphi$,\ \
      no $\neg\neg\varphi \Rightarrow \varphi$,\ \
      no axiom of choice},
    then~$\E \models \psi$.
  \end{hilblock}
\end{frame}

\section[First steps]{First steps in alternate universes}

\newcommand{\intex}[3]{#1 \quad #2\par #3\bigskip\medskip}

\begin{frame}{First steps in alternate universes}
  \begin{changemargin}{-1.1em}{0em}
    \fontsize{10pt}{12pt}\selectfont
    \begin{itemize}
      \item \intex{
        $\Eff \models \text{``Any number is prime or is not prime.''}$
      }{\textcolor{green!90}{\cmark}}{
        There is a \hil{Turing machine} which determines of
        any given number whether it is prime or not.
      }

      \item \intex{
        $\Eff \models \text{``There are infinitely many prime numbers.''}$
      }{\textcolor{green!90}{\cmark}}{
        There is a \hil{Turing machine} producing arbitrarily many prime
        numbers.
      }

      \item \intex{
        $\Eff \models \text{``Any function~$\NN \to \NN$ is the zero function or not.''}$
      }{\textcolor{red!80}{\xmark}}{
        There is a \hil{Turing machine} which, given a Turing
        machine computing a function~$f : \NN \to \NN$, determines whether~$f$
        is the zero function or not.
      }

      \item \intex{
        $\Eff \models \text{``Any function~$\NN \to \NN$ is computable.''}$
      }{\textcolor{green!90}{\cmark}}{}

      \item \intex{
        $\Sh(X) \models \text{``Any cont. function with opposite signs has
        a zero.''}$
      }{\textcolor{red!80}{\xmark}}{
        Zeros can locally be picked \hil{continuously} in
        continuous families of continuous functions.
        \textcolor{red}{(\href{https://rawgit.com/iblech/internal-methods/master/images/zeros-in-families.mp4}{video} for counterexample)}
      }
    \end{itemize}
  \end{changemargin}
\end{frame}

\section[Applications]{Applications in commutative algebra}

\begin{frame}{Applications in commutative algebra}
  Let~$A$ be a reduced commutative ring.

  \hil{For example:}\ \ $\ZZ$,\ \ $\ZZ[X]$,\ \ $\ZZ[X,Y,Z]/(X^n+Y^n-Z^n)$,\ \ $\QQ$,\ \ $\RR$

  \mbox{The \hil{little Zariski topos} of~$A$ contains a \hil{mirror image} of~$A$:
  $A^\sim$.}

  \setlength\columnsep{0pt}
  \begin{multicols*}{2}
    \begin{enumerate}
      \item $A^\sim$ is always a \hil{field}. \columnbreak
      \item $A^\sim$ is still \hil{very close} to~$A$.
    \end{enumerate}
  \end{multicols*}
  % \item Studying~$A^\sim$ is the same as studying~$A$ \hil{from a different
  % point of view}.

  \vspace*{-2em}
  \begin{columns}[t]
    \begin{column}[t]{0.47\textwidth}
      \centering

      \begin{varblock}{\textwidth}{A baby application}
        \justifying
        Let~$M$ be a surjective matrix with more rows than columns over a ring~$A$.
        Then~$A = 0$.
      \end{varblock}

      \scalebox{0.8}{$\begin{pmatrix}
        \cdot & \cdot \\
        \cdot & \cdot \\
        \cdot & \cdot
      \end{pmatrix}$}
    \end{column}

    \begin{column}[t]{0.47\textwidth}
      \centering

      \begin{varblock}{\textwidth}{Generic freeness\phantom{p}}
        \justifying
        Generically, any finitely generated module over a reduced ring is free.
      \end{varblock}
      \vspace*{-0.5em}

      \includegraphics[width=0.6\textwidth]{generic-freeness}
    \end{column}
  \end{columns}
\end{frame}

\section{Conclusion}

\begin{frame}{Topos theory \ldots}
  \begin{itemize}
    \item enriches the platonism debate,
    %\item yields a new view on ``continuous dependence on parameters'',
    \item uncovers further relations between objects,
    \item allows to view objects from a different point of view,
    \item has applications in mathematical practice.
  \end{itemize}

  \centering
  \href{https://www.oliviacaramello.com/Papers/Papers.htm}{\includegraphics[height=0.45\textheight]{olivia-tst}}
  \href{http://math.andrej.com/2014/03/04/intuitionistic-mathematics-and-realizability-in-the-physical-world/}{\includegraphics[height=0.45\textheight]{zenil-computable-universe}}
  \href{https://pizzaseminar.speicherleck.de/skript2/zariski-topos-klein.pdf}{\includegraphics[height=0.45\textheight]{fun-with-the-zariski-topos}}
  \href{https://rawgit.com/iblech/internal-methods/master/notes.pdf}{\includegraphics[height=0.45\textheight]{phd-cover}}
\end{frame}

\end{document}