### 2018-07-10 slides-leipzig2018.tex
Beginning of notes for the Leipzig talk

\documentclass[12pt,utf8,notheorems,compress,t]{beamer}
\usepackage{etex}

\usepackage{pgfpages}
\setbeameroption{show notes on second screen}
\setbeamertemplate{note page}[plain]
\newcommand{\jnote}[2]{\only<#1>{\note{\footnotesize\justifying#2\par}}}

\usepackage[english]{babel}

\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{stmaryrd}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{multicol}
\usepackage{tabto}
\usepackage{xstring}
\usepackage{ifthen}
\usepackage{soul}\setul{0.3ex}{}
\usepackage[all]{xy}
\xyoption{rotate}
\usepackage{tikz}
\usetikzlibrary{mindmap,calc,shapes,shapes.callouts,shapes.arrows,patterns,fit,backgrounds,decorations.pathmorphing}
\hypersetup{colorlinks=false}

\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\graphicspath{{images/}}

\usepackage[protrusion=true,expansion=true]{microtype}

\setlength\parskip{\medskipamount}
\setlength\parindent{0pt}

\title{How topos theory can help algebra and geometry}
\author{Ingo Blechschmidt}
\date{July 10th, 2018}

\useinnertheme[shadow=true]{rounded}
\useoutertheme[subsection=false]{miniframes}
\setbeamerfont{block title}{size={}}

\useinnertheme{rectangles}

\usecolortheme{orchid}
\usecolortheme{seahorse}
\definecolor{mypurple}{RGB}{150,0,255}
\setbeamercolor{structure}{fg=mypurple}
\definecolor{myred}{RGB}{150,0,0}
\setbeamercolor*{title}{bg=myred,fg=white}
\setbeamercolor*{titlelike}{bg=myred,fg=white}
\setbeamercolor{frame}{bg=black}

\usefonttheme{serif}
\usepackage[T1]{fontenc}
\usepackage{libertine}

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\renewcommand{\AA}{\mathbb{A}}
\renewcommand{\C}{\mathcal{C}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\M}{\mathcal{M}}
\renewcommand{\G}{\mathcal{G}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\GG}{\mathbb{G}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\ZZ}{\mathbb{Z}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\fff}{\mathfrak{f}}
\newcommand{\ppp}{\mathfrak{p}}
\newcommand{\mmm}{\mathfrak{m}}
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\defeqv}{\vcentcolon\equiv}
\newcommand{\Sh}{\mathrm{Sh}}
\newcommand{\GL}{\mathrm{GL}}
\newcommand{\Zar}{\mathrm{Zar}}
\newcommand{\op}{\mathrm{op}}
\newcommand{\Set}{\mathrm{Set}}
\newcommand{\Ring}{\mathrm{Ring}}
\newcommand{\LocRing}{\mathrm{LocRing}}
\newcommand{\Eff}{\mathrm{Ef{}f}}
\newcommand{\Sch}{\mathrm{Sch}}
\newcommand{\Aff}{\mathrm{Aff}}
\newcommand{\LRS}{\mathrm{LRS}}
\newcommand{\Hom}{\mathrm{Hom}}
\newcommand{\Spec}{\mathrm{Spec}}
\newcommand{\lra}{\longrightarrow}
\newcommand{\RelSpec}{\operatorname{Spec}}
\renewcommand{\_}{\mathpunct{.}}
\newcommand{\?}{\,{:}\,}
\newcommand{\speak}[1]{\ulcorner\text{\textnormal{#1}}\urcorner}
\newcommand{\ull}[1]{\underline{#1}}
\newcommand{\affl}{\ensuremath{{\ull{\AA}^1}}}
\newcommand{\Ll}{\text{iff}}
\newcommand{\inv}{inv.\@}
\newcommand{\seq}{\vdash_{\!\!\!\vec x}}
\newcommand{\hg}{\mathbin{:}}  % homogeneous coordinates

\setbeamertemplate{blocks}[rounded][shadow=false]

% Adapted from https://latex.org/forum/viewtopic.php?t=2251 (Stefan Kottwitz)
\newenvironment<>{hilblock}{
  \begin{center}
    \begin{minipage}{9.85cm}
      \setlength{\textwidth}{9.85cm}
      \begin{actionenv}#1
        \def\insertblocktitle{}
        \par
        \usebeamertemplate{block begin}}{
        \par
        \usebeamertemplate{block end}
      \end{actionenv}
    \end{minipage}
  \end{center}}

\newcommand{\bignumber}[1]{
  \renewcommand{\insertenumlabel}{#1}\scalebox{1.5}{\usebeamertemplate{enumerate item}}
}

\newenvironment{indentblock}{%
  \list{}{\leftmargin\leftmargin}%
  \item\relax
}{%
  \endlist
}

\newcommand{\hcancel}[5]{%
  \tikz[baseline=(tocancel.base)]{
    \node[inner sep=0pt,outer sep=0pt] (tocancel) {#1};
    \draw[red, line width=0.4mm] ($(tocancel.south west)+(#2,#3)$) -- ($(tocancel.north east)+(#4,#5)$);
  }%
}

\newenvironment{changemargin}[2]{%
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \setlength{\leftmargin}{#1}%
    \setlength{\rightmargin}{#2}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\itemindent}{\parindent}%
    \setlength{\parsep}{\parskip}%
  }%
  \item[]}{\end{list}}

\tikzset{
  invisible/.style={opacity=0,text opacity=0},
  visible on/.style={alt={#1{}{invisible}}},
  alt/.code args={<#1>#2#3}{%
    \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}}}
}

\newcommand{\pointthis}[3]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=0,outer sep=0] (#2) {#2};
    \node[visible on=#1,overlay,rectangle callout,rounded corners,callout relative pointer={(-0.1cm,0.5cm)},fill=blue!20] at ($(#2.north)+(-0.1cm,-1.1cm)$) {#3};
  }%
}

% Adapted from https://latex.org/forum/viewtopic.php?t=2251 (Stefan Kottwitz)
\newenvironment<>{varblock}[2]{
  \begin{center}
    \begin{minipage}{#1}
      %\setlength{\textwidth}{#1}
      \begin{actionenv}#3
  \def\insertblocktitle{\centering #2}
  \par
  \usebeamertemplate{block begin}}{
        \par
        \usebeamertemplate{block end}
      \end{actionenv}
    \end{minipage}
  \end{center}}

\setbeamertemplate{frametitle}{%
  \vskip0.7em%
  \leavevmode%
  \begin{beamercolorbox}[dp=1ex,center]{}%
      \usebeamercolor[fg]{item}{\textbf{{\Large \insertframetitle}}}
  \end{beamercolorbox}%
}

\setbeamertemplate{navigation symbols}{}

\newcounter{framenumberpreappendix}
\newcommand{\backupstart}{
  \setcounter{framenumberpreappendix}{\value{framenumber}}
}
\newcommand{\backupend}{
  \addtocounter{framenumberpreappendix}{-\value{framenumber}}
  \addtocounter{framenumber}{\value{framenumberpreappendix}}
}

\setbeamertemplate{footline}{%
  \begin{beamercolorbox}[wd=\paperwidth,ht=2.25ex,dp=1ex,right,rightskip=1mm,leftskip=1mm]{}%
    % \inserttitle
    \hfill
    \insertframenumber\,/\,\inserttotalframenumber
  \end{beamercolorbox}%
  \vskip0pt%
}

\newcommand{\hil}[1]{{\usebeamercolor[fg]{item}{\textbf{#1}}}}

\newcommand{\bad}[1]{\textcolor{red!90}{\textnormal{#1}}}

\begin{document}

\tikzstyle{topos} = [draw=mypurple, very thick, rectangle, rounded corners, inner sep=5pt, inner ysep=10pt]
\tikzstyle{title} = [fill=mypurple, text=white]

\input{images/primes.tex}
%\renewcommand{\sieve}[2]{SIEVE}
%\renewcommand{\fakesieve}[2]{SIEVE}

\newcommand{\drawbox}[4]{
  \node[topos, #4] [fit = #3] (#1) {};
  \node[title] at (#1.north) {#2};
}

\newcommand{\muchstuff}{
  \includegraphics[height=3em]{filmat}
  \scalebox{0.5}{\sieve{14}{2}}
}

\newcommand{\muchstuffplaceholder}{
  \includegraphics[height=3em]{filmat-placeholder}
  \scalebox{0.5}{\fakesieve{14}{2}}
}

\newcommand{\fewstuff}{
  \includegraphics[height=3em]{filmat}
  \scalebox{0.5}{\sieve{7}{2}}
}

\newcommand{\threeblobs}{
  \colorbox{mypurple}{\ \ }\quad
  \colorbox{mypurple}{\ \ }\quad
  \colorbox{mypurple}{\ \ }
}

\addtocounter{framenumber}{-1}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{4.95cm}\includegraphics[width=\paperwidth]{topos-horses}\end{minipage}}
\begin{frame}[c]
  \centering
  \medskip
  \hspace*{-3.9em}%

  \hil{How topos theory can help algebra and geometry} \\

  \emph{-- an invitation --}

  \vfill

  \scriptsize
  Ingo Blechschmidt \\
  July 10th, 2018
  \par
\end{frame}}

\begin{frame}{Toposes are \ldots}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \centering
      \hil{generalized spaces}
      \medskip

      \includegraphics[height=0.32\textwidth]{grothendieck}
      \quad
      \includegraphics[height=0.32\textwidth]{elliptic-curve}

      \footnotesize
      étale topos of a scheme \\
      field with one element
    \end{column}

    \begin{column}{0.5\textwidth}
      \centering
      \hil{mathematical universes}
      \medskip

      \begin{tikzpicture}
        \node (overview) {
          \scalebox{0.44}{\sieve{6}{4}}
        };
        \def\R{8pt}
        \begin{pgfonlayer}{background}
        \draw[decoration={bumps,segment length=8pt}, decorate, very thick, draw=mypurple]
          ($(overview.south west) + (\R, 0)$) arc(270:180:\R) --
          ($(overview.north west) + (0, -\R)$) arc(180:90:\R) --
          ($(overview.north east) + (-\R, 0)$) arc(90:0:\R) --
          ($(overview.south east) + (0, \R)$) arc(0:-90:\R) --
          cycle;
        \end{pgfonlayer}
      \end{tikzpicture}
    \end{column}
  \end{columns}

  \bigskip
  \bigskip

  \begin{columns}
    \begin{column}{0.5\textwidth}
      \centering
      \hil{categories of sheaves}
      \medskip

      \includegraphics[height=0.304\textwidth]{sheaf}

      \footnotesize
      A topos is a finitely complete cartesian closed category with a
      subobject classifier.
    \end{column}

    \begin{column}{0.5\textwidth}
      \centering
      \hil{embodiments of theories}
      \medskip

      \includegraphics[height=0.32\textwidth]{olivia-lattices}

      \footnotesize
      ``Let~$G$ be a group.''
    \end{column}
  \end{columns}

  \jnote{1}{Toposes were invented in the 1960s by Grothendieck in order to
  solve concrete problems in algebraic geometry. Their raison d'être is the
  following. In algebraic geometry, we want (and sometimes have to) work over
  fields other than~$\RR$ and~$\CC$ and even over arbitrary commutative rings.
  For the geometric objects in such settings, the \emph{schemes}, the Euclidean
  topology is not available; we have to make do with the \emph{Zariski
  topology}. However, important tools as singular cohomology don't work well
  with this topology.

  The problem was solved by inventing \emph{étale topology} as an enhancement
  of the Zariski topology. However, contrary to its name, the étale topology
  isn't a true topology in the usual sense. Putting the étale topology on a
  scheme doesn't yield a refined topological space, but a \emph{topos}.

  Toposes generalize topological spaces in two ways: Firstly, the ``open sets''
  of toposes don't actually have to be sets of points; they can be more general
  kinds of objects such as coverings. Toposes can even have no points at all
  and still be nontrivial (this is for instance the case for the \emph{topos of
  random sequences}). Secondly, while classically a given open
  subset is either contained in a further open subset or not, the opens of
  toposes can be contained in further opens in many different ways.

  Recently, toposes are being used to help the mysterious field with
  one element come into being.}

  \jnote{2}{Since the 1960s, many more aspects of toposes were discovered.
  A popular reference on toposes starts with a list of 13 ways of viewing
  toposes.

  In this talk, we'll focus on the view of toposes as \emph{alternate
  mathematical universes}. We can do mathematics inside these alternate
  universes just as well as we can do mathematics inside the \emph{standard
  universe} (which is represented by a particular topos called~``$\Set$'').
  Each topos contains own versions of all the familiar mathematical objects --
  numbers, functions, manifolds -- but the properties they enjoy differ
  slightly from topos to topos.

  The definition of what a topos is, displayed in the lower left of the slide,
  has two problems. Firstly, it's only useful if one knows the relevant
  category-theoretic jargon. Secondly, a topos has lots of further vital
  structure, which is crucial for a rounded understanding. However, the
  displayed definition is trimmed for minimality and therefore doesn't mention
  it. A more comprehensive definition is: A \emph{topos} is a locally cartesian
  closed, finitely complete and cocomplete Heyting category which is exact,
  extensive and has a subobject classifier. We won't need either definition in
  this talk.}

  \jnote{3}{Toposes can also be thought as reifying the ``semantic essence'' of
  mathematical theories (the theory of groups, the theory of rings, the theory
  of intervals, \ldots). Given any such theory~$T$, there is a so-called
  \emph{classifying topos}~$\Set[T]$. Its points are precisely the set-based
  models of~$T$ (the groups, the rings, the intervals, \ldots), and it contains
  a \emph{generic model} which has exactly those properties which all models
  have. (This generic model is what mathematicians implicitly refer to when
  they say ``Let~$G$ be a group.''.)

  Crucially, two theories~$T$ and~$T'$ can have equivalent classifying
  toposes even when they are not syntactially related in any way. This
  observation is the starting point of Olivia Caramello's \emph{bridge
  technique}, a grand research program with applications in many different
  fields.

  \[ \xymatrix{
    & \Set[T] \simeq \Set[T'] \ar@{--}@/^1pc/[rd] \ar@{--}@/_1pc/[ld] \\
    T && T'
  } \]}
\end{frame}

\section{A glimpse of the toposophic landscape}

\begin{frame}[fragile]{A glimpse of the toposophic landscape}
  \hspace*{-1em}%
  \begin{tikzpicture}
    \node (objs-set0) at (0,0) {
      \only<1-2>{\muchstuffplaceholder}
      \only<3>{\muchstuff}
    };
    \node[scale=0.4] (objs-set1) at (-4.0,-2.5) {
      \only<1->{\fewstuff}
    };
    \node[scale=0.4] (objs-eff1) at (4.0,-2.5) {
      \only<2->{\fewstuff}
    };
    \node[scale=0.4] (objs-sh1)  at (0,-2.5) {
      \only<2->{\fewstuff}
    };

    \node (prop-set1) [below of=objs-set1, align=left] {
      \only<1->{%
        The usual laws \\
        of logic hold.
      }
    };

    \node (prop-eff1) [below of=objs-eff1, align=left] {
      \only<2->{%
        Every function \\
        is computable.
      }
    };

    \node (prop-sh1) [below of=objs-sh1, align=left, inner sep=0pt] {
      \only<2->{%
        The intermediate \\
        value theorem fails.
      }
    };

    \node (more-eff1) [below of=prop-eff1, visible on=<3->] {
      \threeblobs
    };
    \node (more-sh1)  [below of=prop-sh1, visible on=<3->] {
      \threeblobs
    };
    \node (more-set1) [below of=prop-set1, visible on=<3->] {
      \threeblobs
    };

    \begin{scope}[visible on=<1->]
      \drawbox{set1}{$\mathrm{Set}$}{(objs-set1) (prop-set1) (more-set1)}{}
    \end{scope}
    \begin{scope}[visible on=<2->]
      \drawbox{eff1}{Ef{}f}{(objs-eff1) (prop-eff1) (more-eff1)}{tape}
    \end{scope}
    \begin{scope}[visible on=<2->]
      \drawbox{sh1}{$\mathrm{Sh}\, X$}{(objs-sh1) (prop-sh1) (more-sh1)}{draw=none}
      \def\R{8pt}
      \begin{pgfonlayer}{background}
      \draw[decoration={bumps,segment length=8pt}, decorate, very thick, draw=mypurple, visible on=<2->]
        ($(sh1.south west) + (\R, 0)$) arc(270:180:\R) --
        ($(sh1.north west) + (0, -\R)$) arc(180:90:\R) --
        ($(sh1.north east) + (-\R, 0)$) arc(90:0:\R) --
        ($(sh1.south east) + (0, \R)$) arc(0:-90:\R) --
        cycle;
      \end{pgfonlayer}
    \end{scope}
    \begin{scope}[visible on=<3->]
      \drawbox{set0}{$\mathrm{Set}$}{(objs-set0) (set1) (eff1) (sh1)}{}
    \end{scope}
  \end{tikzpicture}

  \jnote{1}{The topos~$\Set$ is the standard topos. This topos is where
  ordinary mathematics takes place.}

  \jnote{2}{Besides~$\Set$, there is a proper class worth of further toposes.
  We'll get to know some of these better during the remaining course of the
  talk.

  For any topological space~$X$, there is a topos~$\Sh(X)$ of \emph{sheaves
  over~$X$}. They are useful in analysis for internalizing
  parameter-dependence. Apart from pathological cases like~$X$
  being discrete, the intermediate value theorem in the form
  \begin{indentblock}Let~$f : \RR \to \RR$ be a continuous function.
  Assume~$f(-1) < 0 < f(1)$. Then there is a number~$x \in \RR$ such
  that~$f(x) = 0$.\end{indentblock}
  fails in these toposes -- for very meaningful reasons discussed below.
  The following, classically equivalent, version does hold:
  \begin{indentblock}Let~$f : \RR \to \RR$ be a continuous function.
  Assume~$f(-1) < 0 < f(1)$. Then, for every~$\varepsilon > 0$, there is
  a number~$x \in \RR$ such that~$|f(x)| < \varepsilon$.\end{indentblock}

  The \emph{effective topos}~$\Eff$ is a computer scientist's dream come true:
  In it, any function~$\NN \to \NN$ is computable by a Turing machine. The
  effective topos and its close cousins can be used to study the differences
  between the many models of computation, particularly those differences which
  are only visible at higher types.}
\end{frame}

\begin{frame}{The internal universe of a topos}
  For any topos~$\E$ and any statement~$\varphi$, we define the meaning of
  \vspace*{-0.5em}
  \[
    \text{``$\E \models \varphi$''} \quad
    \text{(``$\varphi$ holds in the internal universe of~$\E$'')}
  \]

  \vspace*{-1.0em}
  using the \hil{Kripke--Joyal semantics}.

  \vspace*{-1em}
  \begin{columns}
    \def\insertblocktitle{}
    \begin{column}{0.3\textwidth}\usebeamertemplate{block begin}
      \centering
      $\Set \models \varphi$ \\
      ``$\varphi$ holds in the usual sense.''
    \usebeamertemplate{block end}\end{column}

    \begin{column}{0.3\textwidth}\usebeamertemplate{block begin}
      \centering
      $\Sh(X) \models \varphi$ \\
      ``$\varphi$ holds continuously.''
    \usebeamertemplate{block end}\end{column}

    \begin{column}{0.3\textwidth}\usebeamertemplate{block begin}
      \centering
      $\Eff \models \varphi$ \\
      ``$\varphi$ holds computably.''
    \usebeamertemplate{block end}\end{column}
  \end{columns}
  \medskip

  \pause
  Any topos supports \hil{mathematical reasoning}:

  \vspace*{-1.5em}
  \begin{hilblock}
    If~$\E \models \varphi$ and if~$\varphi$ entails~$\psi$
    \pointthis{<3>}{intuitionistically}{%
      no $\varphi \vee \neg\varphi$,\ \
      no $\neg\neg\varphi \Rightarrow \varphi$,\ \
      no axiom of choice},
    then~$\E \models \psi$.
  \end{hilblock}
  \bigskip
\end{frame}

\newcommand{\intex}[3]{#1 \quad #2\par #3\bigskip\medskip}

\begin{frame}{First steps in alternate universes}
  \begin{changemargin}{-1.1em}{0em}
    \fontsize{10pt}{12pt}\selectfont
    \begin{itemize}
      \item \intex{
        $\Eff \models \text{``Any number is prime or is not prime.''}$
      }{\textcolor{green!90}{\cmark}}{
        Meaning: There is a \hil{Turing machine} which determines of
        any given number whether it is prime or not.
      }

      \item \intex{
        $\Eff \models \text{``There are infinitely many prime numbers.''}$
      }{\textcolor{green!90}{\cmark}}{
        Meaning: There is a \hil{Turing machine} producing arbitrarily many
        primes.\\[0em]
      }

      \item \intex{
        $\Eff \models \text{``Any function~$\NN \to \NN$ is the zero function or not.''}$
      }{\textcolor{red!80}{\xmark}}{
        Meaning: There is a \hil{Turing machine} which, given a Turing
        machine computing a function~$f : \NN \to \NN$, determines whether~$f$
        is zero or not.
      }

      \item \intex{
        $\Eff \models \text{``Any function~$\NN \to \NN$ is computable.''}$
      }{\textcolor{green!90}{\cmark}}{}

      \item \intex{
        $\Sh(X) \models \text{``Any cont. function with opposite signs has
        a zero.''}$
      }{\textcolor{red!80}{\xmark}}{
        Meaning: Zeros can locally be picked \hil{continuously} in
        continuous families of continuous functions.
        \textcolor{red!80}{(\href{https://rawgit.com/iblech/internal-methods/master/images/zeros-in-families.mp4}{video} for counterexample)}
      }
    \end{itemize}
  \end{changemargin}
\end{frame}

\newcommand{\appl}[3]{\centering\includegraphics[height=3.5em]{#1}\par\hil{#2}\\\footnotesize#3}

\begin{frame}{Applications}
  \bigskip

  \small
  \begin{columns}
    \begin{column}{0.35\textwidth}
      \appl{analysis}{analysis}{
        internalizing parameter-dependence
      }
    \end{column}

    \begin{column}{0.39\textwidth}
      \appl{calabi-yau}{algebraic geometry}{
        \mbox{reducing geometry to algebra} \\
        \mbox{reducing relative to absolute} \\
        synthetic account
      }
    \end{column}

    \begin{column}{0.35\textwidth}
      \appl{klein-bottle}{differential geometry}{
        reflection principles \\
        synthetic account
      }
    \end{column}
  \end{columns}

  \bigskip
  \bigskip

  \begin{columns}
    \begin{column}{0.35\textwidth}
      \appl{torus-rainbow}{homotopy theory}{
        synthetic account
        computer-assisted proofs
        generalizations
      }
    \end{column}

    \begin{column}{0.39\textwidth}
      \appl{generic-freeness}{commutative algebra}{
        local-to-global principles \\
        reduction techniques \\
        constructive proofs
      }
    \end{column}

    \begin{column}{0.35\textwidth}
      \appl{bohr-topos}{further subjects}{
        synth.~computability th. \\
        synth.~measure theory \\
        Bohr topos for QM
      }
    \end{column}
  \end{columns}
\end{frame}

\section[Analysis]{Applications in analysis}

\begin{frame}{The topos of sheaves over a space}
  \small\vspace*{-0.8em}
  Let~$X$ be a topological space. We recursively define
  \[ U \models \varphi \quad \text{(``$\varphi$ holds on~$U$'')} \]
  for open subsets~$U \subseteq X$ and statements~$\varphi$.
  Write~``$\Sh(X) \models \varphi$'' to mean~$X \models \varphi$.
  Let~$\C(U)$ be the set of continuous functions~$U \to \RR$.

  \vspace*{-1.3em}
  \footnotesize
  \[ \renewcommand{\arraystretch}{1.15}\begin{array}{@{}l@{\ \ }c@{\ \ }l@{}}
  U \models \top &\Ll& \text{true} \\
  U \models \bot &\Ll& \hcancel{\text{false}}{0pt}{3pt}{0pt}{-2pt}\ U = \emptyset \\
  U \models s = t &\Ll& \text{$s(x) = t(x)$ for all $x \in U$} \\
  U \models \varphi \wedge \psi &\Ll&
  \text{$U \models \varphi$ and $U \models \psi$} \\
  U \models \varphi \vee \psi &\Ll&
  \hcancel{\text{$U \models \varphi$ or $U \models \psi$}}{0pt}{3pt}{0pt}{-2pt}\ \text{there exists an open covering $U = \bigcup_i U_i$} \\
  && \quad\quad\text{such that for all~$i$: $U_i \models \varphi$ or $U_i \models \psi$} \\
  U \models \varphi \Rightarrow \psi &\Ll&
  \text{for all open~$V \subseteq U$: }
  \text{$V \models \varphi$ implies $V \models \psi$} \\
  U \models \forall s \? \RR\_ \varphi(s) &\Ll&
  \text{for all open $V \subseteq U$ and functions~$s_0 \in \C(V)$: $V \models \varphi(s_0)$} \\
  U \models \exists s \? \RR\_ \varphi(s) &\Ll&
  \hcancel{\text{there exists $s_0 \in \C(U)$ such that $U \models \varphi(s_0)$}}{0pt}{3pt}{0pt}{-2pt} \\
  &&
  \text{there exists an open covering $U = \bigcup_i U_i$ such that for all~$i$:} \\
  && \quad\quad \text{there exists~$s_0 \in \C(U_i)$ such that $U_i \models \varphi(s_0)$}
  \end{array} \]
\end{frame}

\section[Differential geometry]{Applications in differential geometry}

\begin{frame}{Synthetic differential geometry}
  \vspace*{-1em}
  \begin{center}
    \includegraphics[width=0.5\textwidth]{microaffinity}
  \end{center}

  \vspace*{-2em}
  \begin{varblock}{10.8cm}{The axiom of microaffinity}
    \justifying
    Let~$\Delta = \{ \varepsilon \in \RR \,|\, \varepsilon^2 = 0 \}$.
    For any function $f : \Delta \to \RR$,
    there is a unique number~$a \in \RR$ such that
    $f(\varepsilon) = f(0) + a \varepsilon$
    for all~$\varepsilon \in \Delta$.
  \end{varblock}

  \vspace*{-1em}
  \small
  \begin{itemize}
    \item The \hil{derivative} of a function~$f : \RR \to \RR$ at~$x_0 \in \RR$
    is the unique number~$a \in \RR$ such that~$f(x_0 + \varepsilon) = f(x_0) +
    a \varepsilon$ for all~$\varepsilon \in \Delta$. \\[-1.2em]
    \item Manifolds are \hil{just sets}. \\[-1.2em]
    \item A \hil{tangent vector} to~$M$ is a map~$\Delta \to M$.
  \end{itemize}
\end{frame}

\section[Commutative algebra]{Applications to commutative algebra}

\begin{frame}{The little Zariski topos of a ring}
  \small\vspace*{-0.8em}
  Let~$A$ be a ring (commutative, with unit). We recursively define
  \[ D(f) \models \varphi \quad \text{(``$\varphi$ holds away from the zeros of~$f$'')} \]
  for~$f \in A$ and statements~$\varphi$. Write
  ``$\Spec(A) \models \varphi$'' to mean~$D(1) \models \varphi$.

  \vspace*{-1em}
  \[\footnotesize\renewcommand{\arraystretch}{1.25}\begin{array}{@{}l@{\ }c@{\ }l@{}}
    D(f) \models \top &\text{iff}& \text{true} \\
    D(f) \models \bot &\text{iff}& \text{$f$ is nilpotent} \\
    D(f) \models x = y &\text{iff}& x = y \in M[f^{-1}] \\
    D(f) \models \varphi \wedge \psi &\text{iff}&
      \text{$D(f) \models \varphi$ and $D(f) \models \psi$} \\
    D(f) \models \varphi \vee \psi &\text{iff}&
      \text{there exists a partition~$f^n = fg_1 + \cdots + fg_m$ with,} \\
    &&\quad\text{for each~$i$, $D(fg_i) \models \varphi$ or $D(fg_i) \models \psi$} \\
    D(f) \models \varphi \Rightarrow \psi &\text{iff}&
      \text{for all~$g \in A$, $D(fg) \models \varphi$ implies $D(fg) \models \psi$} \\
    D(f) \models \forall x\?M^\sim\_ \varphi(x) &\text{iff}&
      \text{for all~$g \in A$ and $x_0 \in M[(fg)^{-1}]$, $D(fg) \models \varphi(x_0)$} \\
    D(f) \models \exists x\?M^\sim\_ \varphi(x) &\text{iff}&
      \text{there exists a partition~$f^n = fg_1 + \cdots + fg_m$ with,} \\
    &&\quad\text{for each~$i$, $D(fg_i) \models \varphi(x_0)$ for some~$x_0 \in M[(fg_i)^{-1}]$}
  \end{array} \]
\end{frame}

\backupstart

\section[Algebraic geometry]{Applications in algebraic geometry}

\begin{frame}{Understanding algebraic geometry}
  \vspace*{-1.5em}\small
  \begin{varblock}{0.9\textwidth}{}
    \justifying
    Understand \hil{notions of algebraic geometry} over a scheme~$X$ as
    \hil{notions of algebra} internal to~$\Sh(X)$.
  \end{varblock}

  \only<1>{
    \centering
    \rotatebox{90}{\tiny\scalebox{0.8}{Illustration: Carina Willbold}}\hspace{-0.05cm}%
    \includegraphics[width=0.7\textwidth]{external-internal-small}
    \par
  }

  \pause

  \small\centering
  \scalebox{0.83}{\begin{tabular}{ll}
    \toprule
    externally & internally to $\Sh(X)$ \\
    \midrule
    sheaf of sets & set \\
    % sheaf of rings & ring \\
    sheaf of modules & module \\
    % sheaf of finite type & finitely generated module \\
    finite locally free sheaf & finite free module \\
    % coherent sheaf & coherent module \\
    tensor product of sheaves & tensor product of modules \\
    % sheaf of Kähler differentials & module of Kähler differentials \\
    sheaf of rational functions & total quotient ring of~$\O_X$ \\
    dimension of $X$ & Krull dimension of~$\O_X$ \\
    spectrum of a sheaf of~$\O_X$-algebras & ordinary spectrum [with a twist] \\
    % big Zariski topos of $X$ & big Zariski topos of the ring $\O_X$ [with a twist] \\
    higher direct image & sheaf cohomology \\
    \bottomrule
  \end{tabular}}

  \footnotesize

  \begin{columns}[c]
    \begin{column}{0.47\textwidth}
      \begin{exampleblock}{}
        \justifying
        Let $0 \to \F' \to \F \to \F'' \to 0$ be a short exact sequence
        of sheaves of~$\O_X$-modules. If~$\F'$ and~$\F''$ are of finite type,
        so is~$\F$.
      \end{exampleblock}
    \end{column}

    \begin{column}{0.1\textwidth}
      \vspace*{1.1em}
      \scalebox{3}{$\Leftarrow$}
    \end{column}

    \begin{column}{0.44\textwidth}
      \begin{exampleblock}{}
        \justifying
        Let~$0 \to M' \to M \to M'' \to 0$ be a short exact sequence of
        modules. If~$M'$ and~$M''$ are finitely generated, so is~$M$.
      \end{exampleblock}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Synthetic algebraic geometry}
  Usual approach to algebraic geometry: \hil{layer schemes above ordinary set theory}
  using either
  \begin{itemize}
    \item locally ringed spaces
    \small
    \begin{multline*}
      \text{set of prime ideals of~$\ZZ[X,Y,Z]/(X^n+Y^n-Z^n)$} + {} \\
      \text{Zariski topology} + \text{structure sheaf}
    \end{multline*}
    \normalsize
    \item or Grothendieck's functor-of-points account, where a scheme is a
    functor~$\Ring \to \Set$.
    \small\[ A \longmapsto \{ (x,y,z) \in A^3 \,|\, x^n+y^n-z^n=0 \} \]
  \end{itemize}
  \bigskip

  \hil{Synthetic approach:} model schemes \hil{directly as sets} in
  the internal universe of the \mbox{\hil{big Zariski topos}} of a base scheme.
  \small
  \[ \{ (x,y,z) \? (\affl)^3 \,|\, x^n+y^n-z^n=0 \} \]
\end{frame}

\backupend

\end{document}