### 2019-09-20 slides-herrsching2019.tex
Slides for Lecture I finished

\documentclass[12pt,utf8,notheorems,compress,t]{beamer}
\usepackage{etex}

%\usepackage{pgfpages}
%\setbeameroption{show notes on second screen}
%\setbeamertemplate{note page}[plain]
%\newcommand{\jnote}[2]{\only<#1>{\note{\setlength\parskip{\medskipamount}\justifying\footnotesize#2\par}}}
\newcommand{\jnote}[2]{}

% Workaround for the issue described at
% https://tex.stackexchange.com/questions/164406/beamer-using-href-in-notes.
\newcommand{\fixedhref}[2]{\makebox[0pt][l]{\hspace*{\paperwidth}\href{#1}{#2}}\href{#1}{#2}}

\usepackage[english]{babel}

\usepackage{graphbox}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{stmaryrd}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{multicol}
\usepackage{tabto}
\usepackage{xstring}
\usepackage{ifthen}
\usepackage[normalem]{ulem}
\usepackage[all]{xy}
\xyoption{rotate}
\usepackage{tikz}
\usetikzlibrary{calc,shapes,shapes.callouts,shapes.arrows,patterns,fit,backgrounds,decorations.pathmorphing,positioning}
\hypersetup{colorlinks=true}

\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}
\DeclareSymbolFont{extraup}{U}{zavm}{m}{n}
\DeclareMathSymbol{\varheart}{\mathalpha}{extraup}{86}

\graphicspath{{images/}}

\usepackage[protrusion=true,expansion=true]{microtype}

\setlength\parskip{\medskipamount}
\setlength\parindent{0pt}

\title{Generalized spaces for constructive algebra}
\author{Ingo Blechschmidt}
\date{September 2019}

\useinnertheme[shadow=true]{rounded}
\setbeamerfont{block title}{size={}}

\useinnertheme{rectangles}

\usecolortheme{orchid}
\usecolortheme{seahorse}
\definecolor{mypurple}{RGB}{150,0,255}
\setbeamercolor{structure}{fg=mypurple}
\definecolor{myred}{RGB}{150,0,0}
\setbeamercolor*{title}{bg=myred,fg=white}
\setbeamercolor*{titlelike}{bg=myred,fg=white}
\setbeamercolor{frame}{bg=black}

\usefonttheme{serif}
\usepackage[T1]{fontenc}
\usepackage{libertine}

% lifted from https://arxiv.org/abs/1506.08870
\DeclareFontFamily{U}{min}{}
\DeclareFontShape{U}{min}{m}{n}{<-> udmj30}{}
\newcommand\yon{\!\text{\usefont{U}{min}{m}{n}\symbol{'210}}\!}

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\renewcommand{\C}{\mathcal{C}}
\newcommand{\M}{\mathcal{M}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\renewcommand{\G}{\mathcal{G}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\GG}{\mathbb{G}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\ppp}{\mathfrak{p}}
\newcommand{\fff}{\mathfrak{f}}
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\defeqv}{\vcentcolon\equiv}
\newcommand{\Sh}{\mathrm{Sh}}
\newcommand{\GL}{\mathrm{GL}}
\newcommand{\Zar}{\mathrm{Zar}}
\newcommand{\op}{\mathrm{op}}
\newcommand{\Set}{\mathrm{Set}}
\newcommand{\Eff}{\mathrm{Ef{}f}}
\newcommand{\Sch}{\mathrm{Sch}}
\newcommand{\Aff}{\mathrm{Aff}}
\newcommand{\Ring}{\mathrm{Ring}}
\newcommand{\LocRing}{\mathrm{LocRing}}
\newcommand{\LRS}{\mathrm{LRS}}
\newcommand{\Hom}{\mathrm{Hom}}
\newcommand{\Spec}{\mathrm{Spec}}
\newcommand{\lra}{\longrightarrow}
\newcommand{\RelSpec}{\operatorname{Spec}}
\renewcommand{\_}{\mathpunct{.}}
\newcommand{\?}{\,{:}\,}
\newcommand{\speak}[1]{\ulcorner\text{\textnormal{#1}}\urcorner}
\newcommand{\ul}[1]{\underline{#1}}
\newcommand{\affl}{\ensuremath{{\ul{\ensuremath{\AA}}^1}}}
\newcommand{\Ll}{\text{iff}}
\newcommand{\inv}{inv.\@}
\newcommand{\seq}[1]{\mathrel{\vdash\!\!\!_{#1}}}
\newcommand{\hg}{\mathbin{:}}  % homogeneous coordinates
\newcommand{\brak}[1]{{\llbracket{#1}\rrbracket}}

\setbeamertemplate{blocks}[rounded][shadow=false]

\newenvironment{indentblock}{%
  \list{}{\leftmargin\leftmargin}%
  \item\relax
}{%
  \endlist
}

% Adapted from https://latex.org/forum/viewtopic.php?t=2251 (Stefan Kottwitz)
\newenvironment<>{hilblock}{
  \begin{center}
    \begin{minipage}{9.05cm}
      \setlength{\textwidth}{9.05cm}
      \begin{actionenv}#1
        \def\insertblocktitle{}
        \par
        \usebeamertemplate{block begin}}{
        \par
        \usebeamertemplate{block end}
      \end{actionenv}
    \end{minipage}
  \end{center}}

\newcommand{\bignumber}[1]{
  \renewcommand{\insertenumlabel}{#1}\scalebox{1.5}{\usebeamertemplate{enumerate item}}
}
\newcommand{\bigheart}{\includegraphics{heart}}

\newenvironment{changemargin}[2]{%
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \setlength{\leftmargin}{#1}%
    \setlength{\rightmargin}{#2}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\itemindent}{\parindent}%
    \setlength{\parsep}{\parskip}%
  }%
  \item[]}{\end{list}}

\tikzset{
  invisible/.style={opacity=0,text opacity=0},
  visible on/.style={alt={#1{}{invisible}}},
  alt/.code args={<#1>#2#3}{%
    \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}}}
}

\newcommand{\pointthis}[3]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=0,outer sep=0] (#2) {#2};
    \node[visible on=#1,overlay,rectangle callout,rounded corners,callout relative pointer={(0.3cm,0.5cm)},fill=blue!20] at ($(#2.north)+(-0.1cm,-1.1cm)$) {#3};
  }%
}

\tikzset{
  invisible/.style={opacity=0,text opacity=0},
  visible on/.style={alt={#1{}{invisible}}},
  alt/.code args={<#1>#2#3}{%
    \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}}}
}

\newcommand{\hcancel}[5]{%
  \tikz[baseline=(tocancel.base)]{
    \node[inner sep=0pt,outer sep=0pt] (tocancel) {#1};
    \draw[red!80, line width=0.4mm] ($(tocancel.south west)+(#2,#3)$) -- ($(tocancel.north east)+(#4,#5)$);
  }%
}

\newcommand{\explain}[7]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=2pt,outer sep=0,fill=#3,rounded corners] (label) {#1};
    \node[anchor=north,visible on=<#2>,overlay,rectangle callout,rounded corners,callout
    relative pointer={(0.0cm,0.5cm)+(0.0cm,#6)},fill=#3] at ($(label.south)+(0,-0.3cm)+(#4,#5)$) {#7};
  }%
}

\newcommand{\explainstub}[2]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=2pt,outer sep=0,fill=#2,rounded corners] (label) {#1};
  }%
}

\newcommand{\squiggly}[1]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=0,outer sep=0] (label) {#1};
    \draw[thick,color=red!80,decoration={snake,amplitude=0.5pt,segment
    length=3pt},decorate] ($(label.south west) + (0,-2pt)$) -- ($(label.south east) + (0,-2pt)$);
  }%
}

% Adapted from https://latex.org/forum/viewtopic.php?t=2251 (Stefan Kottwitz)
\newenvironment<>{varblock}[2]{\begin{varblockextra}{#1}{#2}{}}{\end{varblockextra}}
\newenvironment<>{varblockextra}[3]{
  \begin{center}
    \begin{minipage}{#1}
      \begin{actionenv}#4
        {\centering \hil{#2}\par}
	\def\insertblocktitle{}%\centering #2}
        \def\varblockextraend{#3}
	\usebeamertemplate{block begin}}{
        \par
        \usebeamertemplate{block end}
        \varblockextraend
      \end{actionenv}
    \end{minipage}
  \end{center}}

\setbeamertemplate{headline}{%
  \begin{beamercolorbox}[wd=\paperwidth,ht=2.25ex]{}%
    \insertsectionnavigationhorizontal{\paperwidth}{}{}%
  \end{beamercolorbox}%
  \vskip0pt%
}

\setbeamertemplate{frametitle}{%
  \vskip0.4em%
  \leavevmode%
  \begin{beamercolorbox}[dp=1ex,center]{}%
  %   \usebeamercolor[fg]{item}{\textbf{{\Large \insertframetitle}}}
    \begin{tikzpicture}
      \def\R{8pt}
      \node (title) {\hil{\large\,\!\insertframetitle}};
      \begin{pgfonlayer}{background}
        \draw[decorate, very thick, draw=mypurple]
          ($(title.south west) + (\R, 0)$) arc(270:180:\R) --
          ($(title.north west) + (0, -\R)$) arc(180:90:\R) --
          ($(title.north east) + (-\R, 0)$) arc(90:0:\R) --
          ($(title.south east) + (0, \R)$) arc(0:-90:\R) --
          cycle;
      \end{pgfonlayer}
    \end{tikzpicture}
  \end{beamercolorbox}%
  \vskip-0.6em%
}

\setbeamertemplate{navigation symbols}{}

\newcounter{framenumberpreappendix}
\newcommand{\backupstart}{
  \setcounter{framenumberpreappendix}{\value{framenumber}}
}
\newcommand{\backupend}{
  \addtocounter{framenumberpreappendix}{-\value{framenumber}}
  \addtocounter{framenumber}{\value{framenumberpreappendix}}
}

\newcommand{\insertframeextra}{}
\setbeamertemplate{footline}{%
  \begin{beamercolorbox}[wd=\paperwidth,ht=2.25ex,dp=1ex,right,rightskip=1mm,leftskip=1mm]{}%
    % \inserttitle
    \hfill
    \insertframenumber\insertframeextra\,/\,\inserttotalframenumber
  \end{beamercolorbox}%
  \vskip0pt%
}

\newcommand{\hil}[1]{{\usebeamercolor[fg]{item}{\textbf{#1}}}}
\newcommand{\bad}[1]{\textcolor{red!90}{\textnormal{#1}}}

\begin{document}

\addtocounter{framenumber}{-1}

\tikzstyle{topos} = [draw=mypurple, very thick, rectangle, rounded corners, inner sep=5pt, inner ysep=10pt]
\tikzstyle{title} = [fill=mypurple, text=white]

\input{images/primes.tex}
%\renewcommand{\sieve}[2]{SIEVE}
%\renewcommand{\fakesieve}[2]{SIEVE}

\newcommand{\drawbox}[4]{
  \node[topos, #4] [fit = #3] (#1) {};
  \node[title] at (#1.north) {#2};
}

\newcommand{\muchstuff}{
  \includegraphics[height=3em]{filmat}
  \scalebox{0.5}{\sieve{14}{2}}
}

\newcommand{\muchstuffplaceholder}{
  \includegraphics[height=3em]{filmat-placeholder}
  \scalebox{0.5}{\fakesieve{14}{2}}
}

\newcommand{\fewstuff}{
  \includegraphics[height=3em]{filmat}
  \scalebox{0.5}{\sieve{7}{2}}
}

\newcommand{\threeblobs}{
  \colorbox{mypurple}{\ \ }\quad
  \colorbox{mypurple}{\ \ }\quad
  \colorbox{mypurple}{\ \ }
}

%\setbeamertemplate{headline}{\mynav{gray}{gray}{gray}}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{4.95cm}\includegraphics[width=\paperwidth]{topos-horses}\end{minipage}}
\begin{frame}[c]
  \centering

  \begin{tikzpicture}
    \def\R{8pt}
    \node (title) {\hil{\,Generalized spaces for constructive algebra\,}};
    \begin{pgfonlayer}{background}
      \draw[decorate, very thick, draw=mypurple]
        ($(title.south west) + (\R, 0)$) arc(270:180:\R) --
        ($(title.north west) + (0, -\R)$) arc(180:90:\R) --
        ($(title.north east) + (-\R, 0)$) arc(90:0:\R) --
        ($(title.south east) + (0, \R)$) arc(0:-90:\R) --
        cycle;
    \end{pgfonlayer}
  \end{tikzpicture}
  \bigskip

  \begin{columns}
    \small
    \begin{column}{0.35\textwidth}
      \centering
      \bignumber{1}

      \phantom{x}\hil{Locales}
      \medskip

      \includegraphics[height=4.5em]{phantoms.png}
    \end{column}
    \hspace*{0.5em}
    \begin{column}{0.28\textwidth}
      \centering
      \bignumber{2}

      \phantom{x}\hil{Sheaf models}
      \medskip

      \phantom{x}\!\!\begin{tikzpicture}
        \node (overview) {
          \scalebox{0.35}{\sieve{4}{3}}
        };
        \def\R{8pt}
        \begin{pgfonlayer}{background}
        \draw[decoration={bumps,segment length=8pt}, decorate, very thick, draw=mypurple]
          ($(overview.south west) + (\R, 0)$) arc(270:180:\R) --
          ($(overview.north west) + (0, -\R)$) arc(180:90:\R) --
          ($(overview.north east) + (-\R, 0)$) arc(90:0:\R) --
          ($(overview.south east) + (0, \R)$) arc(0:-90:\R) --
          cycle;
        \end{pgfonlayer}
      \end{tikzpicture}
    \end{column}
    \begin{column}{0.4\textwidth}
      \centering
      \bignumber{3}

      \hil{Constructive algebra}
      \medskip

      \includegraphics[height=4.5em]{calabi-yau}
    \end{column}
  \end{columns}
  \bigskip

  \scriptsize
  \textit{-- an invitation --}
  \bigskip

  Autumn schoon on Proof and Computation \\
  September 20th to 26th, 2019 \\
  Herrsching
  \bigskip

  Ingo Blechschmidt \\
  Università di Verona
  \par

\end{frame}}

% \begin{document}

\section{Locales}

\begin{frame}
  \centering\bigskip
  \includegraphics[height=9.5em]{phantoms}

  \large Lecture I:
  \hil{Locales}

  \normalsize

  \begin{hilblock}Locales are a kind of space in which \hil{opens} instead of
  \hil{points} are fundamental.\end{hilblock}

\end{frame}

\begin{frame}{Locales in a nutshell}
  \justifying
  \textbf{Definition:} A \hil{topological space}~$X$ consists of a \hil{set of
  points} together with a set~$\O(X)$ of point sets which are deemed
  \hil{open} such that unions and finite intersections of open sets are open.
  \bigskip

  \centering\small
  \begin{tikzpicture}[sibling distance=8em, level distance=2.5em,
    every node/.style = {shape=rectangle, rounded corners,
      draw, align=center}, semithick, draw=mypurple]
    \node {euclidean spaces} [grow=up]
      child { node {metric spaces}
        child[dashed] { node [solid] {sober topological spaces}
          child[solid] { node {locales}
            child { node {toposes}
              child { node {$\infty$-toposes} } } }
          child[solid] { node {topological spaces} } } };
  \end{tikzpicture}
\end{frame}

\begin{frame}{Nontrivial spaces without points}
  The following locales don't have any points and are nontrivial:
  \begin{enumerate}
    \item the locale of surjections $\NN \twoheadrightarrow \RR$
    \item $\QQ \cap (\RR \setminus \QQ)$
    \item the pairwise intersections in the Banach--Tarski paradox
    \item Alex Simpson's locale of random binary sequences
  \end{enumerate}

  \begin{hilblock}
    Relinquishing points increases flexibility.
  \end{hilblock}

  \centering
  \includegraphics[height=5em]{banach-tarski}
\end{frame}

\begin{frame}{Issues of constructivity}
  \begin{enumerate}\justifying
    \item The unit interval~$[0,1]$ is \hil{compact}.
    \bigskip

    \item \hil{The fundamental theorem of Galois theory:}
    Let~$L | k$ be a Galois extension. Then there is a bijection between
    \[ \small\begin{array}{@{}rcl@{}}
      \begin{minipage}{0.25\textwidth}\centering the intermediate extensions~$L|E|k$\end{minipage}
      &\text{and}&
      \begin{minipage}{0.33\textwidth}\centering the closed subgroups
      of~$\mathrm{Aut}_k(L)$.\end{minipage} \\
      \\
      E &\longmapsto& \{ \sigma \in \mathrm{Aut}_k(L) \,|\, \sigma|_E = \operatorname{id} \} \\
      L^H &\longmapsfrom& H
    \end{array} \]
    \medskip

    \item \hil{Gelfand duality:} The categories of
    \[ \small\begin{array}{@{}rcl@{}}
      \begin{minipage}{0.38\textwidth}\centering compact Hausdorff spaces\end{minipage}
      &\text{and}&
      \left(\begin{minipage}{0.33\textwidth}\centering commutative $C^\star$-algebras with unit\end{minipage}\right)^{\mathrm{op}} \\[1.1em]
      X &\longmapsto& \{ f : X \to \CC \,|\, \text{$f$ continuous} \}
    \end{array} \]
    are equivalent.
  \end{enumerate}
\end{frame}

\begin{frame}{Localic basics}
  \justifying
  \textbf{Definition:} A \hil{topological space}~$X$ consists of a \hil{set of
  points} together with a set~$\O(X)$ of point sets which are deemed
  \hil{open} such that unions and finite intersections of open sets are open.

  \textbf{Observation:} The partial order~$\O(X)$ of open sets has
  \[ \small\begin{array}{@{}ccc@{}}
    \text{arbitrary joins}
    &\text{and}&
    \text{finite meets}, \\
    \bigvee && \wedge
  \end{array} \]
  and finite meets distribute over arbitrary joins:
  \[ U \wedge \bigvee_i V_i = \bigvee_i (U \wedge V_i). \]
  \vspace*{-1.2em}
  \pause

  \textbf{Definition:} A \hil{frame} is a partial order with arbitrary joins
  and finite meets such that the distributive law holds.

  \textbf{Definition:} A \hil{locale}~$X$ is given by a \hil{frame~$\O(X)$ of opens}.
\end{frame}

\begin{frame}{Locales presented by theories}
  \justifying
  \textbf{Definition:} A \hil{geometric signature} consists of
  \begin{enumerate}
    \item a set of sorts: $X$, $Y$, $Z$, \ldots
    \item a set of function symbols: $f : X \times Y \to Z$, \ldots
    \item a set of relation symbols: $R \hookrightarrow X \times Y \times Z$, \ldots
  \end{enumerate}
  % A geometric signature is \hil{propositional} iff its set
  % of sorts is empty.
  A \hil{geometric theory}~$\TT$ over a geometric
  signature~$\Sigma$ consists of a set of axioms (geometric sequents
  over~$\Sigma$).

  \only<1>{
    \textbf{Examples:}
    \begin{enumerate}
      \item the theory of rings
      \item the theory of surjections~$\NN \twoheadrightarrow \RR$
      \item the theory of Dedekind cuts
    \end{enumerate}
  }
  \pause

  \textbf{Definition:} A \hil{set-based model}~$M$ of a theory~$\TT$ consists of
  \begin{enumerate}
    \item a set~$\brak{X}$ for each sort~$X$,
    \item a function~$\brak{f} : \brak{X_1} \times \cdots \times \brak{X_n} \to
    \brak{Y}$ for \\
    each function symbol~$f : X_1 \times \cdots \times X_n \to
    Y$, and
    \item a relation~$\brak{R} \subseteq \brak{X_1} \times \cdots \times
    \brak{X_n}$ for \\
    each relation symbol~$R \hookrightarrow X_1 \times \cdots \times X_n$
  \end{enumerate}
  such that~$M$ validates the axioms of~$\TT$.
\end{frame}

\end{document}

    \begin{column}[t]{0.4\textwidth}
      \justifying
      \textbf{Contrast with:} A \hil{topological space} consists of a \hil{set of
      points}~$X$ together with a set~$\O(X)$ of point sets which are deemed
      \hil{open} such that \ldots
    \end{column}

% Grothendieck quote for the title picture:
%
% Ces “nuages probabilistes”, remplaçant les rassurantes particules matérielles
% d’antan, me rappellent étrangement les élusifs “voisinages ouverts” qui
% peuplent les topos, tels des fantômes évanescents, pour entourer des “points”
% imaginaires, ...
%
% Diese "Wahrscheinlichkeitswolken", welche die beruhigenden materiellen
% Partikel von früher ersetzen, erinnern mich irgendwie an die flüchtigen
% "offenen Umgebungen" der Topoi -- wie dahinschwindende Phantome, um
% die fiktiven "Punkte" zu umgeben, ...
%
% These “probability clouds”, replacing the reassuring material particles of
% before, remind me strangely of the elusive “open neighborhoods” that populate
% the topoi, like evanescent phantoms, to surround the imaginary “points”.