### 2022-09-18 slides-brixen2022b.tex
Slides for informal talk almost finished

\documentclass[12pt,utf8,notheorems,compress,t]{beamer}
\usepackage{etex}

\usepackage{pgfpages}
\usepackage[export]{adjustbox}

% Workaround for the issue described at
% https://tex.stackexchange.com/questions/164406/beamer-using-href-in-notes.
\newcommand{\fixedhref}[2]{\makebox[0pt][l]{\hspace*{\paperwidth}\href{#1}{#2}}\href{#1}{#2}}

\usepackage[english]{babel}

\usepackage[normalem]{ulem}
\usepackage{graphbox}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{stmaryrd}
\usepackage{amssymb}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{multicol}
\usepackage{tabto}
\usepackage{xstring}
\usepackage{proof}
\usepackage{ifthen}
\usepackage[normalem]{ulem}
\usepackage[all]{xy}
\xyoption{rotate}
\usepackage{tikz}
\usetikzlibrary{calc,shapes,shapes.callouts,shapes.arrows,patterns,fit,backgrounds,decorations.pathmorphing,positioning}
\hypersetup{colorlinks=true}

\newcommand*\circled[1]{\tikz[baseline=(char.base)]{%
  \node[shape=circle,draw,inner sep=1pt] (char) {#1};}}

\DeclareFontFamily{U}{bbm}{}
\DeclareFontShape{U}{bbm}{m}{n}
   {  <5> <6> <7> <8> <9> <10> <12> gen * bbm
      <10.95> bbm10%
      <14.4>  bbm12%
      <17.28><20.74><24.88> bbm17}{}
\DeclareFontShape{U}{bbm}{m}{sl}
   {  <5> <6> <7> bbmsl8%
      <8> <9> <10> <12> gen * bbmsl
      <10.95> bbmsl10%
      <14.4> <17.28> <20.74> <24.88> bbmsl12}{}
\DeclareFontShape{U}{bbm}{bx}{n}
   {  <5> <6> <7> <8> <9> <10> <12> gen * bbmbx
      <10.95> bbmbx10%
      <14.4> <17.28> <20.74> <24.88> bbmbx12}{}
\DeclareFontShape{U}{bbm}{bx}{sl}
   {  <5> <6> <7> <8> <9> <10> <10.95> <12> <14.4> <17.28>%
      <20.74> <24.88> bbmbxsl10}{}
\DeclareFontShape{U}{bbm}{b}{n}
   {  <5> <6> <7> <8> <9> <10> <10.95> <12> <14.4> <17.28>%
      <20.74> <24.88> bbmb10}{}
\DeclareMathAlphabet{\mathbbm}{U}{bbm}{m}{n}
\SetMathAlphabet\mathbbm{bold}{U}{bbm}{bx}{n}

\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}
\DeclareSymbolFont{extraup}{U}{zavm}{m}{n}
\DeclareMathSymbol{\varheart}{\mathalpha}{extraup}{86}

\graphicspath{{images/}}

\usepackage[protrusion=true,expansion=true]{microtype}

\setlength\parskip{\medskipamount}
\setlength\parindent{0pt}

\title{Connecting inductive definitions, generic models and a modal multiverse
for algebra and combinatorics}

\author{Ingo Blechschmidt}
\date{September ??th, 2022}

%\useinnertheme[shadow=true]
\setbeamerfont{block title}{size={}}

\useinnertheme{rectangles}

\usecolortheme{orchid}
\usecolortheme{seahorse}
\definecolor{mypurple}{RGB}{253,73,34}
\setbeamercolor{structure}{fg=mypurple}
\definecolor{myred}{RGB}{150,0,0}
\setbeamercolor*{title}{bg=myred,fg=white}
\setbeamercolor*{titlelike}{bg=myred,fg=white}
\setbeamercolor{frame}{bg=black}

\usefonttheme{serif}
\usepackage[T1]{fontenc}
\usepackage{libertine}

% lifted from https://arxiv.org/abs/1506.08870
\DeclareFontFamily{U}{min}{}
\DeclareFontShape{U}{min}{m}{n}{<-> udmj30}{}
\newcommand\yon{\!\text{\usefont{U}{min}{m}{n}\symbol{'210}}\!}

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\M}{\mathcal{M}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\BB}{\mathbb{B}}
\newcommand{\pp}{\mathbbm{p}}
\newcommand{\MM}{\mathbb{M}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\FF}{\mathbb{F}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\GG}{\mathbb{G}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\ppp}{\mathfrak{p}}
\newcommand{\fff}{\mathfrak{f}}
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\defeqv}{\vcentcolon\equiv}
\newcommand{\Sh}{\mathrm{Sh}}
\newcommand{\GL}{\mathrm{GL}}
\newcommand{\Zar}{\mathrm{Zar}}
\newcommand{\op}{\mathrm{op}}
\newcommand{\Set}{\mathrm{Set}}
\newcommand{\Eff}{\mathrm{Ef{}f}}
\newcommand{\Sch}{\mathrm{Sch}}
\newcommand{\Aff}{\mathrm{Aff}}
\newcommand{\Ring}{\mathrm{Ring}}
\newcommand{\LocRing}{\mathrm{LocRing}}
\newcommand{\LRS}{\mathrm{LRS}}
\newcommand{\Hom}{\mathrm{Hom}}
\newcommand{\Spec}{\mathrm{Spec}}
\newcommand{\lra}{\longrightarrow}
\newcommand{\RelSpec}{\operatorname{Spec}}
\renewcommand{\_}{\mathpunct{.}}
\newcommand{\?}{\,{:}\,}
\newcommand{\speak}[1]{\ulcorner\text{\textnormal{#1}}\urcorner}
\newcommand{\ul}[1]{\underline{#1}}
\newcommand{\affl}{\ensuremath{{\ul{\ensuremath{\AA}}^1}}}
\newcommand{\Ll}{\text{iff}}
\newcommand{\inv}{inv.\@}
\newcommand{\seq}[1]{\mathrel{\vdash\!\!\!_{#1}}}
\newcommand{\hg}{\mathbin{:}}  % homogeneous coordinates

\setbeamertemplate{blocks}[rounded][shadow=false]

\newenvironment{indentblock}{%
  \list{}{\leftmargin\leftmargin}%
  \item\relax
}{%
  \endlist
}

% Adapted from https://latex.org/forum/viewtopic.php?t=2251 (Stefan Kottwitz)
\newenvironment<>{hilblock}{
  \begin{center}
    \begin{minipage}{9.05cm}
      \setlength{\textwidth}{9.05cm}
      \begin{actionenv}#1
        \def\insertblocktitle{}
        \par
        \usebeamertemplate{block begin}}{
        \par
        \usebeamertemplate{block end}
      \end{actionenv}
    \end{minipage}
  \end{center}}

\newenvironment{changemargin}[2]{%
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \setlength{\leftmargin}{#1}%
    \setlength{\rightmargin}{#2}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\itemindent}{\parindent}%
    \setlength{\parsep}{\parskip}%
  }%
  \item[]}{\end{list}}

\tikzset{
  invisible/.style={opacity=0,text opacity=0},
  visible on/.style={alt={#1{}{invisible}}},
  alt/.code args={<#1>#2#3}{%
    \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}}}
}

\newcommand{\pointthis}[3]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=0,outer sep=0] (#2) {#2};
    \node[visible on=#1,overlay,rectangle callout,rounded corners,callout relative pointer={(0.3cm,0.5cm)},fill=blue!20] at ($(#2.north)+(-0.1cm,-1.1cm)$) {#3};
  }%
}

\tikzset{
  invisible/.style={opacity=0,text opacity=0},
  visible on/.style={alt={#1{}{invisible}}},
  alt/.code args={<#1>#2#3}{%
    \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}}}
}

\newcommand{\hcancel}[5]{%
  \tikz[baseline=(tocancel.base)]{
    \node[inner sep=0pt,outer sep=0pt] (tocancel) {#1};
    \draw[red!80, line width=0.4mm] ($(tocancel.south west)+(#2,#3)$) -- ($(tocancel.north east)+(#4,#5)$);
  }%
}

\newcommand{\explain}[7]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=2pt,outer sep=0,fill=#3,rounded corners] (label) {#1};
    \node[anchor=north,visible on=<#2>,overlay,rectangle callout,rounded corners,callout
    relative pointer={(0.0cm,0.5cm)+(0.0cm,#6)},fill=#3] at ($(label.south)+(0,-0.3cm)+(#4,#5)$) {#7};
  }%
}

\newcommand{\explainstub}[2]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=2pt,outer sep=0,fill=#2,rounded corners] (label) {#1};
  }%
}

\newcommand{\squiggly}[1]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=0,outer sep=0] (label) {#1};
    \draw[thick,color=red!80,decoration={snake,amplitude=0.5pt,segment
    length=3pt},decorate] ($(label.south west) + (0,-2pt)$) -- ($(label.south east) + (0,-2pt)$);
  }%
}

% Adapted from https://latex.org/forum/viewtopic.php?t=2251 (Stefan Kottwitz)
\newenvironment<>{varblock}[2]{\begin{varblockextra}{#1}{#2}{}}{\end{varblockextra}}
\newenvironment<>{varblockextra}[3]{
  \begin{center}
    \begin{minipage}{#1}
      \begin{actionenv}#4
        {\centering \hil{#2}\par}
	\def\insertblocktitle{}%\centering #2}
        \def\varblockextraend{#3}
	\usebeamertemplate{block begin}}{
        \par
        \usebeamertemplate{block end}
        \varblockextraend
      \end{actionenv}
    \end{minipage}
  \end{center}}

\setbeamertemplate{headline}{}

\setbeamertemplate{frametitle}{%
  \vskip0.5em%
  \leavevmode%
  \begin{beamercolorbox}[dp=1ex,center]{}%
    \begin{tikzpicture}
      \def\R{8pt}
      \node (title) {\hil{{\large\,\!\insertframetitle}}};
      \begin{pgfonlayer}{background}
        \draw[decorate, very thick, draw=mypurple!30]
          ($(title.south west) + (\R, 0)$) arc(270:180:\R) --
          ($(title.north west) + (0, -\R)$) arc(180:90:\R) --
          ($(title.north east) + (-\R, 0)$) arc(90:0:\R) --
          ($(title.south east) + (0, \R)$) arc(0:-90:\R) --
          cycle;
      \end{pgfonlayer}
    \end{tikzpicture}
  \end{beamercolorbox}%
  \vskip-0.2em%
}

\setbeamertemplate{navigation symbols}{}

\newcounter{framenumberpreappendix}
\newcommand{\backupstart}{
  \setcounter{framenumberpreappendix}{\value{framenumber}}
}
\newcommand{\backupend}{
  \addtocounter{framenumberpreappendix}{-\value{framenumber}}
  \addtocounter{framenumber}{\value{framenumberpreappendix}}
}

\newcommand{\insertframeextra}{}
\setbeamertemplate{footline}{%
  \begin{beamercolorbox}[wd=\paperwidth,ht=2.25ex,dp=1ex,right,rightskip=1mm,leftskip=1mm]{}%
    % \inserttitle
    \hfill
    \insertframenumber\insertframeextra\,/\,\inserttotalframenumber
  \end{beamercolorbox}%
  \vskip0pt%
}

\newcommand{\hil}[1]{{\usebeamercolor[fg]{item}{\textbf{#1}}}}
\newcommand{\bad}[1]{\textcolor{red!90}{\textnormal{#1}}}

\newcommand{\bignumber}[1]{%
  \renewcommand{\insertenumlabel}{#1}\scalebox{1.2}{\!\usebeamertemplate{enumerate item}\!}
}
\newcommand{\normalnumber}[1]{%
  {\renewcommand{\insertenumlabel}{#1}\!\usebeamertemplate{enumerate item}\!}
}
\newcommand{\bigheart}{\includegraphics{heart}}

\newcommand{\subhead}[1]{{\centering\textcolor{gray}{\hrulefill}\quad\textnormal{#1}\quad\textcolor{gray}{\hrulefill}\par}}

\newcommand{\badbox}[1]{\colorbox{red!30}{#1}}

% taken from JDH "The modal logic of arithmetic potentialism and the universal algorithm"
\DeclareMathOperator{\possible}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,rotate=45,line width=.1ex]{\draw (-1,-1) rectangle (1,1);}}}
\DeclareMathOperator{\necessary}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,line width=.1ex]{\draw (-1,-1) rectangle (1,1);}}}
\DeclareMathOperator{\xpossible}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,rotate=45,line width=.1ex]{\draw (-1,-1) rectangle (1,1); \draw[very thin] (-.6,-.6) rectangle (.6,.6);}}}
\DeclareMathOperator{\xnecessary}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,line width=.1ex]{\draw (-1,-1) rectangle (1,1); \draw[very thin] (-.6,-.6) rectangle (.6,.6);}}}

\begin{document}

\addtocounter{framenumber}{-1}

%\setbeamertemplate{headline}{\mynav{gray}{gray}{gray}}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{1.59cm}\includegraphics[width=\paperwidth]{forest-light}\end{minipage}}
\begin{frame}[c]
  \centering

  \bigskip
  %\includegraphics[height=0.32\textwidth]{olivia-lattices}
  \bigskip
  \bigskip
  \bigskip
  \bigskip

  \begin{tikzpicture}
    \def\R{8pt}
    \node (title) {\vbox{\vspace*{-0.0em}Connecting \hil{inductive definitions},
    \hil{generic models} and \\ a \hil{modal multiverse} for algebra and combinatorics
    \\[0.8em]}};
    \begin{pgfonlayer}{background}
      \draw[decorate, very thick, draw=mypurple]
        ($(title.south west) + (\R, 0)$) arc(270:180:\R) --
        ($(title.north west) + (0, -\R)$) arc(180:90:\R) --
        ($(title.north east) + (-\R, 0)$) arc(90:0:\R) --
        ($(title.south east) + (0, \R)$) arc(0:-90:\R) --
        cycle;
    \end{pgfonlayer}
  \end{tikzpicture}

  \scriptsize
  \textit{-- an invitation --}
  \bigskip
  \bigskip
  \bigskip

  REDCOM: \\
  \emph{Reducing complexity in algebra, logic, combinatorics}
  \ \\
  September 19th, 2022
  \bigskip

  Ingo Blechschmidt \\
  University of Augsburg
  \par
\end{frame}}

\definecolor{mypurple}{RGB}{150,0,255}
\setbeamercolor{structure}{fg=mypurple}

% \begin{document}

\begin{frame}{Infinite data}
  \begin{block}{}
    \justifying
    \textbf{Def.} A set~$X$ together with a binary relation~$R$ is
    \hil{almost-full$_\infty$} iff every \bad{\uwave{infinite sequence}}~$\alpha : \NN \to X$ is
    \hil{good} in that there exist numbers~$i < j$ such that~$\alpha(i)
    \mathrel{R} \alpha(j)$.
  \end{block}
  \vspace*{-0.6em}
  {\mbox{\small\emph{Examples.} $(\NN,{\leq})$,\ \ $X \times Y$ [Dickson],\ \
  $X^*$ [Higman],\ \ $\mathrm{Tree}(X)$ [Kruskal]} \\
  \bad{Only classically.}
  \pause
  Constructive reformulation:
  \par}
%  {\small\emph{Examples.}\vspace*{-0.8em}
%  \begin{itemize}
%    \item $(\NN,{\leq})$\vspace*{-0.6em}
%    \item Cartesian products of almost-full sets (by \emph{Dickson's Lemma}) \vspace*{-0.6em}
%    \item Finite lists over almost-full sets (by \emph{Higman's Lemma}) \vspace*{-0.6em}
%    \item Finite trees over almost-full sets (by \emph{Kruskal's Tree Theorem})
%  \end{itemize}}

  \begin{block}{}
    \justifying
    \textbf{Def.} For a predicate~$P$ on finite lists over~$X$, inductively
    define:
    \vspace*{-0.6em}
    \[
      \infer{P \,|\, \sigma}{P(\sigma)}
      \qquad
      \infer{P \,|\, \sigma}{\forall x \in X\_ P \,|\, (\sigma :: x)}
    \]
  \end{block}
  \pause

  \begin{block}{}
    \textbf{Def.} A set~$X$ together with a binary relation~$R$ is
    \hil{almost-full\textsubscript{ind}} iff~$\mathsf{Good} \,|\, []$, where
    $\mathsf{Good}(\sigma) \equiv (\exists i < j\_ \sigma[i] \mathrel{R}
    \sigma[j])$.
  \end{block}
  \vspace*{-0.6em}
  \begin{enumerate}
    \item Constructively, almost-full\textsubscript{ind} $\Rightarrow$
    almost-full$_\infty$.
    \item With \badbox{\textsc{lem}+\textsc{dc}}, almost-full\textsubscript{ind} $\Leftarrow$
    almost-full$_\infty$.
    %\pause
    %\item How much stronger is almost-full\textsubscript{ind}?
    %many-valued sequences, \ldots
  \end{enumerate}
\end{frame}

\begin{frame}{Constructive red flags}
  \emph{Well-foundedness:} A transitive relation~$<$ on a set~$X$ is \ldots
  \begin{itemize}
    \item \ \\[-1.2em]\mbox{\hil{well-founded$_\infty$} iff
    there is no \bad{\uwave{infinite chain}} $x_0 > x_1 > \cdots$.}
    \item \hil{well-founded'$_\infty$} iff
    there is no \bad{\uwave{bad set}} (inhabited and for every member there is
    a smaller member).

    \item<4-> \hil{well-founded\textsubscript{ind}} iff for every~$x \in X$,
    $\mathsf{Acc}(x)$: %, where~$\mathsf{Acc}$ is inductively defined by:
    \[
      \infer{\mathsf{Acc}(x)}{\forall y < x\_ \mathsf{Acc}(y)}
    \]
    \vspace*{-1.3em}
  \end{itemize}

  \pause
  \emph{Krull's Lemma:} An element~$x$ of a ring~$A$ is nilpotent if \ldots
  \begin{itemize}
    \item it is contained in
    every \bad{\uwave{prime ideal}}.
    \item<5-> the theory of prime ideals of~$A$ proves~``$x \in \ppp$''.
  \end{itemize}
  \medskip

  \pause
  \justifying
  \emph{Dependent choice:}
  Let~$R$ be a binary relation on a set~$X$
  such that~$\forall a \in X\_ \exists b \in X\_ a \mathrel{R} b$. Let~$x_0
  \in X$. Then there is an \bad{\uwave{infinite chain}} $x_0 \mathrel{R}
  x_1 \mathrel{R} x_2 \mathrel{R} \cdots$.
\end{frame}

%\begin{frame}{Inductive definitions}
%  \begin{block}{}
%    \justifying
%    \textbf{Def.} A transitive relation~$<$ on a set~$X$ is \hil{well-founded} iff
%    for all~$x \in X$, $\mathsf{Acc}(x)$, where the accessibility predicate is
%    inductively generated by:
%  \end{block}
%  \vspace*{-0.6em}
%  {\small\emph{Examples.} $(\NN,{<})$\par}
%
%  \textbf{Prop.} If~$<$ is well-founded, there are no infinite descending
%  chains~$x_0 > x_1 > \ldots$.
%
%  \emph{Proof.} By induction on~$x_0$.
%\end{frame}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{3.59cm}\includegraphics[width=\paperwidth]{staircase}\end{minipage}}
\begin{frame}{The set-theoretic multiverse}
  \begin{block}{}
    \justifying
    \textbf{Def.} A \hil{model of set theory} is a (perhaps class-sized) structure $(M,{\in})$
    satisfying axioms such as those of~\textsc{zfc}.
  \end{block}
  \vspace*{-0.6em}
  {\small\emph{Examples.}\vspace*{-0.8em}
  \begin{itemize}
    \item $V$, the class of all sets \vspace*{-0.6em}
    \item $L$, Gödel's constructible universe \vspace*{-0.6em}
    \item $V[G]$, a forcing extension containing a generic filter of~$G$ \vspace*{-0.6em}
    \item Henkin/term models from consistency of (extensions of)~\textsc{zfc}
  \end{itemize}}
  \pause

  {\centering\emph{Embracing all models of set theory:}\par}
  \begin{block}{}
    \justifying
    \textbf{Def.} $\possible\varphi$ iff~$\varphi$ holds in \hil{some extension} of
    the current universe. \\
    \phantom{\textbf{Def.}} $\necessary\varphi$ iff~$\varphi$ holds in \hil{all extensions}
    of the current universe.
  \end{block}
  \begin{itemize}
    \item $\necessary(\possible\textsc{CH} \wedge \possible\neg\text{CH})$,
    the continuum hypothesis is a \hil{switch}
    \item \ \\[-1.2em]\mbox{$\necessary\possible\necessary(\text{$X$ is countable})$,
    existence of an enumeration is a \hil{button}}
  \end{itemize}
\end{frame}}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{4.95cm}\includegraphics[width=\paperwidth]{topos-horses-lighter}\end{minipage}}
\begin{frame}{A topos-theoretic multiverse}
  \begin{block}{}
    \justifying
    \textbf{Def.} A statement~$\varphi$ holds \ldots
    \begin{enumerate}
      \item\small \hil{everywhere} ($\necessary\varphi$) iff
      it holds in every (Grothendieck) topos (over the current base topos).
      \item \hil{somewhere} ($\possible\varphi$) iff
      it holds in some positive topos.
      \item \hil{subsequently} ($\xpossible\varphi$) iff
      it holds in some positive ouvert topos.
    \end{enumerate}
  \end{block}
  \pause

  \newcommand{\seqbad}{\only<2>{infinite descending sequence}\only<3->{bad set}}
  \begin{itemize}
  \small
    \item\justifying
    \only<2-4>{A relation is well-founded\textsubscript{ind} \\
    iff \emph{everywhere} there are no \seqbad s \\
    iff the \emph{generic \seqbad} validates~$\bot$ \\
    iff the \emph{theory} of \seqbad s proves~$\bot$.}
    \only<5->{A relation is well-founded\textsubscript{ind} iff
    \emph{nowhere} there are \seqbad s.}
    \pause
    \pause

    \item \only<1-4>{A ring element~$x$ is nilpotent \\
    iff \emph{everywhere} it is contained in all prime ideals \\
    iff it is contained in the \emph{generic prime ideal} \\
    iff the \emph{theory} of prime ideals proves~$x \not\in \ppp$.}
    \only<5->{A ring element is nilpotent iff it is contained in all prime
    ideals \emph{everywhere}.}

    \pause
    \pause
    \item Given a set~$X$, \emph{subsequently} there is a surjection~$\NN
    \twoheadrightarrow X$. \\
    \hil{NB:} $(\possible\varphi) \Rightarrow \varphi$, if~$\varphi$ is a
    geometric implication. \\
    \phantom{\hil{NB:}} $(\xpossible\varphi) \Rightarrow \varphi$, if~$\varphi$
    is first-order.

    \pause
    \item Given~$(X,R,x_0)$ as in~\textsc{dc}, \emph{subsequently} there is an
    infinite chain.

    \pause
    \item Somewhere, the law of excluded middle holds.
  \end{itemize}
\end{frame}}

\end{document}

DC
Krull
Countability
Well
Well-founded

Choice:
erst abzählbar machen,
dann DC
nein, funktioniert nicht ganz
aber für diskrete Quelle habe surjektive (nicht offene?) Theorie/Topos

∀x ∈ X. ∃y ∈ Y. φ(x,y).
(Toposwechsel für Abzählbarkeit von X)
∀n ∈ ℕ. ∃y ∈ Y. φ(πn,y).
Yn := { y ∈ Y | φ(πn,y) }
(~) auf M = (coprod_n Y_n): (n,y) ~ (n',y') gdw. n' = n+1
Dann habe: M bewohnt (durch (0,...)) und ∀p ∈ M. ∃q ∈ M. p ~ q.
(Toposwechsel für DC für (~))
Habe f : ℕ → M mit f(0) = (0,...) und f(n) ~ f(n+1) für alle n.
Also ist π₂.f eine Auswahlfunktion, genauer: ∀n. φ(πn, π₂(f(n))).

Zorn für abzählbare Mengen?
C_0 = {}
C_{n+1} = C_n cup { x_n | C_n cup {x_n} Kette }
C = bigcup_n C_n
z = sup(C)
Dann ist z maximal?
Gelte z ≤ e.
Dann ist C cup { e } Kette.
Wenn e = x_n0, so insbesondere C_n0 cup { e } Kette.
Also e ∈ C_{n+1} ⊆ C, also e ≤ z.
Also e = z.
Problem: Kommt eher selten vor; Supremum wird oft nicht zur Menge gehören.
Beispiel: Supremum von Menge endlich erzeugter Ideale ist nicht endlich erzeugt.

Attribution:
Blass(?)
Alexander
Henry