### 2022-09-27 slides-fischbachau2022.tex
Final touches

\documentclass[11pt,utf8,notheorems,compress,t]{beamer}
\usepackage{etex}

\usepackage{pgfpages}
\usepackage[export]{adjustbox}

% Workaround for the issue described at
% https://tex.stackexchange.com/questions/164406/beamer-using-href-in-notes.
\newcommand{\fixedhref}[2]{\makebox[0pt][l]{\hspace*{\paperwidth}\href{#1}{#2}}\href{#1}{#2}}

\usepackage[english]{babel}

\usepackage{graphbox}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{stmaryrd,wasysym}
\usepackage{bussproofs}
\usepackage{xspace}
\usepackage{amssymb}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{multicol}
\usepackage{tabto}
\usepackage{xstring}
\usepackage{ifthen}
\usepackage[normalem]{ulem}
\usepackage[all]{xy}
\xyoption{rotate}
\usepackage{tikz}
\usetikzlibrary{calc,shapes,shapes.callouts,shapes.arrows,patterns,fit,backgrounds,decorations.pathmorphing,positioning}
\hypersetup{colorlinks=true}

\newcommand*\circled[1]{\tikz[baseline=(char.base)]{%
  \node[shape=circle,draw,inner sep=1pt] (char) {#1};}}

\DeclareFontFamily{U}{bbm}{}
\DeclareFontShape{U}{bbm}{m}{n}
   {  <5> <6> <7> <8> <9> <10> <12> gen * bbm
      <10.95> bbm10%
      <14.4>  bbm12%
      <17.28><20.74><24.88> bbm17}{}
\DeclareFontShape{U}{bbm}{m}{sl}
   {  <5> <6> <7> bbmsl8%
      <8> <9> <10> <12> gen * bbmsl
      <10.95> bbmsl10%
      <14.4> <17.28> <20.74> <24.88> bbmsl12}{}
\DeclareFontShape{U}{bbm}{bx}{n}
   {  <5> <6> <7> <8> <9> <10> <12> gen * bbmbx
      <10.95> bbmbx10%
      <14.4> <17.28> <20.74> <24.88> bbmbx12}{}
\DeclareFontShape{U}{bbm}{bx}{sl}
   {  <5> <6> <7> <8> <9> <10> <10.95> <12> <14.4> <17.28>%
      <20.74> <24.88> bbmbxsl10}{}
\DeclareFontShape{U}{bbm}{b}{n}
   {  <5> <6> <7> <8> <9> <10> <10.95> <12> <14.4> <17.28>%
      <20.74> <24.88> bbmb10}{}
\DeclareMathAlphabet{\mathbbm}{U}{bbm}{m}{n}
\SetMathAlphabet\mathbbm{bold}{U}{bbm}{bx}{n}

\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}
\DeclareSymbolFont{extraup}{U}{zavm}{m}{n}
\DeclareMathSymbol{\varheart}{\mathalpha}{extraup}{86}

\graphicspath{{images/}}

\usepackage[protrusion=true,expansion=true]{microtype}

\setlength\parskip{\medskipamount}
\setlength\parindent{0pt}

\title{Revisiting divisible, injective and flabby abelian groups from a
constructive point of view}
\author{Ingo Blechschmidt}
\date{September 21th, 2022}

%\useinnertheme[shadow=true]
\setbeamerfont{block title}{size={}}

\useinnertheme{rectangles}

\usecolortheme{orchid}
\usecolortheme{seahorse}
\definecolor{mypurple}{RGB}{150,0,255}
\setbeamercolor{structure}{fg=mypurple}
\definecolor{myred}{RGB}{150,0,0}
\setbeamercolor*{title}{bg=myred,fg=white}
\setbeamercolor*{titlelike}{bg=myred,fg=white}
\setbeamercolor{frame}{bg=black}

\usefonttheme{serif}
\usepackage[T1]{fontenc}
\usepackage{libertine}

% lifted from https://arxiv.org/abs/1506.08870
\DeclareFontFamily{U}{min}{}
\DeclareFontShape{U}{min}{m}{n}{<-> udmj30}{}
\newcommand\yon{\!\text{\usefont{U}{min}{m}{n}\symbol{'210}}\!}

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\M}{\mathcal{M}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\BB}{\mathbb{B}}
\newcommand{\pp}{\mathbbm{p}}
\newcommand{\MM}{\mathbb{M}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\FF}{\mathbb{F}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\GG}{\mathbb{G}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\ppp}{\mathfrak{p}}
\newcommand{\fff}{\mathfrak{f}}
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\defeqv}{\vcentcolon\equiv}
\newcommand{\Sh}{\mathrm{Sh}}
\newcommand{\GL}{\mathrm{GL}}
\newcommand{\Zar}{\mathrm{Zar}}
\newcommand{\op}{\mathrm{op}}
\newcommand{\Set}{\mathrm{Set}}
\newcommand{\Eff}{\mathrm{Ef{}f}}
\newcommand{\Sch}{\mathrm{Sch}}
\newcommand{\Aff}{\mathrm{Aff}}
\newcommand{\Ring}{\mathrm{Ring}}
\newcommand{\LocRing}{\mathrm{LocRing}}
\newcommand{\LRS}{\mathrm{LRS}}
\newcommand{\Hom}{\mathrm{Hom}}
\newcommand{\Spec}{\mathrm{Spec}}
\newcommand{\lra}{\longrightarrow}
\newcommand{\RelSpec}{\operatorname{Spec}}
\renewcommand{\_}{\mathpunct{.}\,}
\newcommand{\?}{\,{:}\,}
\newcommand{\speak}[1]{\ulcorner\text{\textnormal{#1}}\urcorner}
\newcommand{\ul}[1]{\underline{#1}}
\newcommand{\affl}{\ensuremath{{\ul{\ensuremath{\AA}}^1}}}
\newcommand{\Ll}{\text{iff}}
\newcommand{\inv}{inv.\@}
\newcommand{\seq}[1]{\mathrel{\vdash\!\!\!_{#1}}}
\newcommand{\hg}{\mathbin{:}}  % homogeneous coordinates

\setbeamertemplate{blocks}[rounded][shadow=false]

\newenvironment{indentblock}{%
  \list{}{\leftmargin\leftmargin}%
  \item\relax
}{%
  \endlist
}

% Adapted from https://latex.org/forum/viewtopic.php?t=2251 (Stefan Kottwitz)
\newenvironment<>{hilblock}{
  \begin{center}
    \begin{minipage}{9.05cm}
      \setlength{\textwidth}{9.05cm}
      \begin{actionenv}#1
        \def\insertblocktitle{}
        \par
        \usebeamertemplate{block begin}}{
        \par
        \usebeamertemplate{block end}
      \end{actionenv}
    \end{minipage}
  \end{center}}

\newenvironment{changemargin}[2]{%
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \setlength{\leftmargin}{#1}%
    \setlength{\rightmargin}{#2}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\itemindent}{\parindent}%
    \setlength{\parsep}{\parskip}%
  }%
  \item[]}{\end{list}}

\tikzset{
  invisible/.style={opacity=0,text opacity=0},
  visible on/.style={alt={#1{}{invisible}}},
  alt/.code args={<#1>#2#3}{%
    \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}}}
}

\newcommand{\pointthis}[3]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=0,outer sep=0] (#2) {#2};
    \node[visible on=#1,overlay,rectangle callout,rounded corners,callout relative pointer={(0.3cm,0.5cm)},fill=blue!20] at ($(#2.north)+(-0.1cm,-1.1cm)$) {#3};
  }%
}

\tikzset{
  invisible/.style={opacity=0,text opacity=0},
  visible on/.style={alt={#1{}{invisible}}},
  alt/.code args={<#1>#2#3}{%
    \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}}}
}

\newcommand{\hcancel}[5]{%
  \tikz[baseline=(tocancel.base)]{
    \node[inner sep=0pt,outer sep=0pt] (tocancel) {#1};
    \draw[red!80, line width=0.4mm] ($(tocancel.south west)+(#2,#3)$) -- ($(tocancel.north east)+(#4,#5)$);
  }%
}

\newcommand{\explain}[7]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=2pt,outer sep=0,fill=#3,rounded corners] (label) {#1};
    \node[anchor=north,visible on=<#2>,overlay,rectangle callout,rounded corners,callout
    relative pointer={(0.0cm,0.5cm)+(0.0cm,#6)},fill=#3] at ($(label.south)+(0,-0.3cm)+(#4,#5)$) {#7};
  }%
}

\newcommand{\explainstub}[2]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=2pt,outer sep=0,fill=#2,rounded corners] (label) {#1};
  }%
}

\newcommand{\squiggly}[1]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=0,outer sep=0] (label) {#1};
    \draw[thick,color=red!80,decoration={snake,amplitude=0.5pt,segment
    length=3pt},decorate] ($(label.south west) + (0,-2pt)$) -- ($(label.south east) + (0,-2pt)$);
  }%
}

% Adapted from https://latex.org/forum/viewtopic.php?t=2251 (Stefan Kottwitz)
\newenvironment<>{varblock}[2]{\begin{varblockextra}{#1}{#2}{}}{\end{varblockextra}}
\newenvironment<>{varblockextra}[3]{
  \begin{center}
    \begin{minipage}{#1}
      \begin{actionenv}#4
        {\centering \hil{#2}\par}
	\def\insertblocktitle{}%\centering #2}
        \def\varblockextraend{#3}
	\usebeamertemplate{block begin}}{
        \par
        \usebeamertemplate{block end}
        \varblockextraend
      \end{actionenv}
    \end{minipage}
  \end{center}}

\setbeamertemplate{headline}{}

\setbeamertemplate{frametitle}{%
  \vskip0.5em%
  \leavevmode%
  \begin{beamercolorbox}[dp=1ex,center]{}%
    \begin{tikzpicture}
      \def\R{8pt}
      \node (title) {\hil{{\large\,\!\insertframetitle}}};
      \begin{pgfonlayer}{background}
        \draw[decorate, very thick, draw=mypurple!30]
          ($(title.south west) + (\R, 0)$) arc(270:180:\R) --
          ($(title.north west) + (0, -\R)$) arc(180:90:\R) --
          ($(title.north east) + (-\R, 0)$) arc(90:0:\R) --
          ($(title.south east) + (0, \R)$) arc(0:-90:\R) --
          cycle;
      \end{pgfonlayer}
    \end{tikzpicture}
  \end{beamercolorbox}%
  \vskip-0.2em%
}

\setbeamertemplate{navigation symbols}{}

\newcounter{framenumberpreappendix}
\newcommand{\backupstart}{
  \setcounter{framenumberpreappendix}{\value{framenumber}}
}
\newcommand{\backupend}{
  \addtocounter{framenumberpreappendix}{-\value{framenumber}}
  \addtocounter{framenumber}{\value{framenumberpreappendix}}
}

\newcommand{\insertframeextra}{}
\setbeamertemplate{footline}{%
  \begin{beamercolorbox}[wd=\paperwidth,ht=2.25ex,dp=1ex,right,rightskip=1mm,leftskip=1mm]{}%
    % \inserttitle
    \hfill
    \insertframenumber\insertframeextra\,/\,\inserttotalframenumber
  \end{beamercolorbox}%
  \vskip0pt%
}

\newcommand{\hil}[1]{{\usebeamercolor[fg]{item}{\textbf{#1}}}}
\newcommand{\bad}[1]{\textcolor{red!90}{\textnormal{#1}}}
\newcommand{\good}[1]{\textcolor{mypurple}{\textnormal{#1}}}

\newcommand{\bignumber}[1]{%
  \renewcommand{\insertenumlabel}{#1}\scalebox{1.2}{\!\usebeamertemplate{enumerate item}\!}
}
\newcommand{\normalnumber}[1]{%
  {\renewcommand{\insertenumlabel}{#1}\!\usebeamertemplate{enumerate item}\!}
}
\newcommand{\bigheart}{\includegraphics{heart}}

\newcommand{\subhead}[1]{{\centering\textcolor{gray}{\hrulefill}\quad\textnormal{#1}\quad\textcolor{gray}{\hrulefill}\par}}

\newcommand{\badbox}[1]{\colorbox{red!30}{#1}}
\newcommand{\infobox}[1]{\colorbox{yellow!70}{\color{black}#1}}
\newcommand{\grayline}{\textcolor{gray}{\hspace*{-3em}\hrulefill\hspace*{-3em}}}
\newcommand{\notnot}{\emph{not~not}\xspace}

\begin{document}

\addtocounter{framenumber}{-1}

%\setbeamertemplate{headline}{\mynav{gray}{gray}{gray}}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{1.59cm}\includegraphics[width=\paperwidth]{forest-light}\end{minipage}}
\definecolor{mypurple}{RGB}{253,73,34}
\begin{frame}[c]
  \centering

  \bigskip
  \bigskip
  \bigskip
  \bigskip

  \scriptsize
  \textit{-- an invitation --}

  \setbeamercolor{block body}{bg=black!100}
  \begin{block}{}
    \centering\normalsize\color{white}
    \hil{Extraction} of \hil{programs} from \hil{proofs}
  \end{block}

  \bigskip
  \bigskip
  \bigskip
  \bigskip
  \bigskip
  \bigskip
  \bigskip

  Autumn school on \\
  \emph{Proof and Computation} \\
  in Fischbachau \\
  \ \\
  September 26th to October 1st, 2022

  Ingo Blechschmidt \\
  University of Augsburg
  \par
\end{frame}}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{5.95cm}\includegraphics[width=\paperwidth]{fr1-lighter}\end{minipage}}
\begin{frame}{}
  \medskip
  \textbf{Thm.}
  For every number~$n \in \NN$, there is a prime larger than~$n$.

  {\emph{Proof.} Any prime factor of~$n! + 1$ will do.\par}
  \medskip
  \pause

  {\centering\emph{``Every constructive theorem has a computable witness.''}
  \pause
  \[
    \begin{array}{c@{\qquad}c@{\qquad}c}
      \mathrm{HA} \vdash \varphi &\Longrightarrow&
      \exists e\_ e \Vdash \varphi \\
      \text{\small constructive proof} &\longmapsto&
      \text{\small realizer}
    \end{array}
  \]}
  \pause
  \vspace*{-2.2em}

  \begin{columns}
    \begin{column}{0.45\textwidth}
      \begin{itemize}
        \item Integrated developments \\ \emph{SAT checking, \ldots}
        \item Computability theory \\ \emph{induction $\widehat{=}$ recursion, \ldots}
      \end{itemize}
    \end{column}

    \hspace*{-1em}
    \begin{column}{0.65\textwidth}
      \begin{itemize}
        \item Metatheory of constructive systems \\ \emph{provability results, \ldots}
        \item Philosophy of proof and computation \\ \emph{realizability in the real world, \ldots}
      \end{itemize}
    \end{column}
  \end{columns}

  \grayline
  \pause

  \justifying
  \textbf{Thm.} Every infinite sequence~$\alpha : \NN \to \NN$ is \emph{good}
  in that there are numbers~$i < j$ such that~$\alpha(i) \leq \alpha(j)$.

  {\emph{Proof.} By~\badbox{\textsc{lem}}, there is a minimal value~$\alpha(i)$.
  Set~$j \defeq i + 1$.\par}

  {\centering\emph{``Every theorem has a computable* witness.''} \\ \scriptsize
  * with monadic side effects\par}
\end{frame}}

% \includegraphics[width=3cm]{lovelace-babbage}
% \includegraphics[width=3cm]{first-program}
\addtocounter{framenumber}{-1}
{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\includegraphics[height=\paperheight,valign=t]{ada-lovelace}\includegraphics[width=8cm,valign=t]{first-program}\end{minipage}}
\begin{frame}[plain]
  \vspace*{6cm}\hspace*{6cm}\begin{minipage}{4.7cm}
    \huge\hil{Ada Lovelace},

    \Large
    the world's first \\
    computer programmer
    \medskip

    * 1815 \ \ † 1852
  \end{minipage}
\end{frame}}

\section{Realizability theory}

\begin{frame}
  \centering\bigskip
  \includegraphics[height=9.5em]{turing-machine}
  \medskip

  \large Lecture I: \\
  \hil{Realizability theory}

  \normalsize

  \emph{for extracting programs from constructive proofs}
\end{frame}

\begin{frame}{Heyting arithmetic}
  The \hil{language of arithmetic} has
  \begin{itemize}
    \item as its single sort: $N$
    \item as function symbols: $0$, $S$, $+$, $\cdot$
    \item as its single relation symbol: $=$
  \end{itemize}

  \hil{Heyting arithmetic} has as axioms (the universal closure of)
  \begin{align*}
    &&& \neg(0 = Sx) \\
    &&& S(x) = S(y) \Rightarrow x = y \\
    x + 0 &= x &&&
    x \cdot 0 &= 0 \\
    x + S(y) &= S(x+y) &&&
    x \cdot S(y) &= (x \cdot y) + x
  \end{align*}
  together with the \hil{induction scheme} (one axiom for each
  formula~$\varphi$)
  \[
    \varphi(0) \wedge \bigl(\forall x\?N\_ \varphi(x) \Rightarrow
    \varphi(S(x))\bigr) \quad\Longrightarrow\quad \forall x\?N\_ \varphi(x)
  \]
  and the rules of \hil{sequence calculus}.
\end{frame}

\begin{frame}{Sequence calculus}
  \begin{center}
    \vspace{-0.5em}
    \phantom{a}\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}\UnaryInfC{$\varphi \seq{\vec x} \varphi$}\DisplayProof\hfill
    \AxiomC{$\varphi \seq{\vec x} \psi$}\UnaryInfC{$\varphi[\vec s/\vec x]
    \seq{\vec y} \psi[\vec s/\vec x]$}\DisplayProof\hfill
    \AxiomC{$\varphi \seq{\vec x} \psi$}\AxiomC{$\psi \seq{\vec x}
    \chi$}\BinaryInfC{$\varphi \seq{\vec x} \chi$}\DisplayProof
    \phantom{a}\hfill
    \bigskip\medskip

    \phantom{a}\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}\UnaryInfC{$\varphi \seq{\vec x} \top$}\DisplayProof\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}\UnaryInfC{$\varphi \wedge \psi \seq{\vec x} \varphi$}\DisplayProof\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}\UnaryInfC{$\varphi \wedge \psi \seq{\vec x} \psi$}\DisplayProof\hfill
    \AxiomC{$\varphi \seq{\vec x} \psi$}\AxiomC{$\varphi \seq{\vec x} \chi$}\BinaryInfC{$\varphi \seq{\vec x} \psi \wedge \chi$}\DisplayProof
    \phantom{a}\hfill
    \bigskip\medskip

    \phantom{a}\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}\UnaryInfC{$\bot \seq{\vec x} \varphi$}\DisplayProof\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}\UnaryInfC{$\varphi \seq{\vec x} \varphi \vee \psi$}\DisplayProof\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}\UnaryInfC{$\psi \seq{\vec x} \varphi \vee \psi$}\DisplayProof\hfill
    \AxiomC{$\varphi \seq{\vec x} \chi$}\AxiomC{$\psi \seq{\vec x} \chi$}\BinaryInfC{$\varphi \vee \psi \seq{\vec x} \chi$}\DisplayProof
    \phantom{a}\hfill
    \bigskip\medskip

    \phantom{a}\hfill
    \Axiom$\varphi \wedge \psi\ \fCenter\seq{\vec x} \chi$
    \doubleLine
    \UnaryInf$\varphi\ \fCenter\seq{\vec x} \psi \Rightarrow \chi$
    \DisplayProof
    \phantom{a}\hfill
    \bigskip\medskip

    \phantom{a}\hfill
    \Axiom$\varphi\ \fCenter\seq{\vec x, y} \psi$
    \doubleLine
    \UnaryInf$\exists y\?Y\_\! \varphi\ \fCenter\seq{\vec x} \psi$
    \DisplayProof
    {\tiny ($y$ not occurring in~$\psi$)}
    \hfill
    \Axiom$\varphi\ \fCenter\seq{\vec x, y} \psi$
    \doubleLine
    \UnaryInf$\varphi\ \fCenter\seq{\vec x\phantom{, y}} \forall y\?Y\_\! \psi$
    \DisplayProof
    {\tiny ($y$ not occurring in~$\varphi$)}
    \hfill\phantom{a}

    \phantom{a}\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}
    \UnaryInfC{$\top \seq{x} x = x$}
    \DisplayProof
    \hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}
    \UnaryInfC{$(\vec x = \vec y) \wedge \varphi \seq{\vec z} \varphi[\vec y/\vec x]$}
    \DisplayProof
    \hfill\phantom{a} \\[0.5em]
    (``$\vec x = \vec y\,$'' is short for~``$x_1 = y_1 \wedge \cdots \wedge x_n =
    y_n$''.)
  \end{center}
\end{frame}

\newcommand{\realizes}{\Vdash}

\begin{frame}{Number realizability}
  \vspace*{-1em}
  \begin{changemargin}{-1.5em}{-0.5em}
  \begin{tabbing}
    $e \models (\forall f\?\NN^\NN\_ \varphi(n))$ \= \kill
    $e \realizes s = t$ \> iff $s = t$. \\
    $e \realizes \top$ \> iff true. \\
    $e \realizes \bot$ \> iff false. \\
    $e \realizes (\varphi \wedge \psi)$ \> iff~$\pi_1 \cdot e \downarrow$
    and~$\pi_2 \cdot e \downarrow$ and $\pi_1 \cdot e \realizes \varphi$
    and~$\pi_2 \cdot e \realizes \psi$. \\
    $e \realizes (\varphi \vee \psi)$ \> iff~$\pi_1 \cdot e \downarrow$
    and~$\pi_2 \cdot e \downarrow$ and \\ \> \qquad if~$\pi_1 \cdot e = 0$
    then~$\pi_2 \cdot e \realizes
    \varphi$, and \\ \> \qquad if~$\pi_1 \cdot e \neq 0$ then~$\pi_2 \cdot e \realizes \psi$. \\
    $e \realizes (\varphi \Rightarrow \psi)$ \> iff for every~$r \in \NN$
    such that~$r \realizes \varphi$, $e \cdot r \downarrow$ and~$e \cdot r \realizes \psi$. \\
    $e \realizes (\forall n\?N\_ \varphi(n))$ \> iff for every~$n_0
    \in \NN$, $e \cdot n_0 \downarrow$ and~$e \cdot n_0 \realizes \varphi(n_0)$. \\
    $e \realizes (\exists n\?N\_ \varphi(n))$ \> iff~$\pi_1 \cdot e
    \downarrow$ and~$\pi_2 \cdot e \downarrow$
    and~$\pi_2 \cdot e \realizes \varphi(\pi_1 \cdot e)$. \\
    $e \realizes (\forall f\?N^N\_ \varphi(f))$ \> iff for every~$f_0
    : \NN \to \NN$ and every~$r_0 \in \NN$ such that \\ \> \qquad $f_0$ is computed by the~$r_0$-th
    machine, \\ \> \qquad
    $e \cdot r_0 \downarrow$ and~$e \cdot r_0 \realizes \varphi(f_0)$. \\
    $e \realizes (\exists f\?N^N\_ \varphi(f))$ \> iff~$\pi_1 \cdot e \downarrow$
    and~$\pi_2 \cdot e \downarrow$ and
    the $(\pi_1 \cdot e)$-th machine \\ \> \qquad computes a function~$f_0 : \NN \to \NN$
    and $\pi_2 \cdot e \realizes \varphi(f_0)$.
  \end{tabbing}\end{changemargin}

  \mbox{\textbf{Thm.} If~$\text{HA} \vdash \varphi$, then there is a
  number~$e \in \NN$ such that~\only<1>{$e \realizes
  \varphi$}\only<2>{$\text{HA} \vdash (\underline{e} \realizes \varphi)$}.}
\end{frame}

\newcommand{\expl}[2]{
  \justifying
  ``$\realizes \!\text{\normalnumber{#1}}$'' amounts to: #2
}

\newcommand{\qswitch}[3]{\only<1-#1>{\includegraphics[height=0.7em]{question-mark}}\only<#2->{#3}}
\newcommand{\ccmark}{\good{\cmark}}
\newcommand{\cxmark}{\bad{\xmark}}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{0.00cm}\hfill\mbox{\includegraphics[width=0.2\textwidth]{turing-machine}\;\;}\end{minipage}}
\begin{frame}{Exploring the realizability model}
  \medskip
  \medskip
  \medskip
  \begin{tabular}{@{\!\!\!\!\!\!}l@{\,}llp{1.8cm}}
    \toprule
    & statement & classical? & realizable? \\
    \midrule
    \normalnumber{1} & Every number is prime or not prime. & \ccmark{}
    (trivially) & \ccmark \\
    \normalnumber{2} & After every number there is a prime. & \ccmark & \ccmark \\
    \normalnumber{3} & Every map $\NN \to \NN$ has a zero or not. & \ccmark{} (trivially) & \cxmark \\
    \normalnumber{4} & Every map $\NN \to \NN$ is computable. & \cxmark &
    \qswitch{4}{5}{\ccmark}\only<1-4>{\,} \visible<5->{(trivially)} \\
    \normalnumber{5} & Every map $\RR \to \RR$ is continuous. & \cxmark &
    \qswitch{5}{6}{\ccmark} \\
    \normalnumber{6} & Markov's principle holds. & \ccmark{} (trivially) &
    \qswitch{6}{7}{\ccmark{} (if MP)} \\
    \normalnumber{7} & Countable choice holds. & \ccmark &
    \qswitch{7}{8}{\ccmark{} (always!)} \\
    \normalnumber{8} & Heyting arithmetic is categorical. & \cxmark &
    \qswitch{8}{9}{\ccmark{} (if MP)} \\
    \bottomrule
  \end{tabular}
  \medskip

  \only<1>{\ \\\ \\\ \\\ }
  \only<2>{\expl{1}{There is a machine which determines of any given
  number whether it is prime or not. \\\ }}
  \only<3>{\expl{2}{There is a machine which, given a number~$n$, computes a
  prime larger than~$n$. \\\ \\\ }}
  \only<4>{\expl{3}{There is a machine which, given a machine
  computing a map~$f : \NN \to \NN$, determines whether~$f$ has a
  zero or not. \\\ }}
  \only<5>{\expl{4}{There is a machine which, given a machine
  computing a map~$f : \NN \to \NN$, outputs a machine
  computing~$f$. \\\ }}
  \only<6>{\ \\\ \\\ \\\ }
  \only<7>{\expl{6}{There is a machine which, given a machine
  computing a map~$f : \NN \to \NN$ and given the promise that it is \notnot
  the case that~$f$ has a zero, determines a zero of~$f$.}}
  \only<8->{\expl{7}{\justifying There is a machine which, given a machine
  computing for every~$x \in \NN$ some~$y \in A$ together with a realizer
  of~$\varphi(x,y)$, outputs a machine computing a suitable choice
  function~$\NN \to A$.}}
\end{frame}}

\begin{frame}{Metatheory of Heyting arithmetic}
  \begin{enumerate}
    \item \hil{Unprovability results:}
    \medskip

    There are instances of~\badbox{\textsc{lem}} which HA does not prove,
    such as ``every Turing machine terminates or does not terminate''.
    \bigskip

    \item \hil{Disjunction property:}
    \medskip

    If HA proves~$\varphi \vee \psi$, then HA
    proves~$\varphi$ or HA proves~$\psi$.
    \bigskip

    \item \hil{Existence property:}
    \medskip

    If HA proves~$\exists n\?N\_ \varphi(n)$, then there is a number~$n_0 \in
    \NN$ such that HA proves~$\varphi(\underline{n_0})$.
    \bigskip

    \item \hil{Growth rate:}
    \medskip\justifying

    If HA proves~$\forall x\?N\_ \exists y\?N\_ \varphi(x,y)$,
    then there exists a \hil{higher primitive recursive} function $f_0 : \NN
    \to \NN$ such that for all~$x_0 \in \NN$, HA proves
    $\varphi(\underline{x_0}, \underline{f_0(x_0)})$.
  \end{enumerate}
\end{frame}

\begin{frame}{Range of machine models}
  \begin{enumerate}
    \item \hil{Turing machines}

    \visible<2->{``Every map $\NN \to \NN$ is computable'' is realized by
    \textsf{cat}.}
    \medskip

    \item \hil{Untyped lambda calculus}

    \visible<3->{``Every map $\NN \to \NN$ is computable'' is \emph{not}
    realized.}
    \medskip

    \item \hil{Infinite-time Turing machines}

    \visible<4->{\mbox{``Every map $\NN \to \NN$ has a zero or not'' is realized by
    infinite search.}\\[0em]}
    \medskip

    \item \hil{Gödel's System T}

    \visible<5->{Markov's principle is not realized.}
    \medskip

    \item \hil{Machines in the real world} -- philosophical

    \justifying
    \visible<6->{``Every map $\RR \to \RR$ is continuous'' is realized if,
    in the physical world, only finitely many computational
    steps can be carried out in finite time and if it is possible to form
    tamper-free private communication channels.}
  \end{enumerate}
\end{frame}

\section{Proof transformations}

\addtocounter{framenumber}{-1}
\begin{frame}[plain]\frametitle{A classical logic fairy tale}
  \vspace*{-0.5em}
  \begin{changemargin}{-1em}{-1em}
    \setlength{\columnsep}{15pt}
    \begin{multicols}{2}
      \justifying
      \scriptsize

      \textbf{Narrator.} Once upon a time, in a kingdom far, far away, the
      queen of the land and of all Möbius strips called for her royal
      philosopher.

      \textbf{Queen.} Philosopher! I ask you to carry out the following order.
      Get me the Philosopher's Stone, or alternatively find out how one could
      produce arbitrary amounts of gold with it!

      \textbf{Philosopher.} But my queen! I haven't studied anything useful!
      How could I fulfill this order?

      \textbf{Queen.} That is not my concern. I'll see you again tomorrow.
      Should you not accomplish the task, I will take your head off.

      \textbf{Narrator.} After a long and wakeful night the philosopher was
      called to the queen again.

      \textbf{Queen.} Tell me! What do you have to report?

      \textbf{Philosopher.} It was not easy and I needed to follow lots of
      obscure references, but finally I actually found out how to use the
      Philosopher's Stone to produce arbitrary amounts of gold. But only I can
      conduct this procedure, your royal highness.

      \textbf{Queen.}
      Alright. So be it.

      \textbf{Narrator.} And so years passed by, during which the philosopher
      imagined herself to be safe. The queen searched for the stone on her own,
      but as long as she hadn't found it, the philosopher didn't need to worry.
      Yet one day the impossible happened: The queen has found the stone! And
      promptly called for her philosopher.

      \textbf{Queen.} Philosopher, look! I have found the Philosopher's Stone!
      Now live up to your promise! \emph{[She hands over the stone.]}

      \textbf{Philosopher.} Thank you. \emph{[She inspects the stone.]} This is
      indeed the Philosopher's Stone. Many years ago you asked me to either
      acquire the Philosopher's Stone or find out how to produce arbitrary
      amounts of gold using it. Now it's my pleasure to present to you the
      Philosopher's Stone. \emph{[She returns the stone.]}
    \end{multicols}
  \end{changemargin}
\end{frame}

\begin{frame}
  \centering\bigskip
  \includegraphics[height=9.5em]{philosophers-stone}
  \medskip

  \large Lecture II: \\
  \hil{Proof transformations}

  \normalsize

  \emph{for extracting constructive proofs from classical proofs}
\end{frame}

\begin{frame}{A case study in double negation}
  \vspace*{-0.8em}
  \mbox{How to extract \good{constructive proofs} from
  \bad{classical proofs} as the following?}

  \begin{enumerate}
    \item \textbf{Thm.} Every infinite sequence~$\alpha : \NN \to \NN$ is \emph{good}
    in that there are numbers~$i < j$ such that~$\alpha(i) \leq \alpha(j)$.
    \medskip

    {\emph{Proof.} By~\badbox{\textsc{lem}}, there is a minimal value~$\alpha(i)$.
    Set~$j \defeq i + 1$.\par}
    \medskip
    \pause

    \item \textbf{Thm.} Every infinite binary sequence contains repeated terms.
    \medskip

    {\emph{Proof.} By the \bad{infinite box principle},
    infinitely many terms are zeros or are ones. In either case the claim
    follows.\par}
    \pause
  \end{enumerate}
  \vspace*{-0.7em}

  \grayline

  \textbf{Thm.} Let~$X$ be an inhabited set of natural numbers.
  \begin{enumerate}
    \item Assuming~\badbox{\textsc{lem}}, the set~$X$ contains a minimal element.
    \item<5-> If~$X$ is \hil{detachable}, then~$X$ contains a minimal element.
    \item<7-> It is \hil{not~not} the case that~$X$ contains a minimal element.
  \end{enumerate}

  \justifying
  \only<1-3>{\ \\\ \\\ \\\ }
  \only<4-5>{\emph{Proof of~\normalnumber{1}.} There is some~$n \in X$.
  By~\badbox{\textsc{lem}}, either~$\exists k \in X\_ k < n$ \\ or not. In the
  first case, we continue by induction. Else~$n$ is minimal.\\\ \par}
  \only<6-7>{\mbox{\emph{Proof of~\normalnumber{2}.} There is some~$n \in X$.
  By~\good{assumption}, either~$\exists k \in X\_ k < n$} or not. In the
  first case, we continue by induction. Else~$n$ is minimal.\\\ \par}
  \only<8>{\emph{Proof of~\normalnumber{3}.} There is some~$n \in X$. Assume that~$X$ does not
  contain a minimum. Then it is not the case that~$\exists k \in X\_ k <
  n$, as else~$\bot$ by induction. Hence~$n$ is minimal. This is a
  contradiction.\par}
\end{frame}

\newcommand{\hnegneg}{\hil{$\neg\neg$}}
\newcommand{\negg}{{\neg}\!\!\!{\neg}}
\newcommand{\neggv}{{\neg}\!\!\!\phantom{{\neg}}}
\newcommand{\bott}{\bot\!\!\!\!\bot}
\newcommand{\bottv}{\bot\!\!\!\!\phantom{\bot}}

\begin{frame}{The double-negation embedding}
  \justifying
  \vspace*{-1em}
  \small
  \textbf{Def.} For formulas over a fixed first-order signature, the
  \hil{$\neg\neg$-translation} $\varphi \mapsto \varphi^{\neg\neg}$ is defined
  by the following clauses.
  \begin{align*}
    (\varphi_\text{atomic})^{\neg\neg} &\defeqv \hnegneg \varphi_\text{atomic} &
    (\varphi \Rightarrow \psi)^{\neg\neg} &\defeqv (\varphi^{\neg\neg} \Rightarrow \psi^{\neg\neg}) \\
    \bot^{\neg\neg} &\defeqv \hnegneg\bot &
    \top^{\neg\neg} &\defeqv \top \\
    (\varphi \vee \psi)^{\neg\neg} &\defeqv \hnegneg(\varphi^{\neg\neg} \vee \psi^{\neg\neg}) &
    (\varphi \wedge \psi)^{\neg\neg} &\defeqv (\varphi^{\neg\neg} \wedge \psi^{\neg\neg}) \\
    (\exists x\?X\_ \varphi)^{\neg\neg} &\defeqv \hnegneg{(\exists x\?X\_ \varphi^{\neg\neg})} &
    (\forall x\?X\_ \varphi)^{\neg\neg} &\defeqv (\forall x\?X\_ \varphi^{\neg\neg}) \\
  \end{align*}
  \vspace*{-3em}

  \mbox{\textbf{Ex.} $(\forall a{:}X\_\! \exists b{:}X\_\! a = b \vee \ldots)^{\neg\neg}
  \equiv (\forall a{:}X\_\! \hnegneg\exists b{:}X\_\! \hnegneg(\hnegneg(a=b) \vee
  (\ldots)^{\neg\neg})).$}
  \pause

  \textbf{Prop.} Classically, $\varphi \Leftrightarrow \varphi^{\neg\neg}$.
  \pause
  \vspace*{-0.5em}

  \grayline

  \textbf{Thm.} For every formula~$\varphi$ and set of formulas~$\Gamma$:
  \vspace*{-0.6em}
  \begin{enumerate}
    \item Minimally, $\neg\neg(\varphi^{\neg\neg}) \Rightarrow
    \varphi^{\neg\neg}$.
    \item Minimally, $\varphi^{\neg\neg} \Leftrightarrow \neg\neg\varphi$ in
    case that~$\varphi$ is geometric
    ($R{\top}{\bot}{\wedge}{\vee}{\exists}{\bigvee}$).
    \item If~$\Gamma$ entails~$\varphi$ classically,
    then~$\Gamma^{\neg\neg}$ entails~$\varphi^{\neg\neg}$ minimally.
  \end{enumerate}
  \pause
  \vspace*{-0.3em}

  \textbf{Cor.} If PA proves~$\varphi$, then~HA proves~$\varphi^{\neg\neg}$.
  \pause

  \textbf{Rem.} Theorem and corollary hold for every \hil{local
  operator}~$\nabla$ in place of~$\neg\neg$, in particular for~$\negg\negg\varphi \defeq
  ((\varphi \Rightarrow \bott) \Rightarrow \bott)$ for some arbitrary
  formula~$\bott$.
\end{frame}

\newcommand{\negx}{\hil{${\only<1>{\neggv}\only<2->{\negg}}$}}
\newcommand{\botx}{\hil{${\only<1>{\bottv}\only<2>{\bott}\only<3->{\beta}}$}}

\begin{frame}{Barr's theorem / Friedman's trick / A-translation}
  \textbf{Thm.} Let~$\Gamma$ be a set of geometric sequents over a fixed
  signature. Let~$\sigma$ be a geometric sequent. Then the following are
  equivalent:
  \begin{enumerate}
    \addtocounter{enumi}{-1}
    \item $\sigma$ holds for the \good{generic model} of~$\Gamma$ (in its
    classifying topos).
    \item $\sigma$ is provable from~$\Gamma$ in \good{geometric logic}.
    \item $\sigma$ is provable from~$\Gamma$ in \good{intuitionistic logic}.
    \item $\sigma$ is provable from~$\Gamma$ in \bad{classical logic}.
    \item (Assuming \badbox{\textsc{ac}}) $\sigma$ is provable from~$\Gamma$ in
    \bad{classical logic with~\textsc{ac}}.
  \end{enumerate}

  \emph{Proof of~``\,\,$\normalnumber{3} \Rightarrow \normalnumber{2}$''.}
  Write~$\sigma \equiv (\alpha \seq{\vec x} \beta)$. Then intuitionistically,
  \[ \alpha \Longrightarrow
    \negx\negx\alpha \Longleftrightarrow
    \alpha^{\negx\negx} \Longrightarrow
    \beta^{\negx\negx} \Longleftrightarrow
    \negx\negx\beta \equiv
    ((\beta{\Longrightarrow}\botx){\Rightarrow}\botx)
    \visible<4->{\Longrightarrow \beta.}
  \]
\end{frame}

\section{Constructive proofs from invalid proofs}

\begin{frame}
  \centering\bigskip
  \includegraphics[height=9.5em]{phantoms}
  \medskip

  \large Lecture III: \\
  \hil{Extracting constructive proofs from invalid* proofs}

  \normalsize
\end{frame}

\addtocounter{framenumber}{-1}

\end{document}

start:
- infinitude of primes
- Dickson's lemma
- "Every constructive theorem has a computable witness."
- induction = recursion, Markov = unbounded search, ...
- side effects
- applications:
  - integrated developments (for instance SAT checking)
  - metatheory of constructive systems
  - philosophy
  - exploring computability theory

definition:
- HA
- TM
- realizability (as formalization of BKH)
- proof

examples:
- as in filmat paper
  - also with lambda terms!

special kinds of statements:
- negated: uninformative
- double negation
- Π⁰₁

metatheory of HA:
- disjunction property
- existence property
- unprovability of (for instance) halting
- growth rate

variants:
- ITTM
- real world
- completion to Eff

----

- play
- double-negation embedding
- Friedman's trick
- perhaps in Agda

\begin{block}{}
  \justifying
  \textbf{Thm.}
  Every pair of infinite sequences~$\alpha,\beta : \NN \to \NN$ is \emph{good} in that
  there are~$i < j$ such that~$\alpha(i)
  \leq \alpha(j)$ and~$\beta(i) \leq \beta(j)$.
\end{block}
\vspace*{-0.6em}
{\small\emph{Proof.} By~\badbox{\textsc{lem}}+\badbox{\textsc{dc}}, there is a
monotonic subsequence
\[ \alpha(i_0) \leq \alpha(i_1) \leq \alpha(i_2) \leq \cdots \]
with~$i_0 < i_1 < \cdots$. By~\badbox{\textsc{lem}}, there is a minimal
value~$\beta(i_{k_0})$ among all values of~$(\beta(i_k))_k$. Then
$\alpha(i_{k_0}) \leq \alpha(i_{k_0+1})$ and
$\beta(i_{k_0}) \leq \beta(i_{k_0+1})$. \qed
\par}

Claim: (∃n ∈ X) ⇒ X has a minimum.

Classical proof:
By LEM, ∃(k ∈ X). k < n or not.
In the first case, we conclude by induction.
In the second case, we conclude that n is minimal:
  Given k ∈ X, we have k < n or n ≤ k.
  In the first case, ⊥ and hence conclusion.
  In the second case, good.

Translated proof:
Assume that X does not have a minimum, show ⊥.
May conclude that ¬(∃(k ∈ X). k < n). For else, X inhabited by <n, so ⊥ by induction.
Thus n is a minimum:
  Given k ∈ X, we have k < n or n ≤ k.
  In the first case, ⊥ and hence conclusion.
  In the second case, good.

\begin{tabular}{ll@{\qquad\qquad}ll}
  \toprule
  $\varphi$ & $\varphi^{\neg\neg}$ &
  $\varphi$ & $\varphi^{\neg\neg}$ \\\midrule
  $\alpha_\text{atomic}$ & $\hnegneg \alpha_\text{atomic}$ &
  $\alpha \Rightarrow \beta$ & $\alpha^{\nabla} \Rightarrow \beta^{\nabla}$ \\
  $\bot^{\neg\neg}$ & $\hnegneg\bot$ &
  $\top^{\neg\neg}$ & $\top$ \\
  $\alpha \vee \beta$ & $\hnegneg(\alpha^{\neg\neg} \vee \beta^{\neg\neg})$ &
  $\alpha \wedge \beta$ & $\alpha^{\neg\neg} \wedge \beta^{\neg\neg}$ \\
  \bottomrule
\end{tabular}