### 2022-11-03 slides-paris2022.tex
Structure of Paris talk

A modal logical multiverse for commutative algebra and combinatorics

In the spirit of the set-theoretic multiverse philosophy put forward by
Joel David Hamkins, we explore a related modal multiverse populated by
Kripke models and more general worlds. In this
multiverse—as the talk will explain—the law of excluded middle can be
switched on and off like a light bulb and countability is a button (for
every set X of every world, there is a larger world containing a
surjection ℕ → X). Our interest in this multiverse is because of
concrete applications in commutative algebra and combinatorics,
including the endeavor of extracting algorithms from proofs utilizing
transfinite techniques. The talk will be framed by several examples of
this kind.

=== Motivation

Infinitude of primes
- algorithm visible from the proof

Two motivating questions:
- Dickson's lemma
- Surjectivity of matrices
How to extract algorithms here?
Spoil elementary proof of surjective case

=== Set-theoretic multiverse

Quick sketch of history including dream solution to CH

Multiverse position: acknowledge that notion of set is underdetermined by ZFC
axioms, embrace multiverse of models

Switches and buttons

Analogy with field extensions

=== Relaxation to topos-theoretic case

"Parametric mathematics" (esp. visible with toposes over spaces)

Modal operators

Examples of modal statements