### 2025-06-20 slides-types2025.lagda.tex
Notes slides

% compile using: agda --latex; result will then be in the latex/ subfolder
\documentclass[11pt,utf8,notheorems,compress,t,aspectratio=169]{beamer}
\usepackage{etex}

\usepackage{pgfpages}
\usepackage[export]{adjustbox}

% Workaround for the issue described at
% https://tex.stackexchange.com/questions/164406/beamer-using-href-in-notes.
\newcommand{\fixedhref}[2]{\parbox[t][0pt][t]{0pt}{\vspace*{-\paperheight}\vspace*{-0.6em}\href{#1}{#2}}\href{#1}{#2}}

\usepackage[english]{babel}

\usepackage{mathtools}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{stmaryrd}
\usepackage{amssymb}
\usepackage{manfnt}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{multicol}
\usepackage{tabto}
\usepackage{xstring}
\usepackage{proof}
\usepackage{agda}
\usepackage[all]{xy}
\xyoption{rotate}
\usepackage{tikz}
\usetikzlibrary{calc,shapes,shapes.callouts,shapes.arrows,patterns,fit,backgrounds,decorations.pathmorphing,positioning,svg.path}
\hypersetup{colorlinks=true}
\usepackage[dvipsnames]{xcolor}
\usepackage{fontawesome5}

\DeclareFontFamily{U}{bbm}{}
\DeclareFontShape{U}{bbm}{m}{n}
   {  <5> <6> <7> <8> <9> <10> <12> gen * bbm
      <10.95> bbm10%
      <14.4>  bbm12%
      <17.28><20.74><24.88> bbm17}{}
\DeclareFontShape{U}{bbm}{m}{sl}
   {  <5> <6> <7> bbmsl8%
      <8> <9> <10> <12> gen * bbmsl
      <10.95> bbmsl10%
      <14.4> <17.28> <20.74> <24.88> bbmsl12}{}
\DeclareFontShape{U}{bbm}{bx}{n}
   {  <5> <6> <7> <8> <9> <10> <12> gen * bbmbx
      <10.95> bbmbx10%
      <14.4> <17.28> <20.74> <24.88> bbmbx12}{}
\DeclareFontShape{U}{bbm}{bx}{sl}
   {  <5> <6> <7> <8> <9> <10> <10.95> <12> <14.4> <17.28>%
      <20.74> <24.88> bbmbxsl10}{}
\DeclareFontShape{U}{bbm}{b}{n}
   {  <5> <6> <7> <8> <9> <10> <10.95> <12> <14.4> <17.28>%
      <20.74> <24.88> bbmb10}{}
\DeclareMathAlphabet{\mathbbm}{U}{bbm}{m}{n}
\SetMathAlphabet\mathbbm{bold}{U}{bbm}{bx}{n}

\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}
\DeclareSymbolFont{extraup}{U}{zavm}{m}{n}
\DeclareMathSymbol{\varheart}{\mathalpha}{extraup}{86}

\graphicspath{{images/}}

\usepackage[protrusion=true,expansion=true]{microtype}

\setlength\parskip{\medskipamount}
\setlength\parindent{0pt}

\title{Towards topological type theory for decrypting transfinite methods in classical mathematics}

\author{Ingo Blechschmidt}
\date{June 9th, 2025}

\setbeameroption{show notes on second screen=bottom}
\newcommand{\jnote}[2]{\only<#1>{\note{\setlength\parskip{\medskipamount}\footnotesize\justifying#2\par}}}
\makeatletter
\setbeamertemplate{note page}
{%
  {%
    \scriptsize
    \insertvrule{.25\paperheight}{white!90!black}
    \vskip-.25\paperheight
    \nointerlineskip
    \vbox{
      %\hfill\insertslideintonotes{0.25}\hskip-\Gm@rmargin\hskip0pt%
      %\vskip-0.25\paperheight%
      %\nointerlineskip
      \begin{pgfpicture}{0cm}{0cm}{0cm}{0cm}
        \begin{pgflowlevelscope}{\pgftransformrotate{90}}
          {\pgftransformshift{\pgfpoint{-2cm}{0.2cm}}%
          \pgftext[base,left]{\footnotesize\the\year-\ifnum\month<10\relax0\fi\the\month-\ifnum\day<10\relax0\fi\the\day}}
        \end{pgflowlevelscope}
      \end{pgfpicture}}
    \nointerlineskip
    \vbox to .25\paperheight{\vskip0.5em
      \hbox{\insertshorttitle[width=8cm]}%
      \setbox\beamer@tempbox=\hbox{\insertsection}%
      \hbox{\ifdim\wd\beamer@tempbox>1pt{\hskip4pt\raise3pt\hbox{\vrule
            width0.4pt height7pt\vrule width 9pt
            height0.4pt}}\hskip1pt\hbox{\begin{minipage}[t]{7.5cm}\def\breakhere{}\insertsection\end{minipage}}\fi%
      }%
      \setbox\beamer@tempbox=\hbox{\insertsubsection}%
      \hbox{\ifdim\wd\beamer@tempbox>1pt{\hskip17.4pt\raise3pt\hbox{\vrule
            width0.4pt height7pt\vrule width 9pt
            height0.4pt}}\hskip1pt\hbox{\begin{minipage}[t]{7.5cm}\def\breakhere{}\insertsubsection\end{minipage}}\fi%
      }%
      \setbox\beamer@tempbox=\hbox{\insertshortframetitle}%
      \hbox{\ifdim\wd\beamer@tempbox>1pt{\hskip30.8pt\raise3pt\hbox{\vrule
            width0.4pt height7pt\vrule width 9pt
            height0.4pt}}\hskip1pt\hbox{\insertshortframetitle[width=7cm]}\fi%
      }%
      \vfil}%
  }%
  \vskip.25em
  \nointerlineskip
  \insertnote
}
\makeatother

%\useinnertheme[shadow=true]
\setbeamerfont{block title}{size={}}

\useinnertheme{rectangles}

\usecolortheme{orchid}
\usecolortheme{seahorse}
\definecolor{mypurple}{RGB}{253,73,34}
\definecolor{mypurpledark}{RGB}{100,0,150}
\setbeamercolor{structure}{fg=mypurple}
\setbeamercolor*{title}{bg=mypurple,fg=white}
\setbeamercolor*{titlelike}{bg=mypurple,fg=white}
\setbeamercolor{frame}{bg=black}

\usefonttheme{serif}
\usepackage[T1]{fontenc}
\usepackage{libertine}

% lifted from https://arxiv.org/abs/1506.08870
\DeclareFontFamily{U}{min}{}
\DeclareFontShape{U}{min}{m}{n}{<-> udmj30}{}
\newcommand\yon{\!\text{\usefont{U}{min}{m}{n}\symbol{'210}}\!}

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\M}{\mathcal{M}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\BB}{\mathbb{B}}
\newcommand{\pp}{\mathbbm{p}}
\newcommand{\MM}{\mathbb{M}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\FF}{\mathbb{F}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\GG}{\mathbb{G}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\bbb}{\mathfrak{b}}
\newcommand{\ccc}{\mathfrak{c}}
\newcommand{\ppp}{\mathfrak{p}}
\newcommand{\fff}{\mathfrak{f}}
\newcommand{\mmm}{\mathfrak{m}}
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\defeqv}{\vcentcolon\equiv}
\newcommand{\Cov}{\mathrm{Cov}}
\renewcommand{\_}{\mathpunct{.}}
\newcommand{\?}{\,{:}\,}
\newcommand{\speak}[1]{\ulcorner\text{\textnormal{#1}}\urcorner}
\newcommand{\inv}{inv.\@}
\newcommand{\forces}{\vDash}
\newcommand{\ind}{\ensuremath{_\text{ind}}\xspace}
\newcommand{\sinf}{\ensuremath{_\infty}}
\newcommand{\impl}{\ensuremath{_\text{impl}}\xspace}
\newcommand{\formalized}{{\color{NavyBlue!75!White}{\raisebox{-0.5pt}{\scalebox{0.8}{\faCog}}}}}

\setbeamertemplate{blocks}[shadow=false]

\newenvironment{indentblock}{%
  \list{}{\leftmargin\leftmargin}%
  \item\relax
}{%
  \endlist
}

% Adapted from https://latex.org/forum/viewtopic.php?t=2251 (Stefan Kottwitz)
\newenvironment<>{hilblock}{
  \begin{center}
    \begin{minipage}{9.05cm}
      \setlength{\textwidth}{9.05cm}
      \begin{actionenv}#1
        \def\insertblocktitle{}
        \par
        \usebeamertemplate{block begin}}{
        \par
        \usebeamertemplate{block end}
      \end{actionenv}
    \end{minipage}
  \end{center}}

\newenvironment{changemargin}[2]{%
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \setlength{\leftmargin}{#1}%
    \setlength{\rightmargin}{#2}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\itemindent}{\parindent}%
    \setlength{\parsep}{\parskip}%
  }%
  \item[]}{\end{list}}

\tikzset{
  invisible/.style={opacity=0,text opacity=0},
  visible on/.style={alt={#1{}{invisible}}},
  alt/.code args={<#1>#2#3}{%
    \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}}}
}

% https://tex.stackexchange.com/questions/172336/drawing-roman-laurel-leaves-spqr-in-tikz
\tikzset{
  laurel-wreath/.pic = {
    \fill svg{M14.4-24.6c-1.5-1.5-2.6-3.3-3.1-5.3l-.4-1.7c-.2-1.1-.2-4.1 .2-5.7 .2-.9 .3-1.3 .5-1.3l1.4 1.1 2.5 2.4c2.7 2.5 5.2 6 5.8 8 .2 .6-.5 .3-2.2-.9-1.6-1.3-3.3-2.6-5-3.8l.1 1.4c.2 1.4 .5 2.7 1.1 4.6s.8 2.5 .5 2.5l-1.4-1.3zm69.6 1.1 .3-1.2c.8-2.3 1.3-4.8 1.6-7.3l-1.5 1.1c-1.3 .9-2.6 1.9-3.7 3-1.6 1.1-2 1.3-2.1 1 .7-1.8 1.6-3.4 2.8-4.9 1.3-1.7 6.5-6.8 7-6.8 .2 0 .3 .2 .3 .5l.3 1.6c.3 2.2 .2 5.7-.5 7.4-.8 1.9-1.6 3.1-3 4.7-1.1 1.1-1.4 1.3-1.5 .9z};
    \fill svg{M10-29.4c-.8-1.1-1.4-2.2-2-4.1l-.7-3.5c-.2-3 .2-4.4 1.4-8.3l.5-1.4c.2-1.3 .3-1.9 .6-1.9 .3-.2 .6 .3 .7 .8s.9 2.2 1.9 3.6c1.4 2.2 2.7 4.4 3.9 6.6l.9 2.7c0 .6 0 .6-.3 .6-.6 0-4.9-4.4-5.8-6l-.2-.6-.1 1.7-.3 2.8c-.3 2.7-.3 3.8 0 5.5 .6 2 .5 2.4-.5 1.5zm79.2 .3 .4-2.4c.2-1.3 .2-2.7-.1-4.9l-.3-2.8v-1.6l-.7 1c-.8 1.3-5 5.5-5.5 5.5s-.5-.3 .2-1.9c.5-1.7 1.4-3.3 3.3-6.5 2.4-3.6 2.7-3.9 2.8-4.7 .5-1.3 .5-1.4 .8-1.2 .3 0 .6 .8 .6 1.5l.7 2.4c.9 2.7 1.1 3.6 1.2 6 .2 3.1-.5 6-2 8.2-.8 1.3-1.3 1.7-1.4 1.5z};
    \fill svg{M5-40c-.4-3.2-.1-6.5 .9-9.6 .5-1.1 1.6-2.8 2.2-3.4l1.3-1.6 2-2.7 .2 .6c.1 1.3 .4 2.6 .9 3.8l.3 1c.8 1.7 1.1 2.7 1.6 5.3 .6 2.5 .6 4.6 .2 4.6-.3 0-.9-.8-1-1.1l-.5-.8c-1.4-2-3-5.2-2.9-6.5-.9 2.7-2 5.4-3.5 7.9l-.3 .8-.3 .8c0 .5-.6 1.6-.8 1.6l-.3-.7zm89.2 .2-.2-.5-.3-.9-1.1-2.7-1.1-2.4c-.6-1.4-1.2-2.8-1.6-4.2l-.3 .9c-.3 1.3-1.6 3.9-3 6-1.3 2-1.6 2-1.5 0s1.1-6.3 2.2-9c.8-1.7 1.1-3.1 .9-4.1-.2-1.1 .5-.8 2.2 1.8 3.3 4.4 3.8 5.4 4.4 7.8 .6 2.4 .5 7.7-.3 7.8l-.3-.5z};
    \fill svg{M13.9-50.1c-.5-1.9-.8-3.9-.9-5.8-.2-1.6-.1-3.3 .1-4.9-.3 .8-1.7 2.5-4.2 5.1l-3 4.9-.3 .1c-.3 0-.3-2.2 0-3.3 .8-3 1.4-4.6 2.5-6.1 .9-1.3 1.7-1.9 2.5-2.5 1.1-.6 2.7-1.9 3.5-2.7 .9-.9 1.9-1.4 2.2-1.4v1.1l-.3 6.6c0 6.8 .2 6.3-1 8.9-.5 1.1-.8 1.1-1.1 0zm70.8-.4c-.8-2.2-.8-2.5-.7-6.3-.1-2.7-.1-5.5-.2-8.2-.3-1.6-.3-1.9 .5-1.6l.6 .5c1.4 1.4 3 2.5 3.9 3.1 1.3 .9 1.9 1.6 2.7 2.6l.6 .7 .2 .4 .2 .3c.8 .9 2 4.9 2 6.9 .2 1.9-.2 1.9-.9 .5-.7-1.4-1.5-2.7-2.6-4-1.6-1.5-3-3.2-4.2-5 .4 3 .3 6-.5 9 0 .8-.5 2.2-.8 2.3-.2 0-.5-.3-.8-1.2z};
    \fill svg{M16.4-58.5l.2-1.5 .3-3.7c.2-2.8 .3-3.5 1.1-5.4l.7-1.3-.5 .4-1 .7c-.5 .4-1.1 .8-1.5 1.3l-.5 .3-1.9 1.6c-2.2 1.6-2.7 2-3.9 3.6-.5 .8-1.1 1.3-1.3 1.3-.5 0 0-2.4 1.1-4.7 1.5-3.4 4.3-6 7.7-7.4l1.3-.4 1.9-.4 2-.5c1.4 0 1.4 0 1 1.1-.5 .8-.8 2-1.1 4.2-.3 2.3-1.1 4.5-2.2 6.5l-.4 .6c-.6 1.1-1.3 2.1-2 3.2-.5 .6-.8 .8-1 .5zm66.3-.2c-.8-.9-2.8-4.4-3.5-6.1-.6-1.3-.9-2.5-1.1-3.5-.2-2.1-.7-4.1-1.5-6 0-.3 0-.3 1.2-.3l2.1 .5 1.9 .4 1.2 .4 .6 .1 1 .6c3 1.4 5.7 4.6 6.8 8.5l.7 2.6c-.2 .6-.5 .5-1.4-.7-2.2-2.7-4.8-5-7.7-6.9l-1.7-1.3 .6 1.3c.3 .6 .6 1.2 .8 1.9l.3 2.5 .3 3.9c.3 2.4 .2 2.8-.6};
    \fill svg{M21.6-66.1l.4-1.1 .9-3.2c.3-1.9 1.1-3.3 2.4-4.7l.4-.8-1.2 .2-2.2 .3c-2.7 .3-5.3 1.2-7.7 2.5-.6 .5-1.3 .6-1.3 .3 0-.5 .9-1.9 2-2.9 .8-.9 2-1.9 3.2-2.6l.9-.4 2.2-1c.3-.2 1.3-.3 3.2-.1 3 0 4.1 .2 6.3 .7l1.1 .4c.5 .2 .6 .6 .3 .6-.5 0-1.4 .9-1.9 1.7l-1.2 1.8c-1.7 2.8-2.2 3.5-4.6 5.9l-3 2.7-.2-.3zm53.9-2c-2.7-2.8-3.5-3.8-5.4-6.8-.9-1.6-1.4-2.4-1.9-2.5l-.8-.5c-.3 0-.2-.5 .4-.6l1.1-.4c1.9-.6 3-.8 5.6-.9l3.3 .2c2 .6 3.8 1.5 5.4 2.8 .3 0 1.9 1.6 2.5 2.4l.9 1.8c0 .3-.3 .2-1.9-.6-2.8-1.4-4.4-1.9-7.7-2.2l-2.2-.5c-.9-.2-.9-.2-.6 .2 .6 .5 1.7 2 2.1 2.8l.9 2.5c.3 1.5 .6 3 .9 4.6l-2.6-2.3z};
    \fill svg{M34.1-78.7c-3.4-1.3-6.9-2.1-10.6-2.5-.9 0-1.4 0-2.3 .3-2 .5-2 0 0-1.3l2.8-1.2c1.4-.5 1.9-.5 3.8-.6 3.8-.2 6.1 .3 9.3 1.7l3.6 1.1 2.2 .3c1.3 0 1.7 0 2.7-.3 1.1-.3 2.8-1.1 2.8-1.3l-1.3-.9c-1.9-1.4-3.1-2.7-3.1-3.2l.8-.6c.9-.3 1.3-.2 2 .8 .5 .8 1.1 1.4 2.9 2.7 .2 .3 .3 .2 1.1-.3 .9-.8 2.4-2 2.6-2.7 .5-.6 .9-.8 1.8-.5l.8 .6c0 .5-1.4 1.7-3.2 3.2l-1.3 .9c0 .2 1.7 .9 2.9 1.3 .9 .3 1.4 .3 2.7 .3l2.2-.3c1.7-.4 3.4-1 5-1.7 2-.8 4.4-1.3 7.7-1.1 2 .2 2.5 .2 3.8 .6 .9 .3 2.2 .8 2.8 1.2 2 1.1 2 1.6 .2 1.3-1.6-.3-1.9-.3-4.4 0-2.4 .3-4.7 .8-7 1.6l-1.5 .6c-2.9 .3-5.9 .2-8.8-.3-1.7-.3-3.6-.9-6-2.1l-1.1-.4-1.3 .6c-4.5 2.2-9.6 3-14.6 2.2zm-6.3-9.1c};
  }
}

\newcommand{\pointthis}[3]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=0,outer sep=0] (#2) {#2};
    \node[visible on=#1,overlay,rectangle callout,rounded corners,callout relative pointer={(0.3cm,0.5cm)},fill=blue!20] at ($(#2.north)+(-0.1cm,-1.1cm)$) {#3};
  }%
}

\tikzset{
  invisible/.style={opacity=0,text opacity=0},
  visible on/.style={alt={#1{}{invisible}}},
  alt/.code args={<#1>#2#3}{%
    \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}}}
}

\newcommand{\hcancel}[5]{%
  \tikz[baseline=(tocancel.base)]{
    \node[inner sep=0pt,outer sep=0pt] (tocancel) {#1};
    \draw[red!80, line width=0.4mm] ($(tocancel.south west)+(#2,#3)$) -- ($(tocancel.north east)+(#4,#5)$);
  }%
}

\newcommand{\explain}[7]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=2pt,outer sep=0,fill=#3,rounded corners] (label) {#1};
    \node[anchor=north,visible on=<#2>,overlay,rectangle callout,rounded corners,callout
    relative pointer={(0.0cm,0.5cm)+(0.0cm,#6)},fill=#3] at ($(label.south)+(0,-0.3cm)+(#4,#5)$) {#7};
  }%
}

\newcommand{\explainstub}[2]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=2pt,outer sep=0,fill=#2,rounded corners] (label) {#1};
  }%
}

\newcommand{\squiggly}[1]{%
  \tikz[remember picture,baseline]{
    \node[anchor=base,inner sep=0,outer sep=0] (label) {#1};
    \draw[thick,color=red!80,decoration={snake,amplitude=0.5pt,segment
    length=3pt},decorate] ($(label.south west) + (0,-2pt)$) -- ($(label.south east) + (0,-2pt)$);
  }%
}

% Adapted from https://latex.org/forum/viewtopic.php?t=2251 (Stefan Kottwitz)
\newenvironment<>{varblock}[2]{\begin{varblockextra}{#1}{#2}{}}{\end{varblockextra}}
\newenvironment<>{varblockextra}[3]{
  \begin{center}
    \begin{minipage}{#1}
      \begin{actionenv}#4
        {\centering \hil{#2}\par}
	\def\insertblocktitle{}%\centering #2}
        \def\varblockextraend{#3}
	\usebeamertemplate{block begin}}{
        \par
        \usebeamertemplate{block end}
        \varblockextraend
      \end{actionenv}
    \end{minipage}
  \end{center}}

\setbeamertemplate{headline}{}

\setbeamertemplate{frametitle}{%
  \leavevmode%
  \vskip-1.6em%
  \begin{beamercolorbox}[dp=1ex,center,wd=\paperwidth,ht=2.25ex]{title}%
    \vskip0.5em%
    \bf\insertframetitle
  \end{beamercolorbox}%

  \vskip-0.77em\hspace*{-2em}%
  \textcolor{mypurpledark}{\rule[0em]{1.1\paperwidth}{2.4pt}}

  \vskip-0.4em%
}

\setbeamertemplate{navigation symbols}{}

\newcounter{framenumberpreappendix}
\newcommand{\backupstart}{
  \setcounter{framenumberpreappendix}{\value{framenumber}}
}
\newcommand{\backupend}{
  \addtocounter{framenumberpreappendix}{-\value{framenumber}}
  \addtocounter{framenumber}{\value{framenumberpreappendix}}
}

\newcommand{\insertframeextra}{}
\setbeamertemplate{footline}{%
  \begin{beamercolorbox}[wd=\paperwidth,ht=2.25ex,dp=1ex,right,rightskip=1mm,leftskip=1mm]{}%
    % \inserttitle
    \hfill
    \insertframenumber\insertframeextra\,/\,\inserttotalframenumber
  \end{beamercolorbox}%
  \vskip0pt%
}

\newcommand{\hil}[1]{{\usebeamercolor[fg]{item}{\textbf{#1}}}}
\newcommand{\hill}[1]{{\usebeamercolor[fg]{item}{#1}}}
\newcommand{\bad}[1]{\textcolor{red!90}{\textnormal{#1}}}
\newcommand{\good}[1]{\textcolor{mypurple}{\textnormal{#1}}}

\newcommand{\bignumber}[1]{%
  \renewcommand{\insertenumlabel}{#1}\scalebox{1.2}{\!\usebeamertemplate{enumerate item}\!}
}
\newcommand{\normalnumber}[1]{%
  {\renewcommand{\insertenumlabel}{#1}\!\usebeamertemplate{enumerate item}\!}
}
\newcommand{\bigheart}{\includegraphics{heart}}

\newcommand{\subhead}[1]{{\centering\textcolor{gray}{\hrulefill}\quad\textnormal{#1}\quad\textcolor{gray}{\hrulefill}\par}}

\newcommand{\badbox}[1]{\colorbox{red!30}{#1}}
\newcommand{\infobox}[1]{\colorbox{yellow!70}{\color{black}#1}}

% taken from JDH "The modal logic of arithmetic potentialism and the universal algorithm"
\DeclareMathOperator{\possible}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,rotate=45,line width=.1ex]{\draw (-1,-1) rectangle (1,1);}}}
\DeclareMathOperator{\necessary}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,line width=.1ex]{\draw (-1,-1) rectangle (1,1);}}}
\DeclareMathOperator{\xpossible}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,rotate=45,line width=.1ex]{\draw (-1,-1) rectangle (1,1); \draw[very thin] (-.6,-.6) rectangle (.6,.6);}}}
\DeclareMathOperator{\xnecessary}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,line width=.1ex]{\draw (-1,-1) rectangle (1,1); \draw[very thin] (-.6,-.6) rectangle (.6,.6);}}}

\input{images/primes.tex}

\newcommand{\triang}{\hil{$\blacktriangleright$}}
\newcommand{\concat}{\mathbin{{+}\mspace{-8mu}{+}}}

\newcommand{\astikznode}[2]{\tikz[baseline,remember picture]{\node[anchor=base,inner sep=0,outer sep=0.1em] (#1) {#2};}}
\newcommand{\astikznodecircled}[3]{\tikz[baseline,remember picture]{\node[anchor=base,circle,draw=#2,thick,inner sep=0.05em,outer sep=0.05em] (#1) {#3};}}
\newcommand{\astikznodetransparentlycircled}[2]{\tikz[baseline,remember picture]{\node[anchor=base,circle,opacity=0,draw=white,text opacity=1,thick,inner sep=0.05em,outer sep=0.05em] (#1) {#2};}}

\setbeamersize{text margin left=1.60em,text margin right=1.60em}

\newlength\stextwidth
\newcommand\makesamewidth[3][c]{%
  \settowidth{\stextwidth}{#2}%
  \makebox[\stextwidth][#1]{#3}%
}

\newcommand{\dnote}[1]{%
  \begin{tabular}{@{}m{2em}@{}m{0.83\textwidth}@{}}%
    \textdbend &#1%
  \end{tabular}%
  \par
}

\newcommand{\genalpha}{\mbox{$\hspace{0.12em}\shortmid\hspace{-0.62em}\alpha$}}

\usepackage{newunicodechar}

\newunicodechar{∇}{\ensuremath{\nabla}}
\newunicodechar{λ}{\ensuremath{\lambda}}
\newunicodechar{σ}{\ensuremath{\sigma}}
\newunicodechar{τ}{\ensuremath{\tau}}
\newunicodechar{∷}{\ensuremath{\dblcolon}}
\newunicodechar{⧺}{\ensuremath{\!\!\!}}
\newunicodechar{₁}{\ensuremath{_1}}
\newunicodechar{₂}{\ensuremath{_2}}
\newunicodechar{∈}{\ensuremath{\in}}
\newunicodechar{ℕ}{\ensuremath{\NN}}
\newunicodechar{↠}{\ensuremath{\twoheadrightarrow}}
\newunicodechar{ʳ}{\ensuremath{^r}}

\begin{document}

\addtocounter{framenumber}{-1}

\definecolor{mypurpleblack}{RGB}{30,0,50}
\setbeamercolor{structure}{fg=black}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\centering\vspace*{-7em}\includegraphics[width=\paperwidth]{forest2}\end{minipage}}
\begin{frame}[c]
  \centering
  \color{white}

  \bigskip
  \bigskip
  \bigskip
  \bigskip
  \bigskip
  \bigskip
  \bigskip
  \bigskip
  \bigskip

  \scriptsize

  \setbeamercolor{block body}{bg=black!100}
  \begin{minipage}{0.52\textwidth}
    \begin{block}{}
      \centering\normalsize\color{white}
      \hil{\color{white}Towards topological type theory for decrypting
      transfinite methods in classical~mathematics} \\[-0.4em]
      \
    \end{block}
  \end{minipage}

  \bigskip
  \bigskip
  \bigskip
  \bigskip
  \bf
  \colorbox{black}{\begin{minipage}{0.2\textwidth}
    \centering
    TYPES 2025 \\
    June 9th, 2025
  \end{minipage}}
  \bigskip

  \colorbox{black}{\begin{minipage}{0.2\textwidth}
    \centering
    Ingo Blechschmidt \\
    University of Antwerp
  \end{minipage}}
\end{frame}}

\definecolor{mypurple}{RGB}{150,0,255}
\setbeamercolor{structure}{fg=mypurple}

\addtocounter{framenumber}{-1}
\begin{frame}\frametitle{A classical logic fairy tale}
  \centering\bigskip\bigskip\bigskip
  \includegraphics[height=9.5em]{philosophers-stone}

  \jnote{1}{
    \vspace*{-1em}
    \begin{changemargin}{-1.5em}{-1.5em}
      \setlength{\columnsep}{15pt}
      \begin{multicols}{2}
        \justifying
        \scriptsize

        \textbf{Narrator.} Once upon a time, in a kingdom far, far away, the
        queen of the country and of all Möbius strips called for her royal
        philosopher.

        \textbf{Queen.} Philosopher! I ask you to carry out the following order.
        Get me the Philosopher's Stone, or alternatively find out how one could
        produce arbitrary amounts of gold with it!

        \textbf{Philosopher.} But my queen! I haven't studied anything useful!
        How could I fulfill this order?

        \textbf{Queen.} That is not my concern. I'll see you again tomorrow.
        Should you not accomplish the task, I will take your head off.

        \textbf{Narrator.} After a long and wakeful night the philosopher was
        called to the queen again.

        \textbf{Queen.} Tell me! What do you have to report?

        \textbf{Philosopher.} It was not easy and I needed to follow lots of
        obscure references, but finally I actually found out how to use the
        Philosopher's Stone to produce arbitrary amounts of gold. But only I can
        conduct this procedure, your royal highness.

        \textbf{Queen.}
        Alright. So be it.

        \textbf{Narrator.} And so years passed by, during which the philosopher
        imagined herself to be safe. The queen searched for the stone on her own,
        but as long as she hadn't found it, the philosopher didn't need to worry.
        Yet one day the impossible happened: The queen has found the stone! And
        promptly called for her philosopher.

        \textbf{Queen.} Philosopher, look! I have found the Philosopher's Stone!
        Now live up to your promise! \emph{[She hands over the stone.]}

        \textbf{Philosopher.} Thank you. \emph{[She inspects the stone.]} This is
        indeed the Philosopher's Stone. Many years ago you asked me to either
        acquire the Philosopher's Stone or find out how to produce arbitrary
        amounts of gold using it. Now it's my pleasure to present to you the
        Philosopher's Stone. \emph{[She returns the stone.]}
      \end{multicols}
    \end{changemargin}
  }

  \jnote{2}{
    Adapted from
    \fixedhref{http://blog.ezyang.com/2013/04/a-classical-logic-fairy-tale/}{Edward
    Yang's blog}.

    The idea conveyed in the fairy tale is at the core of the
    \emph{constructive} proof of $\neg\neg(A \vee \neg A)$. While the law of
    excluded middle (\textsc{lem}) is not available in constructive
    mathematics, up to a double negation every instance of it is.

    This observation gives rise to the double negation translation, a
    transformation of (statements and) proofs in classical logic to proofs in
    intuitionistic logic, at the only expense of introducing double negations;
    by following this up with Friedman's trick with the nontrivial exit
    continuation (``Baby Barr Theorem''), we can in many cases even get rid of
    these double negations in a second step.

    Getting rid of applications of the law of excluded middle is in many
    contexts (an amazing mind-boggingly wonder at first, but then) a routine
    matter---see \href{https://rt.quasicoherent.io/}{here} for more details and
    pointers to the literature.
  }
\end{frame}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{1.7em}\includegraphics[height=\paperheight]{sea-of-clouds-1}\end{minipage}}
\begin{frame}{A case study in Hilbert's program}
  \jnote{1}{
    Well quasi-orders are an important notion in proof theory and termination
    analysis.
  }

  \jnote{2-3}{
    \vspace*{-0.5em}
    The presented proof rests on the law of excluded middle and hence cannot
    immediately be interpreted as a program for finding suitable indices~$i <
    j$. However, constructive proofs are also possible (for instance by
    induction on the value of a given term of the sequence, see
    \href{https://www.math.lmu.de/~petrakis/Dickson.pdf}{Constructive
    combinatorics of Dickson's Lemma} by Iosif Petrakis for several fine
    quantitative results). And even more:
    There is a procedure for regarding this proof---and many others in the
    theory of well quasi-orders---as \emph{blueprints} for more informative
    constructive proofs. This shall be our motto for today:

    \emph{Do not take classical proofs literally, instead ask which
    constructive proofs they are blueprints for.}
  }

  \jnote{3}{
    The displayed stability results, along with several others, provide a
    flexible toolbox for constructing new well quasi-orders from given ones.
    However, with the classical formulation of \emph{well},
    shortly to be renamed~``well\sinf'', these results are
    \href{https://lets-play-agda.quasicoherent.io/Padova2025.Explorations.Forcing.Orthogonal.html}{inherently
    classical} (even the weaker claim ``the product of two streamless sets is
    streamless'' admits countermodels).

    (In Higman's lemma, the set~$X^*$ of finite lists of elements of~$X$ is
    equipped with the following ordering: We have~$(x_0 \dblcolon \ldots
    \dblcolon x_{n-1} \dblcolon [\,]) \leq (y_0 \dblcolon \ldots \dblcolon
    y_{m-1} \dblcolon [\,])$ iff there is an increasing injection~$f : \{
    0,\ldots,n-1 \} \to \{ 0,\ldots,m-1 \}$ such that~$x_i \leq y_{f(i)}$ for all~$i < n$.
    Kruskal's theorem is the result behind the celebrated enormous number
    $\mathsf{TREE}(3)$.)
  }

  \jnote{7}{
    Thanks to work by Thierry Coquand, Daniel Fridlender and Monika
    Seisenberger, a constructive substitute is available, the notion well\ind.
    In classical mathematics (where bar induction is available), this notion is
    equivalent to well\sinf.

    The predicate~$\mathsf{Good}$ of finite lists~$\sigma = (x_0 \dblcolon
    \ldots \dblcolon x_{n-1} \dblcolon [\,])$ appearing in the definition
    of~$G$ expresses that some earlier term is at most some later term,
    i.\,e.\@ that there merely exist~$i < j$ such that~$x_i \leq x_j$.

    The operator $\dblcolon^r$ is adding an element at the end.
  }

  \jnote{8}{
    \begin{columns}[t]
      \begin{column}{0.65\textwidth}
        The original notion well\sinf:
        \begin{itemize}
          \justifying
          \item[\cmark] short and simple
          \\[-1.3em]
          \item[\cmark] constructively satisfied for the main examples (but
          only because of the theory around well\ind)
          \\[-1.3em]
          \item[\cmark] concise abstract proofs (albeit employing transfinite methods)
          \\[-1.3em]
          \item[\xmark] main results not constructively attainable
          \\[-1.3em]
          \item[\xmark] philosophically strenuous by the quantification over all sequences
          \\[-1.3em]
          \item[\xmark] not stable under ``change of base''---a forcing extension
          of the universe may well contain more sequences than the base universe
          \\[-1.3em]
          \item[\xmark] negative (universal) condition
        \end{itemize}
      \end{column}

      \begin{column}{0.33\textwidth}
        The constructive substitute well\ind:
        \begin{itemize}
          \justifying
          \item[\cmark] main results constructive
          \\[-1.3em]
          \item[\cmark] stable under change of base
          \\[-1.3em]
          \item[\cmark] positive (existential) condition
          \\[-1.3em]
          \item[\xmark] proofs intriguing, but also somewhat alien, not just some
          trivial reshuffling of the classical arguments; classical sequence
          language cannot be used
        \end{itemize}
      \end{column}
    \end{columns}
  }

  \textbf{Def.} Let~$(X,{\leq})$ be a quasi-order.
  \vspace*{-0.3em}
  \begin{itemize}
    \item A sequence~$\alpha : \NN \to X$ is \hil{good} iff
    there merely exist~$i < j$ with~$\alpha\,i \leq \alpha\,j$.
    \item The quasi-order~$X$ is \hil{well\only<4->{\sinf}} iff every sequence~$\NN \to X$ is good.
  \end{itemize}
  \pause
  \vspace*{-1.3em}

  \begin{columns}[t]
    \begin{column}{0.46\textwidth}
      \begin{block}{Natural numbers\phantom{y}}
        \textbf{Prop.} $(\NN, {\leq})$ is well\only<4->{\sinf}. \formalized
        \smallskip

        \emph{Proof.} Let~$\alpha : \NN \to \NN$. By~\badbox{\textsc{lem}}, there is a
        \bad{minimum}~$\alpha\,i$.
        Set~$j \defeq i + 1$. \qed

        \hfill\color{gray}offensive?
      \end{block}
      \only<2-6>{
        \vspace*{-2em}
        \[ \astikznodetransparentlycircled{xm}{7}\!,
        \quad \astikznodetransparentlycircled{x0}{4}\!,
        \quad \astikznodecircled{t1}{mypurple}{3}\!,
        \quad \astikznodetransparentlycircled{x1}{1}\!,
        \quad \astikznodecircled{t2}{mypurple}{8}\!,
        \quad \astikznodetransparentlycircled{x2}{2}\!,
        \quad \ldots \]
      {\centering\begin{tikzpicture}[remember picture,overlay]
        \node[draw=mypurple, circle, thick, inner sep=0.1em] (t3) {\scriptsize$\leq$};
        \path[draw=mypurple,thick]
          (t1)
          to [out=-90, in=180] (t3)
          to [out=0, in=-90] (t2);
      \end{tikzpicture}\par}}
    \end{column}

    \pause

    \begin{column}{0.50\textwidth}
      \begin{block}{Key stability results}
        \justifying
        \only<1-5>{Assuming~\badbox{\textsc{lem}}
        and~\badbox{\textsc{dc}}}\only<6->{\good{Constructively}}, \ldots

        \hil{Dickson:} \tabto{1.7cm} If~$X$ and~$Y$ are well\only<4-5>{\sinf}\only<6->{\ind}, so is~$X \times Y$. \\
        \hil{Higman:}  \tabto{1.7cm} If~$X$ is well\only<4-5>{\sinf}\only<6->{\ind}, so is~$\mathrm{List}\,X$. \\
        \hil{Kruskal:} \tabto{1.7cm} If~$X$ is well\only<4-5>{\sinf}\only<6->{\ind}, so is~$\mathrm{Tree}\,X$.
      \end{block}
    \end{column}
  \end{columns}
  \pause
  \pause
  \pause
  \pause

  \textbf{Def.} A quasi-order~$X$ is \hil{well\ind} iff~$G\,[\,]$, where $G$ is
  the following inductively defined predicate on \hil{finite lists}.~\formalized{}
  (In presence of \bad{bar induction}, $\text{well}\ind \Leftrightarrow
  \text{well}\sinf$.)
  \[
    \small
    \infer{\mathsf{now}\,p : G\,\sigma}{p : \mathsf{Good}\,\sigma}
    \qquad
    \infer{\mathsf{later}\,f : G\,\sigma}{f : (x : X) \to G\,(\sigma \dblcolon^r x)}
  \]
  \vspace*{-1.3em}

  \only<8>{
    \begin{columns}[c]
      \begin{column}{0.01\textwidth}
	\includegraphics[height=2.4em]{question-mark}
      \end{column}
      \begin{column}{0.9\textwidth}
      Is there a procedure for reinterpreting \hil{classical proofs}
      regarding well\sinf{} as \newline \hil{blueprints for constructive proofs} regarding well\ind?
      \end{column}
    \end{columns}
  }
\end{frame}}

\begin{frame}{Missing functions in the type of all functions?}
  \jnote{1}{
    The double negation translation technique fails to constructivize the two
    proofs presented on the slide.

    The reason is that, after applying the translation, the ``$I$-extracted
    subsequence'' appearing in the proof on the left will no longer actually be
    a function~$\NN \to X$. Instead, it will be a partially defined function
    with the property that each number~$n$ does \emph{not~not} belong to the
    domain. However, the assumption that~$X$ is well\sinf{} does not apply to
    these kinds of generalized sequences.

    In contrast, the notion well\ind{} is much stronger. If~$X$ is well\ind,
    then not only will every actual sequence~$\NN \to X$ be good, but (in a
    suitable sense) so will every up-to-$\neg\neg$-partially-defined sequence,
    every multivalued sequence and many more kinds of generalized sequences.
  }

  \begin{columns}[t]
    \begin{column}{0.57\textwidth}
      \begin{block}{Behold: A transfinite tool \ldots\phantom{g}}
        \justifying
        \textbf{Lemma.} \badbox{\textsc{lem}} Let~$X$ be well\sinf.
        Let~$\alpha : \NN \to X$. Then there is an increasing
        subsequence~$\alpha\,i_0 \leq \alpha\,i_1 \leq \cdots$.
      \end{block}

      \small
      \emph{Proof.}
      \begin{changemargin}{-0.8em}{0em}\begin{enumerate}\justifying
      \item The type
      $I \defeq \sum_{i \? \NN} \neg\sum_{j \? \NN} i < j \times \alpha\,i \leq
      \alpha\,j$
      is not in bijection with~$\NN$, as else the~$I$-extracted subsequence
      of~$\alpha$ would not be good.\smallskip
      \item By~\badbox{\textsc{lem}}, the type~$I$ is finite.\smallskip
      \item Any index~$i_0$ larger than all the indices in~$I$ is a suitable
      starting point for an increasing subsequence. \qed
      %  \item Let $\mathsf{Sad}\,i \defeq
      %  \neg \mathsf{Happy}\,i$ be the proposition
      %  expressing that~$i$ does not appear as the first component
      %  of a good pair, where
      %  $\mathsf{Happy}\,i \defeq \sum_{j \? \NN} i < j \times \alpha\,i \leq
      %  \alpha\,j$.
      %  \item If~$\mathsf{Sad}$ is anonymously dense in~$\NN$, i.e. $\forall a
      %  \? \NN\_ \neg\neg\exists b\?\NN\_ b \geq a \times \mathsf{Sad}\,b$, then
      %  \badbox{\textsc{lem}} and \badbox{\textsc{dc}} provide us with a
      %  subsequence~$\alpha\,k_0 \leq \alpha\,k_1 \leq \ldots$
      %  with~$\mathsf{Sad}\,k_0, \mathsf{Sad}\,k_1, \ldots$. This is a contradiction to wellness.
      %  \item
      %  Hence there is~\bad{not~not} an index~$a \? \NN$ such that~$\mathsf{Sad}$ is
      %  strictly bounded by~$a$; hence every index~$b \geq a$ does~\bad{not not} have
      %  a partner: $\forall b\geq a\_ \neg\neg\exists c>b\_
      %  \alpha\,b \leq \alpha\,c$. Thus~\bad{\textsc{lem}} and \bad{\textsc{dc}}
      %  provide us with the desired subsequence. \qed
      \end{enumerate}\end{changemargin}
    \end{column}

    \begin{column}{0.43\textwidth}
      \begin{block}{\ldots{} implying a concrete consequence}
        \textbf{Cor.} \badbox{\textsc{lem}}
        Let~$X$ and~$Y$ be well\sinf. Then~$X \times Y$ is
        well\sinf.\phantom{g}
      \end{block}

      \small
      \emph{Proof.}
      \begin{changemargin}{-0.8em}{0em}\begin{enumerate}\justifying
      \item Let a sequence~$\gamma : \NN \to X \times Y$ be
      given. Write~$\gamma\,k = (\alpha\,k,\beta\,k)$.\smallskip
      \item By the lemma, there is an
      increasing subsequence~$\alpha\,i_0 \leq \alpha\,i_1 \leq \cdots$.\smallskip
      \item Because~$Y$
      is well, there are indices~$n < m$ such that~$\beta\,i_n \leq
      \beta\,i_m$.\smallskip
      \item As
      also~$\alpha\,i_n \leq \alpha\,i_m$, the sequence~$\gamma$ is good. \qed
      \end{enumerate}\end{changemargin}
    \end{column}
  \end{columns}

  \dnote{\emph{We cannot trust~\bad{\textsc{lem}}-provided sequences to be
  available in the type~$\NN \to X$. Similarly with~\bad{\textsc{dc}}.}}
\end{frame}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{0.29cm}\hspace*{-1cm}\includegraphics[height=\paperheight]{tower-cropped}\end{minipage}}
\begin{frame}{Where do many cherished inductive definitions come from?}
  \begin{changemargin}{8.0em}{-0.5em}
    In mathematics, we routinely enlarge structures:
    \begin{itemize}
      \small
      \item Pass from~$\mathbb{Q}$ to~$\mathbb{R}$, to embrace irrationals.
      \item Pass from~$\mathbb{R}$ to~$\mathbb{C}$, to obtain $\sqrt{-1}$.
      \item Pass from~$\mathbb{C}$ to~$\mathbb{C}[X]$, to obtain a ``generic number''.
    \end{itemize}

    In set and type theory, we can also enlarge \hil{the mathematical universe}:
    \begin{itemize}
      \small
      \item Force a \hil{generic sequence} $\NN \to X$.
            \fixedhref{https://lets-play-agda.quasicoherent.io/Padova2025.Explorations.Forcing.Cohen.html}{\formalized}
      \item Force a \hil{generic enumeration} $\NN \twoheadrightarrow X$
            (even if $X$ is uncountable).
            \fixedhref{https://lets-play-agda.quasicoherent.io/Padova2025.Explorations.Forcing.Levy.html}{\formalized}
      \item Force a \hil{generic prime ideal} of a given ring.
            \fixedhref{https://iblech.github.io/constructive-maximal-ideals/Forcing.Prime.html}{\formalized}
    \end{itemize}
    \pause

    Central observations of the multiversal yoga:
    \begin{itemize}
      \small
      \item A quasi-order~$X$ is well\ind iff the generic sequence~$\NN \to X$
      is good.
      \item A set is Noetherian iff the generic sequence~$\NN \to X$
      has two repeated terms.
      \item A relation is well-founded iff for the generic descending chain,
      $\bot$.
      \item A ring element is nilpotent iff it is contained in the generic
      prime ideal.
      \fixedhref{https://iblech.github.io/constructive-maximal-ideals/Forcing.Prime.html\#corollary}{\formalized}
    \end{itemize}
    \pause

    The mystery of nongeometric sequents:
    \begin{itemize}
    \small
      \item The generic ring is a field.
      \item \ \\[-1.2em]\mbox{For the generic surjection $f : \NN \twoheadrightarrow \RR$,
      $\neg\neg \exists n \? \NN\_ f(n) = \pi \wedge f(n+1) = e$.
      \fixedhref{https://lets-play-agda.quasicoherent.io/Padova2025.Explorations.Forcing.Levy.html}{\formalized}}
    \end{itemize}
  \end{changemargin}

  \jnote{1}{
    The technique for enlarging the mathematical universe is called
    \emph{forcing} and was originally pioneered by Paul Cohen. Set theorists
    use forcing to construct new models of set theory from given ones, in order
    to explore the range of set-theoretic possibility. For instance, by forcing
    we can construct models of~\textsc{zfc} validating the continuum hypothesis
    and also models which falsify it.

    We here refer to a simplification of original forcing which is useful in a
    constructive metatheory. At its core, every forcing extension is just a
    formula and proof translation of a certain form. For instance, there is a
    forcing extension validating~\bad{\textsc{lem}} even if the base universe
    does not; this forcing extension is not a deep mystery, for a statement holds
    in that forcing extension iff its double negation translation holds in the
    base universe and it is well-known that the double negation translation
    of every instance of~\bad{\textsc{lem}} is an intuitionistic tautology.

    \href{https://www.speicherleck.de/iblech/stuff/slides-herrsching2023.pdf}{Here
    is a set of slides on constructive forcing}, and Section~4 of
    \href{https://raw.githubusercontent.com/iblech/constructive-maximal-ideals/master/tex/extended.pdf}{this
    joint paper with Peter Schuster} contains a written summary of constructive
    forcing.
  }

  \jnote{2}{
    Forcing the existence of a new function~$\NN \to X$ is also called
    \emph{Cohen forcing}. Forcing the existence of a new surjection is called
    \emph{Lévy collapse}. Forcing the existence of a prime ideal does not have
    an established name, perhaps it should be called \emph{Zariski forcing}; it
    has been explored by Thierry Coquand and others of the constructive algebra
    crew.

    The multiversal yoga allows us to reason with the good notion well\ind{}
    using the simple language of well\sinf, and similarly with several other
    notions.
  }

  \jnote{3}{
    For more on the mystery of nongeometric sequents, see
    \href{https://www.speicherleck.de/iblech/stuff/slides-birmingham2019.pdf}{this
    set of slides}.
  }
\end{frame}}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{4.95cm}\includegraphics[width=\paperwidth]{topos-horses-lighter}\end{minipage}}
\begin{frame}{A modal language for harnessing the multiverse}
  \jnote{1}{
    By \emph{topos}, we mean \emph{Grothendieck topos}. In constructive
    forcing, a ``forcing extension of the base universe'' is exactly the same
    thing as a Grothendieck topos.

    A particular member of the rich and varied landscape of toposes is the
    \emph{trivial topos}, in which every statement whatsoever holds. By
    restricting to positive toposes, we exclude this special case.

    For positive toposes~$\E$, a geometric implication holds in~$\E$ iff it
    holds in the base universe. For positive overt toposes~$\E$, we even have
    that a bounded first-order formula holds in~$\E$ iff it holds in the base.
    Hence, for the purpose of verifying a bounded first-order assertion about
    the base, we can freely pass to a positive overt topos with problem-adapted
    better higher-order properties (such as that some uncountable set from the
    base now appears countable, or that an infinite sequence whose existence is
    predicted by failing dependent choice now actually exists).

    \href{https://www.speicherleck.de/iblech/stuff/early-draft-modal-multiverse.pdf}{Here
    is a rough early draft of a preprint with more details about the modal
    multiverse.}
  }

  \jnote{2}{
    The idea to study the modal multiverse of toposes in a principled manner
    was proposed by Alexander Oldenziel in 2016. \emph{Foreshadowed by:}
    \begin{itemize}
      \item[1984\phantom{s}] André Joyal, Miles Tierney. ``An extension of the Galois theory of Grothendieck''.
      \\[-1em]
      \item[1987\phantom{s}] Andreas Blass. ``Well-ordering and induction in intuitionistic logic and topoi''.
      \\[-1em]
      \item[2010s] Milly Maietti, Steve Vickers. Ongoing work on arithmetic universes.
      \\[-1em]
      \item[2011\phantom{s}] Joel David Hamkins. ``The set-theoretic multiverse''.
      \\[-1em]
      \item[2013\phantom{s}] Shawn Henry. ``Classifying topoi and preservation of higher order logic by geometric morphisms''.
    \end{itemize}
  }

  \jnote{3}{
    With the modal language we seek to provide an accessible and modular
    framework for constructivization results.

    For instance, conservativity of classical logic over intuitionistic logic
    for geometric implications (known under various names such as Barr's
    theorem, Friedman's trick, escaping the continuation monad, \ldots) is
    packaged up by the observation that \emph{somewhere}, the law of excluded
    middle holds.

    Another example:
    In the community around Krull's lemma, it is well-known that
    we can constructively infer that a given ring element~$x \in A$ is nilpotent from
    knowing that it is contained in the \emph{generic prime ideal} of~$A$. This entity
    is not actually an honest prime ideal of the ring~$A$ in the base
    universe, but a certain combinatorial notion (efficiently dealt with using
    \emph{entailment relations}). Constructive forcing allows us to reify the
    generic prime ideal as an actual prime ideal in a suitable forcing extension,
    so in a suitable topos (a version of the little Zariski topos of the ring).
  }

  \textbf{Def.} A statement~$\varphi$ holds \ldots
  \vspace*{-0.3em}
  \begin{itemize}
    \small
    \item \hil{everywhere} \tabto{2.05cm}($\necessary\varphi$)\tabto{2.98cm}
    iff it holds \hil{in every topos}
    (over the current base).
    \item \hil{somewhere} \tabto{2.05cm}($\possible\varphi$)\tabto{2.98cm} iff
    it holds \hil{in some positive topos}.
    \item \hil{proximally} \tabto{2.05cm}($\xpossible\varphi$)\tabto{2.98cm}
    iff it holds \hil{in some positive overt topos}.
  \end{itemize}

  We then have:
  \begin{enumerate}
    \item $\mathsf{Well}\ind(X,{\le}) \Longleftrightarrow
    \necessary\mathsf{Well}\sinf(X,{\le})$.
    \item $(\possible\varphi) \Longleftrightarrow \varphi$,
    if $\varphi$ is a geometric implication (``$\forall\ldots\forall (\%
    \Rightarrow \%)$'', with no~$\forall$ nor~$\Rightarrow$
    in~$\%$).~\fixedhref{https://iblech.github.io/constructive-maximal-ideals/Forcing.Semantics.html\#CoherentEquivalence}{\formalized}
    \item $(\xpossible\varphi) \Longleftrightarrow \varphi$,
    if~$\varphi$ is a bounded first-order statement.
    \fixedhref{https://iblech.github.io/constructive-maximal-ideals/Forcing.Semantics.html\#1\%CB\%A2\%E1\%B5\%97OrderEquivalence}{\formalized}
    \item For every inhabited type~$X$, $\xpossible\necessary
    \mathsf{Countable}(X)$, where~$\mathsf{Countable}(X) \defeqv (\exists f : \NN
    \to X\_ \mathsf{Surjective}(f))$.
    \fixedhref{https://lets-play-agda.quasicoherent.io/Padova2025.Explorations.Forcing.Levy.html}{\formalized}

    So being countable is a \hil{button}.
    \item $\possible \badbox{\textsc{lem}}$.
    (Baby Barr / Friedman's trick / nontrivial exit continuation)

    In fact, \badbox{\textsc{lem}} is a \hil{switch}:
    $\necessary\bigl((\possible\badbox{\textsc{lem}}) \wedge
    (\possible\neg\badbox{\textsc{lem}})\bigr)$

    \item $\badbox{\textsc{zorn}} \Rightarrow
    \necessary\possible\badbox{\textsc{ac}}$. (Great Barr)
  \end{enumerate}
\end{frame}}

\begin{frame}{The plain combinatorics of toposes}
  \jnote{3-}{
    Proofs (about the base universe) are (pure) programs; proofs about a forcing
    extension are effectful programs unfolding in the~$\nabla$ monad.

    The $\nabla$ monad is a refinement of the (monotonic) state monad.
  }

  \begin{enumerate}\justifying
    \item Realize a generic gadget as some kind of limit of \hil{approximations}
    from the base universe. For the generic function $f_0 : \NN \to X$:

    \AgdaHide{
      \begin{code}
open import Data.List
open import Agda.Primitive using (Set)

Prop = Set
postulate
  X : Set
  _∈_ : X → List X → Prop
      \end{code}
    }

    {\small\begin{code}
L : Set
L = List X
    \end{code}}

    \pause
    \item \hil{Reinterpret}, in a mechanical fashion, assertions about the generic
    gadget as assertions about its
    approximations~(\fixedhref{https://iblech.github.io/constructive-maximal-ideals/Forcing.Semantics.html}{\formalized}).

    {\small\begin{itemize}
    \item We have the \hil{stage-dependent proposition} ``$f_0\,n = x$'',
    a certain function $L \to \mathsf{Prop}$: $\lambda \sigma.\
    (\mathsf{lookupMaybe}\,\sigma\,n = \mathsf{just}\,x)$\end{itemize}}
    \medskip

    \pause
    \item Be prepared to \hil{evolve approximations}
    (\fixedhref{https://iblech.github.io/constructive-maximal-ideals/Forcing.Base.html}{\formalized}).

    {\small\begin{code}
data ∇ (P : L → Prop) : L → Prop where
  now   : {σ : L} → P σ → ∇ P σ
  later : {σ : L} → ((x : X) → ∇ P (σ ∷ʳ x)) → ∇ P σ
    \end{code}

    \begin{itemize}
    \item
    For a stage-dependent proposition~$P : L \to \mathsf{Prop}$, $\nabla\,P\,\sigma$ expresses
    that no matter how~$\sigma$ will evolve over a time to a better
    approximation~$\tau$, eventually~$P\,\tau$ will hold.
  % later₂ : (a : X) → ((τ : X*) → a ∈ σ ⧺ τ → ∇ P (σ ⧺ τ)) → ∇ P σ
    \item That~$f_0$ is defined on input~$n$ can be expressed
    as~$\nabla\,P\,[\,]$ where~$P\,\sigma \defeqv (\mathsf{length}\,\sigma > n)$.\end{itemize}}
    \medskip
    \pause

    \item Crucially, this interpretation is \hil{sound} with respect to constructive
    reasoning.
  \end{enumerate}
\end{frame}

\begin{frame}{Formal metatheory}
  \jnote{2}{
    \href{https://msp.cis.strath.ac.uk/types2025/slides/TYPES2025-slides57.pdf}{Exciting
    ongoing work on split type theory} by Martin Baillon, Assia Mahboubi and
    Pierre-Marie Pédrot!
  }

  \small
  \good{\cmark{}} \emph{There are type-theoretic multiverses,} such as
  \begin{itemize}
    \item the collection of all $\mathrm{PSh}(\mathcal{C} \times \mathcal{B})$,
    where~$\mathcal{B}$ ranges over cube categories and~$\mathcal{C}$ over
    arbitrary small categories, and their corresponding sheaf models

    {\scriptsize
    Coquand. ``A survey of constructive presheaf models of univalence''. \emph{ACM SIGLOG News}, 5.3 (2018).
    \par}
  \end{itemize}
  \medskip
  \pause

  \bad{\xmark} \emph{Accessing the multiverse from within intensional type theory is tricky:}
  \begin{itemize}
    \item Given a model of $\mathfrak{s}$CIC and a category~$\mathcal{C}$ in
    it, we have a syntactic presheaf model of CIC.

    {\scriptsize
    Coquand, Jaber. ``A note on forcing and type theory''. \emph{Fundamenta Informaticae 100} (2010). \\
    Jaber, Lewertowski, Pédrot, Sozeau, Tabareau. ``The definitional side of the forcing''. \emph{Proceedings of LICS '16} (2016). \\
    Pédrot. ``Russian constructivism in a prefascist theory''. \emph{Proceedings of LICS '20} (2020).
    \par}
    \item Given a suitable lex modality, we have a syntactic sheaf model (model of modal types).

    {\scriptsize
    Coquand, Ruch, Sattler. ``Constructive sheaf models of type theory.'' \emph{Math. Struct. Comput. Sci.} 31.9 (2021). \\
    Escardó, Xu. ``Sheaf models of type theory in type theory''. Unpublished (2016). \\
    Quirin. ``Lawvere--Tierney sheafification in Homotopy Type Theory''. PhD thesis (2016).
    \par}
    \item (I believe) we have syntactic sheaf models in certain special cases,
    when no coherence issues arise in defining the notion of presheaves.
  \end{itemize}

  Note: We can use~$\nabla$ even without a proper metatheoretic backing.
\end{frame}

{\renewcommand{\insertframeextra}{a}
\begin{frame}{Increasing subsequences as convenient fictions}
  \jnote{1}{
    On this slide~$\nabla$ is the monad described on slide 5,
    and~$\nabla^\nnearrow$ is the variant where lists are only allowed to grow
    in a monotonically increasing manner; so~$\nabla$ is the monad for the
    generic sequence~$\NN \to X$, while~$\nabla^\nnearrow$ is the monad for the
    generic increasing sequence~$\NN \to X$.

    The displayed multiversal proof gives a positive answer to a question posed
    by Stefano Berardi, Gabriele Buriola and Peter Schuster, see
    \href{https://www.speicherleck.de/iblech/stuff/slides-abmv2024.pdf\#page=16}{this
    set of slides}.

    At the end of the day, this proof yields a concrete algorithm (see
    \href{https://www.speicherleck.de/iblech/stuff/dickson/implicational.pl}{here}
    for manually-extracted Perl code). Given an infinite sequence, this
    algorithm seeks, in a certain systematic but not exhaustive manner, a
    sufficiently long monotonic subsequence, making judicious use of
    backtracking. This search terminates, though this fact is not obvious just
    by looking at the algorithm.
  }

  \justifying
  Let~$X$ be a quasi-order.
  Let~$B : \mathsf{List}\,X \to \mathsf{Prop}$ be a monotone predicate.

  \begin{block}{Classical blueprint}
    \textbf{Thm.} \badbox{\textsc{lem}} If~$X$ is well\sinf{} and
    if every increasing sequence~$\alpha : \NN \to X$ has a prefix validating~$B$,
    then every sequence has a prefix validating~$B$.
  \end{block}
  \vspace*{-0.4em}
  {\small\emph{Proof.} Let~$\alpha : \NN \to X$ be a sequence. By the lemma,
  there is an increasing subsequence~$\alpha\,i_0 \leq \alpha\,i_1 \leq \cdots$.
  By assumption, this subsequence has a prefix validating~$B$. This prefix is part
  of a prefix of the original sequence~$\alpha$. Hence we conclude by
  monotonicity. \qed\par}
  \medskip

  \begin{block}{Constructive reimagination}
    \textbf{Thm.} If~$X$ is well\ind and if~$\nabla^\nnearrow\,B\,[\,]$,
    then~$\nabla\,B\,[\,]$.
    \fixedhref{https://www.speicherleck.de/iblech/stuff/dickson//Dickson.html}{\formalized}
  \end{block}
  \vspace*{-0.4em}
  {\small\emph{Proof.} Let~$\alpha : \NN \to X$ be a sequence (in an arbitrary topos).
  \emph{Somewhere,} \textsc{lem} holds. \emph{There}~$X$
  is still well\sinf, so that we have an increasing subsequence~$\alpha\,i_0 \leq \alpha\,i_1 \leq \cdots$.
  By assumption, this subsequence has a finite prefix validating~$B$. This
  prefix is part of a prefix of the original sequence~$\alpha$. Hence~$\alpha$
  has a prefix validating~$B$ by monotonicity \emph{there}. So
  \emph{somewhere} there is a finite prefix validating~$B$. Thus there
  actually is a finite prefix validating~$B$. \qed\par}
\end{frame}}

\addtocounter{framenumber}{-1}
{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{5.95cm}\includegraphics[width=\paperwidth]{fr1}\end{minipage}}
\renewcommand{\insertframeextra}{b}
\begin{frame}{Maximal ideals as convenient fictions}
  \jnote{1-}{
    The displayed classical proof is quite efficient from the point of view of
    organizing mathematical knowledge, as it quickly reduces the general
    situation of dealing with an arbitrary ring to dealing with a field.
    Alas, read literally, it is hopeless ineffective.
  }

  \jnote{2-}{
    By employing modal language, we can closely mimic the original proof and be
    fully constructive at the same time.
  }

  \jnote{3-}{
    By unwinding all modal definitions, the modal proof can be unrolled to a
    fully explicit computation.
  }

  \justifying
  Let~$A$ be a commutative ring with unit.

  \textbf{Thm.}
  Let~$M$ be a surjective matrix with more rows than columns over~$A$. Then~$1 = 0$ in~$A$.
  \medskip

  \emph{Classical proof.}
  \bad{Assume not.} Then there is~a \bad{maximal ideal} $\mmm$.
  The matrix~$M$ is surjective over~$A/\mmm$. Since~$A/\mmm$ is a field, this
  is a contradiction to basic linear algebra. \qed
  \medskip

  \only<1>{\medskip\par\centering\scalebox{0.9}{\centering\begin{tikzpicture}
    \node (0) at (0,1) {$(0) = \{0\}$};
    \node (1) at (0,5) {$(1) = \ZZ$};
    \node (2) at (-2,4) {$(2)$};
    \node [right of=2] (3) {$(3)$};
    \node [below of=2] (4) {$(4)$};
    \node [below of=2, xshift=0.7cm] (6) {$(6)$};
    \node [right of=3] (5) {$(5)$};
    \node [right of=5] (7) {$(7)$};
    \node [right of=7] (7d) {$\ldots$\phantom{(}};
    \node [right of=7d, xshift=3cm, yshift=-2cm] (max)
    {\vbox{\small{\it maximal among the proper ideals} \\ \medskip \hspace*{-6.75em}\textbullet \quad $\neg(1 \in
    \mmm)$ \\ \medskip \textbullet \quad $\neg\bigl(1 \in \mmm + (x)\bigr) \Rightarrow x \in \mmm$}};
    \node [below of=4] (8) {$(8)$};
    \node [right of=8, xshift=3cm] (8d) {$\ldots$};
    \draw (0) -- (8);
    \draw (0) -- (8d);
    \draw (0) -- (6);
    \draw (2) -- (1);
    \draw (3) -- (1);
    \draw (5) -- (1);
    \draw (7) -- (1);
    \draw (7d) -- (1);
    \draw (4) -- (2);
    \draw (8) -- (4);
    \draw (6) -- (2);
    \draw (6) -- (3);
    \draw [mypurple!30, thick, shorten <=-2pt, shorten >=-2pt, ->] (max) to [out=120, in=-30] (7d);
    \begin{pgfonlayer}{background}
      \draw[decorate, very thick, draw=mypurple!30]
        ($(2.south west) + (8pt, 0)$) arc(270:180:8pt) --
        ($(2.north west) + (0, -8pt)$) arc(180:90:8pt) --
        ($(7d.north east) + (-8pt, 0)$) arc(90:0:8pt) --
        ($(7d.south east) + (0, 8pt)$) arc(0:-90:8pt) --
        cycle;
    \end{pgfonlayer}
  \end{tikzpicture}\par}\par}
  \pause

  \emph{Multiversal constructive proof.}
  We may work \emph{somewhere} where \good{\textsc{lem}} holds. So \good{assume not}.
  \emph{Proximally,} there is a \good{maximal
  ideal}~$\mmm$~(\fixedhref{https://iblech.github.io/constructive-maximal-ideals/index.html\#§4.§4-4}{\formalized}).
  The matrix~$M$ is still surjective \emph{there}, and also over~$A/\mmm$. Since~$A/\mmm$ is a field, this
  is a contradiction to basic linear algebra. \qed
  \medskip
  \pause

  \emph{Unrolled constructive proof (special case,
  \fixedhref{https://iblech.github.io/constructive-maximal-ideals/index.html\#§4.TestCase}{\formalized}).}
  Write~$M =
  \left(\begin{smallmatrix}x\\y\end{smallmatrix}\right)$. By surjectivity,
  have~$u, v$ with
  \vspace*{-0.6em}
  \[
    u \left(\begin{smallmatrix}x\\y\end{smallmatrix}\right) = \left(\begin{smallmatrix}1\\0\end{smallmatrix}\right)
    \quad\text{and}\quad
    v \left(\begin{smallmatrix}x\\y\end{smallmatrix}\right) = \left(\begin{smallmatrix}0\\1\end{smallmatrix}\right).
  \]
  Hence
  $
    1 = (vy) (ux) = (uy) (vx) = 0
  $.
  \qed
\end{frame}}

\addtocounter{framenumber}{-1}
{\renewcommand{\insertframeextra}{c}
\begin{frame}{Stabilization as convenient fiction}
  A commutative ring~$A$ with unit is \ldots
  \begin{itemize}
    \item Noetherian\sinf{} iff for every sequence~$x_0,x_1,\ldots$ of ring elements,
    there is a number~$n : \NN$ such that~$x_n,x_{n+1},x_{n+2},\ldots \in
    (x_0,\ldots,x_{n-1})$.
    \item Noetherian\ensuremath{_\text{RS}} iff for every sequence~$x_0,x_1,\ldots$ of ring elements,
    there is a number~$n : \NN$ such that~$x_n \in
    (x_0,\ldots,x_{n-1})$.
    \item Noetherian\ind{} iff a certain inductively defined condition holds
    \fixedhref{https://dl.acm.org/doi/10.5555/646538.696017}{[Coquand–Persson]}.
  \end{itemize}

  We then have:
  \begin{align*}
    \mathsf{Noetherian\ind}(A) &\Longleftrightarrow
    \necessary\mathsf{Noetherian}_\text{RS}(A) \\
    &\Longleftrightarrow
    \necessary\bigl(\badbox{\textsc{lem}} \Rightarrow
    \mathsf{Noetherian}\sinf(A)\bigr) \\
    &\Longleftrightarrow
    \necessary\possible\mathsf{Noetherian}\sinf(A).
  \end{align*}

  With this observation, from their classical counterparts, we can extract
  constructive proofs of:
  \begin{enumerate}
    \item Hilbert's basis theorem: If $A$ is Noetherian\ind, then so is $A[X]$.
    \item Locality: If~$1 = f_1+\ldots+f_n$ and if each
    ring~$A[f_i^{-1}]$ is Noetherian\ind, then so is~$A$. \\
    (As a consequence, $A$ is Noetherian\ind
    iff algebraic geometry's~$\mathcal{O}_{\mathrm{Spec}(A)}$ is.)
  \end{enumerate}
\end{frame}

\backupstart

\begin{frame}[plain]
  \jnote{1}{
    \href{https://lets-play-agda.quasicoherent.io/}{Let's play Agda}, an
    interactive course on Agda, complementing the
    \href{https://agdapad.quasicoherent.io/}{Agdapad}.

    \href{https://www.uni-augsburg.de/en/fakultaet/mntf/math/prof/geom/workshop-aimath/}{AI
    Transforms Maths Research}, a workshop in Augsburg.

    \href{https://conferences.cirm-math.fr/3377.html}{Synthetic mathematics}, a
    conference in Luminy.
  }

  \begin{columns}[c]
    \begin{column}{0.45\textwidth}
      \includegraphics[width=0.95\textwidth]{lets-play-agda}
    \end{column}
    \begin{column}{0.5\textwidth}
      \includegraphics[width=0.8\textwidth]{ai-2025}

      \vspace*{-1.5cm}\hspace*{2cm}
      \includegraphics[width=0.8\textwidth]{cirm-2025}

      \vspace*{1cm}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \emph{Backup slides.}
\end{frame}

\section{Basics of forcing}

\begin{frame}{Ingredients for forcing}
  \vspace*{-0.3em}
  To construct a forcing extension, we require:
  \begin{enumerate}
    \item a base universe~$V$
    \item a preorder~$L$ of \hil{forcing conditions} in~$V$\!,
    pictured as \hil{finite approximations}

    (\emph{convention:} $\tau \preccurlyeq \sigma$ means that~$\tau$ is a
    better finite approximation than~$\sigma$)
    \item a \hil{covering system} governing how finite approximations evolve to
    better ones

    (for each~$\sigma \in L$, a set~$\Cov(\sigma) \subseteq
    P({\downarrow}\sigma)$, with a simulation condition)
  \end{enumerate}
  In the forcing extension~$V^\nabla$, there will then be a \hil{generic filter} (ideal
  object).
  \pause

  \vspace*{-1em}
  \begin{columns}
    \begin{column}{0.49\textwidth}
      \small
      \begin{block}{For the generic surjection~$\NN \twoheadrightarrow X$}
        \justifying
        Use \hil{finite lists}~$\sigma \in X^*$ as forcing conditions,
        where $\tau \preccurlyeq \sigma$ iff~$\sigma$ is an initial segment of~$\tau$,
        and be prepared to grow~$\sigma$ to \ldots
        \footnotesize
        \begin{enumerate}
          \item[(a)] one of~$\{ \sigma x \,|\, x \in X \}$, to make~$\sigma$ more defined
          \item[(b)] one of~$\{ \sigma \tau \,|\, \tau \in X^*, a \in
          \sigma\tau \}$, for any~$a \in X$, to make~$\sigma$ more surjective
        \end{enumerate}
        \vspace*{-0.4em}
      \end{block}
    \end{column}
    \pause

    \begin{column}{0.45\textwidth}
      \small
      \begin{block}{For the generic prime ideal of a ring~$A$}
        \justifying
        Use \hil{f.g.\@ ideals} as forcing conditions, where $\bbb \preccurlyeq
        \aaa$ iff~$\bbb \supseteq \aaa$, and be prepared to grow~$\aaa$ to \ldots
        \footnotesize
        \begin{enumerate}
          \item[(a)] one of~$\emptyset$, if~$1 \in \aaa$, to make~$\aaa$ more proper
          \item[(b)] one of~$\{ \aaa+(x), \aaa+(y) \}$, if~$xy \in \aaa$, to
          make~$\aaa$ more prime
        \end{enumerate}
        \vspace*{-0.4em}
      \end{block}
    \end{column}
  \end{columns}
\end{frame}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{-1cm}\includegraphics[width=\paperwidth]{forest-light-colored}\end{minipage}}
\begin{frame}{The eventually monad}
  Let~$L$ be a forcing notion.

  Let~$P$ be a monotone predicate on~$L$
  (if $\tau \preccurlyeq \sigma$, then $P\sigma \Rightarrow P\tau$). \\
  For instance, in the case~$L = X^*$:
  \begin{itemize}
    \item $\mathsf{Repeats}\, x_0\ldots x_{n-1} \defeqv \exists i\_ \exists j\_ i < j \wedge x_i = x_j$
    \item $\mathsf{Good}\, \,\,\,\,\,\,x_0\ldots x_{n-1} \defeqv \exists i\_ \exists j\_ i < j \wedge x_i \leq x_j$
    \quad (for some preorder~$\leq$ on~$X$)
  \end{itemize}
  \pause

  We then define~``\hil{$\nabla\,P\,\sigma$}'' (``$P$ bars~$\sigma$'') inductively by the following clauses:
  \begin{enumerate}
    \item If~$P\sigma$, then~$\nabla\,P\,\sigma$.
    \item If~$\nabla\,P\,\tau$ for all~$\tau \in R$, where~$R$ is some covering
    of~$\sigma$, then~$\nabla\,P\,\sigma$.
  \end{enumerate}
  So~$\nabla\,P\,\sigma$ expresses in a \hil{direct inductive fashion}:
  \[ \text{``No matter
  how~$\sigma$ evolves to a better approximation~$\tau$, eventually~$P\,\tau$
  will hold.''} \]

  \pause
  We use quantifier-like notation: ``$\nabla(\tau \preccurlyeq \sigma)\_
  P\tau$'' means ``$\nabla\,P\,\sigma$''.
\end{frame}}
% BOARD:
% - examples for P | σ
% - abuse of notation

\begin{frame}{Proof translations}
  \textbf{Thm.} Every~\textsc{iqc}-proof remains correct, with at most a
  polynomial increase in length, if throughout we
  replace
  \[\begin{array}{rcl@{\quad\text{where}\quad}rcl}
    \exists & \leadsto & \exists^\mathrm{cl},
    & \exists^\mathrm{cl} &\defeqv& \neg\neg\exists, \\
    \vee & \leadsto & \vee^\mathrm{cl},
    & \alpha \vee^\mathrm{cl} \beta &\defeqv& \neg\neg(\alpha \vee \beta), \\
    = & \leadsto & =^\mathrm{cl},
    & s =^\mathrm{cl} t &\defeqv& \neg\neg(s = t).
  \end{array} \]
  \pause

  \begin{columns}[c]
    \begin{column}{0.01\textwidth}
      \includegraphics[height=2.4em]{sheafification-man-2}
    \end{column}
    \quad
    \begin{column}{0.9\textwidth}
      \hil{When we say:}\ \ some statement ``holds in~$V^{\neg\neg}$'', \\
      \makesamewidth[l]{\hil{When we say:}}{\hil{we mean:}}\ \ its translation holds in~$V$.
    \end{column}
  \end{columns}
  \bigskip

  Similarly for arbitrary forcing extensions~$V^\nabla$, ``just with~$\nabla$
  instead of~$\neg\neg$''.
  \bigskip

  \pause
  \textbf{Ex.} As~$\neg\neg(\varphi \vee \neg\varphi)$ is a theorem
  of~\textsc{iqc}, the law of excluded middle holds in~$V^{\neg\neg}$.
\end{frame}

\newcommand{\defeqvi}{\quad iff\quad}
\begin{frame}{The $\nabla$-translation}
  \small
  \only<1>{For bounded first-order formulas over the (large) first-order signature which has
  \begin{enumerate}
    \scriptsize
    \item one sort~$\underline{X}$ for each set~$X$ in the base universe,
    \\[-1.2em]
    \item one~$n$-ary function symbol~$\underline{f} : \underline{X_1} \times
    \cdots \times \underline{X_n} \to \underline{Y}$ for each map~$f : X_1 \times
    \cdots \times X_n \to Y$,
    \\[-1.2em]
    \item one~$n$-ary relation symbol~$\underline{R} \hookrightarrow
    \underline{X_1} \times \cdots \times \underline{X_n}$ for each relation~$R
    \subseteq X_1 \times \cdots \times X_n$, and
    \\[-1.2em]
    \item an additional unary relation symbol~$G \hookrightarrow \underline{L}$
    (for the \emph{generic filter} of~$L$),
  \end{enumerate}
  we recursively define:}
  \scriptsize
  \only<2->{\vspace*{-0.4em}}
  \begin{tabbing}
    \quad \= $\sigma \forces \forces \forall(x\?\underline{X})\_ \varphi$ \=
    \defeqvi $\textcolor{gray}{\forall(\tau \preccurlyeq \sigma)\_}\
    \forall(x_0 \in X)\_ \tau \forces \varphi[\underline{x_0}/x]$.\qquad\quad \=
    $\sigma \forces \exists(x\?\underline{X})\_ \varphi$
    \= $\sigma \forces \underline{R}(\underline{s_1},\ldots,\underline{s_n})$ \= \defeqvi $s = t$. \= \kill

    \> $\sigma \forces s = t$
    \> \defeqvi $\nabla \sigma\_ \llbracket s \rrbracket = \llbracket t \rrbracket$.
    \> $\sigma \forces \underline{R}(s_1,\ldots,s_n)$
    \> \defeqvi $\nabla\sigma\_ R(\llbracket s_1 \rrbracket,\ldots,\llbracket s_n \rrbracket)$. \\[0.3em]

    \> $\sigma \forces \varphi \Rightarrow \psi$
    \> \defeqvi $\textcolor{gray}{\forall(\tau \preccurlyeq \sigma)\_}\ (\tau \forces \varphi) \Rightarrow
    (\tau \forces \psi)$.
    \> $\sigma \forces G\tau$
    \> \defeqvi $\nabla\sigma\_ \sigma \preccurlyeq \llbracket\tau\rrbracket$. \\[0.3em]

    \> $\sigma \forces \top$ \> \defeqvi $\top$.
    \> $\sigma \forces \bot$ \> \defeqvi $\hil{$\nabla\sigma\_$}\ \bot$ \\[0.3em]

    \> $\sigma \forces \varphi \wedge \psi$
    \> \defeqvi $(\sigma \forces \varphi) \wedge (\sigma \forces \psi)$.
    \> $\sigma \forces \varphi \vee \psi$
    \> \defeqvi $\hil{$\nabla\sigma\_$}\ (\sigma \forces \varphi) \vee (\sigma \forces \psi)$. \\[0.3em]

    \> $\sigma \forces \forall(x\?\underline{X})\_ \varphi$
    \> \defeqvi $\textcolor{gray}{\forall(\tau \preccurlyeq \sigma)\_}\ \forall(x_0 \in X)\_ \tau \forces
    \varphi[\underline{x_0}/x]$.
    \> $\sigma \forces \exists(x\?\underline{X})\_ \varphi$
    \> \defeqvi $\hil{$\nabla\sigma\_$}\ \exists(x_0 \in X)\_ \sigma \forces \varphi[\underline{x_0}/x]$.
  \end{tabbing}
  \small
  \only<1>{Finally, we say that~$\varphi$ ``holds in~$V^\nabla$'' iff for all~$\sigma
  \in L$, $\sigma \forces \varphi$.}

  \footnotesize
  \begin{tabular}{@{}lp{0.27\textwidth}p{0.48\textwidth}@{}}
    \toprule
    forcing notion & statement about~$V^\nabla$ & external meaning \\
    \midrule
    surjection $\NN \twoheadrightarrow X$ &
    ``the gen.\@ surj.\@ is surjective'' &
    $\forall(\sigma{\in}X^*)\_ \forall(a{\in}X)\_ \nabla(\tau{\preccurlyeq}\sigma)\_ \exists(n{\in}\NN)\_ \tau[n] = a$. \\[0.4em]
    \pause

    map $\NN \to X$ &
    ``the gen.\@ sequence is good'' &
    $\mathsf{Good} \mid [\,]$. \\[0.4em]

    frame of opens &
    ``every complex number has a square root'' &
    For every open~$U \subseteq X$ and every cont.\@
    function $f : U \to \CC$, there is an open covering $U = \bigcup_i U_i$ such
    that for each index~$i$, there is a cont.\@ function $g : U_i \to \CC$
    such that~$g^2 = f$. \\[4.3em]

    big Zariski &
    ``$x \neq 0 \Rightarrow \text{$x$ inv.}$'' &
    If the only f.p.\@ $k$-algebra in which~$x = 0$ is the zero algebra,
    then~$x$ is invertible in~$k$.\\[1.5em]
    \pause

    little Zariski &
    ``every f.g. vector space does \emph{not~not} have a basis'' &
    \makesamewidth[l]{}{\phantom{x}}\hil{Grothendieck's generic freeness lemma}
  \end{tabular}
\end{frame}

\begin{frame}{Outlook}
  \begin{block}{Passing to and from extensions}
    \justifying\small
    \textbf{Thm.} Let~$\varphi$ be a \hil{bounded first-order formula} not
    mentioning~$G$. In each of the following situations, we have that
    $\varphi$ holds in~$V^\nabla$ iff~$\varphi$
    holds in $V$:

    \vspace*{-0.5em}
    \begin{enumerate}
      \item $L$ and all coverings are inhabited (proximality). \\[-1em]
      \item $L$ contains a top element, every covering of the
      top element is inhabited, and~$\varphi$ is a coherent implication
      (positivity).
    \end{enumerate}
    \vspace*{-0.8em}
  \end{block}

  \vspace*{-1.5em}
  \begin{columns}
    \begin{column}{0.46\textwidth}
      \begin{block}{The mystery of nongeometric sequents}
        \justifying
        The \hil{generic ideal} of a ring is maximal:
        \vspace*{-1em}
        \[ (x \in \aaa \Rightarrow 1 \in \aaa) \Longrightarrow 1 \in \aaa + (x). \]

        The \hil{generic ring} is a field:
        \vspace*{-0.7em}
        \[ (x = 0 \Rightarrow 1 = 0) \Longrightarrow (\exists y\_ xy = 1). \]
        \vspace*{-1.4em}
      \end{block}

    \end{column}

    \begin{column}{0.50\textwidth}
      \begin{block}{Traveling the multiverse \ldots}
        \textsc{lem} is a \hil{switch} and \hil{holds positively};
        being countable is a \hil{button}.
        \medskip

        Every instance of \textsc{dc} \hil{holds proximally}.
        \medskip

        A geometric implication is provable iff it holds \hil{everywhere}.
        \vspace*{-0.2em}
      \end{block}
      \vspace*{-0.3em}
      \hfill\footnotesize\ldots{} upwards, but always keeping ties to the
      base.{\ }
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{More on forcing notions}
  \small
  \textbf{Def.} A \hil{forcing notion} consists of
  a preorder~$L$ of \hil{forcing conditions}, and
  for every~$\sigma \in L$, a set~$\Cov(\sigma) \subseteq
  P({\downarrow}\sigma)$ of \hil{coverings} of~$\sigma$
  such that: If~$\tau \preccurlyeq \sigma$ and~$R \in \Cov(\sigma)$, there
  should be a covering~$S \in \Cov(\tau)$ such that~$S \subseteq
  {\downarrow}R$.
  \bigskip

  {\centering\footnotesize\begin{tabular}{llll}
    \toprule
    & preorder~$L$ & coverings of an element~$\sigma \in L$ & filters of~$L$ \\
    \midrule
    \normalnumber{1} & $X^*$ & $\{ \sigma x \,|\, x \in X \}$ & maps~$\NN \to X$ \\
    \normalnumber{2} & $X^*$ & $\{ \sigma x \,|\, x \in X \}$,\ \ $\{ \sigma\tau \,|\, \tau \in X^*, a \in \sigma\tau \}$ for each~$a \in X$ & surjections~$\NN \twoheadrightarrow X$ \\
    \normalnumber{3} & f.g. ideals & --- & ideals \\
    \normalnumber{4} & f.g. ideals & $\{ \sigma+(a), \sigma+(b) \}$ for each~$ab \in \sigma$,\ \ $\{\}$ if~$1 \in \sigma$ & prime ideals \\
    \normalnumber{5} & opens & $\mathcal{U}$ such that~$\sigma = \bigcup \mathcal{U}$ & points \\
    \normalnumber{6} & $\{\star\}$ & $\{ \star \,|\, \varphi \} \cup \{ \star \,|\, \neg\varphi \}$ &
    witnesses of~\textsc{lem}
    \\
    \bottomrule
  \end{tabular}\par}
  \bigskip

  \textbf{Def.} A \emph{filter} of a forcing notion~$(L,\mathrm{Cov})$
  is a subset~$F \subseteq L$ such that
  \vspace*{-0.4em}
  \begin{enumerate}
    \scriptsize
    \item $F$ is upward-closed: if~$\tau \preccurlyeq \sigma$ and if~$\tau \in F$, then~$\sigma \in F$; \\[-3.0em]
    \item $F$ is downward-directed: $F$ is inhabited, and if~$\alpha,\beta \in F$,
    then there is a common refinement~$\sigma \preccurlyeq \alpha,\beta$ such
    that~$\sigma \in F$; and \\[-2.0em]
    \item $F$ splits the covering system: if~$\sigma \in F$ and~$R \in
    \Cov(\sigma)$, then~$\tau \in F$ for some~$\tau \in R$.
  \end{enumerate}
\end{frame}

\backupend

\end{document}

0. A play in classical logic

1. Hilbert's program
   - basics about wqo's
   - subsequence lemma
   - Dickson
   - inductive reimagination
   - questions: where from? why work? how much stronger?

2. Where do our most cherished definitions come from?
   - a familiar business: enlarging mathematical structures
   - adjoining generic function, surjection, prime ideal, ...
   - observations:
     - well iff generic sequence is good iff everywhere, every sequence is good
       (in particular, every multivalued and every up-to-¬¬-defined sequence is good)
     - wellfounded iff for the generic decreasing sequence, ⊥ iff ...
   - fun facts about generic gadgets

3. Modal language for harnessing the multiverse
   - definition
   - examples
   - implicational wellness
   - dialogue

4. Implementation
   - idea: realize generic gadget as some kind of limit of approximations from
     the base
   - reinterpret, in a mechanic fashion, assertions about generic objects as
     assertions about approximations
     - ex.: "the generic function is defined on input n" as "no matter how the
       empty list evolves over a time to a better approximation xs, eventually
       xs will have length at least succ n"
   - crucially, this interpretation is sound with respect to intuitionistic
     reasoning

5. Let's play Agda, CIRM, AI

---

Towards topological type theory for decrypting transfinite methods in classical mathematics

In constructive mathematics, an ongoing challenge is to extracting
computational content from infinitary arguments in classical
mathematics—especially those that yield concrete results via abstract tools
such as minimal bad sequences or maximal ideals. Today, a wide range of such
techniques exists, constituting a late partial fulfillment of Hilbert's
program.

This talk presents work in progress on porting the modal operators "everywhere"
and "somewhere", as developed in set theory by Joel David Hamkins, Victoria
Gitman and their collaborators, into type theory. This modal enrichment
provides a uniform language for expressing several established
constructivization techniques, and is at the core of a relatively recent new
technique that offers insights on the origins of some of our most cherished
inductive definitions. These developments are closely connected to recent
advances in type-theoretic presheaf and sheaf models, dialogue
continuity, synthetic algebraic geometry and well-quasi-order theory.

---

Which functions are we excluding when we form the type of all functions between
two types? Does every field have an algebraic closure?

Inspired by the modal approach to the set-theoretic multiverse of Joel David
Hamkins, Victoria Gitman and their collaborators, we aim to introduce the modal
operators "everywhere" and "somewhere" to homotopy type theory, proposing a
multiversal perspective on these motivating questions.

---

Well quasi-orders are a combinatorial notion with featureful applications in
graph theory, termination checking, commutative algebra and other subjects.
However, their classical definition, though concise and elegant, poses
significant challenges in constructive mathematics and frameworks that lack
function sets. In response, several constructive substitutes have been
developed, most recently an implicational definition by Stefano Berardi,
Gabriele Buriola and Peter Schuster. In this talk, we investigate how a modal
approach can reinterpret classical proofs involving transfinite methods as
blueprints for constructive proofs grounded in these alternatives, thereby
combining the best of both worlds: Short and abstract proofs, but with
constructive content. We also indicate how the modal approach allows us to
compare the implicational definition with an earlier inductive one.

---

Combinatorics and commutative algebra abound with situations where we prove
quite concrete results by quite abstract transfinite methods, such as minimal
bad sequences or maximal ideals. Amazingly, such infinitary arguments can often
be understood as blueprints for quite explicit computations—as called for by
Hilbert’s programme. In this talk, we will travel the modal toposophic
multiverse to facilitate this kind of mining abstract proofs for refined
quantitative results. We will use a celebrated theorem from order theory about
sequences N→N that everybody can relate with as a running example.

---

Where do some of our most cherished inductive definitions come from? Which
functions are we excluding when we form the type of all functions between two
types? Does every field have an algebraic closure?

Inspired by the modal approach to the set-theoretic multiverse of Joel David
Hamkins, Victoria Gitman and their collaborators, we aim to introduce the modal
operators "everywhere" and "somewhere" to homotopy type theory, proposing a
multiversal perspective on these motivating questions.

Unlike the set-theoretic role model, we focus less on exploring the range of
foundational possibility and more on concrete applications in constructive
mathematics, with the goal of porting results of classical mathematics to
homotopy type theory. It will turn out that every field has an algebraic
closure somewhere; that a transitive relation is well-founded iff nowhere there
is an infinite descending chain; and that somewhere, the law of excluded middle
holds.

This is ongoing joint work with Alexander Oldenziel and connected to the recent
advances with type-theoretic presheaf models and sheaf models.